### 2020-05-18 docs/paper/LaTex/Paper.tex
Sorting files and folders again

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   8888888b.                                         888      888
%   888   Y88b                                        888      888
%   888    888                                        888      888
%   888   d88P 888d888 .d88b.   8888b.  88888b.d88b.  88888b.  888  .d88b.
%   8888888P"  888P"  d8P  Y8b     "88b 888 "888 "88b 888 "88b 888 d8P  Y8b
%   888        888    88888888 .d888888 888  888  888 888  888 888 88888888
%   888        888    Y8b.     888  888 888  888  888 888 d88P 888 Y8b.
%   888        888     "Y8888  "Y888888 888  888  888 88888P"  888  "Y8888

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% vscode extension LaTex-Workshop uses these magic comments to 
% define which program will be used to build the project or bibliography
% I do need lualatex to use markdown package. Also, when I use lualatex
% in need to include `pdfencoding=auto` in hyperref setup.
% !TEX program = lualatex
%% !TEX options = -synctex=1 -interaction=nonstopmode -shell-scape "%DOC%"
% `-shell-scape` argument doesn't work in Windows 10 lualatex
% !TEX options = -synctex=1 -interaction=nonstopmode "%DOC%"
% !BIB program = biber

\documentclass[11pt,a4paper,dvipsnames,twoside]{article}

% Paper geometry
\usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
% \usepackage[showframe,a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}

% inputenc is ignored when I'm using lualatex as builder
% \usepackage[utf8]{inputenc}

% To produce a spanish automated output, for instance, TOC is called 'Índice' and
% References is called 'Referencias'
\usepackage[english]{babel}

% Modifier for space between lines (=\baselinestretch*\baselineskip)
% \renewcommand{\baselinestretch}{1.5}
% Modifies the space between notes in footnotes 
% IMPORTANT: just the same than paragraphs (parskip package below)
% \setlength{\footnotesep}{.9\baselineskip}

% It is better to use setspace package
\usepackage{setspace}
\setstretch{1.5}

% Modifies the space over footnote line
\addtolength{\skip\footins}{\baselineskip}

% To convert fonts already installed in the system and using them as main font in the document
% To check the names that fontspec admits use `otfinfo -i </path/to/font/file/even.ttf>`
% otfinfo does admit also ttf files! For instance:
% `otfinfo -i /usr/share/fonts/liberation/LiberationSerif-Regular.ttf`
\usepackage{fontspec}
\setmainfont{Liberation Serif}
\setsansfont{Liberation Sans}
\setmonofont{CourierPrimeCode-Regular} 
% \setmonofont{Liberation Mono}
% A different font for titles
\newfontfamily\headingfont[]{Liberation Sans}

% To allow custom headings
\usepackage[compact]{titlesec}
\titleformat*{\section}{\Large\headingfont\bfseries}
\titleformat*{\subsection}{\large\headingfont\bfseries}
\titleformat*{\subsubsection}{\normalsize\headingfont\bfseries}
% To control spacing for sections, I'm not using it, it just an example
% \titlespacing*{\section}{3em}{0em}{-2em}

% To change title font family
% \usepackage{titling}
% \renewcommand{\maketitlehooka}{\headingfont}

% To change TOC appearance; see 2.1 section in tocloft.pdf for info about 'titles' option
% it does not mention fancyhdr package but it do the work to flushright page numbers for 
% table of contents pages. see: https://tex.stackexchange.com/a/137301/83189
% \usepackage[titles]{tocloft}
\usepackage{tocloft}
\renewcommand{\cftsecfont}{\headingfont\bfseries}
\renewcommand{\cftsecpagefont}{\headingfont\bfseries}
\renewcommand{\cftsubsecfont}{\headingfont}
\renewcommand{\cftsubsecpagefont}{\headingfont}
\renewcommand{\cftsubsubsecfont}{\headingfont}
\renewcommand{\cftsubsubsecpagefont}{\headingfont}
\tocloftpagestyle{fancy} % However, I find better this way bc 'titles' option does not 
% produce a compact TOC, it does expand bc probably it uses the same spacing rules in 
% the TOC than in (subsub)(sub)sections titles at the rest of the text.
% -------------------
% Remove the TOC title (it is redundant if you have included a TOC section inside TOC)
\makeatletter
\renewcommand{\@cftmaketoctitle}{}
\makeatother
% But in the case you still want to print the TOC title, this other one modifies TOC title font
\renewcommand{\cfttoctitlefont}{\Large\headingfont\bfseries}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   8888888b.                   888
%   888   Y88b                  888
%   888    888                  888
%   888   d88P 8888b.   .d8888b 888  888  8888b.   .d88b.   .d88b.  .d8888b
%   8888888P"     "88b d88P"    888 .88P     "88b d88P"88b d8P  Y8b 88K
%   888       .d888888 888      888888K  .d888888 888  888 88888888 "Y8888b.
%   888       888  888 Y88b.    888 "88b 888  888 Y88b 888 Y8b.          X88
%   888       "Y888888  "Y8888P 888  888 "Y888888  "Y88888  "Y8888   88888P'
%                                                      888
%                                                 Y8b d88P
%                                                  "Y88P"

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Basic packages (they don't need extra configurations)
\usepackage{% `esvect` will not work without texlive-fontsextra
  amsmath,
  amsfonts,
  amssymb,
  graphicx,
  multirow,
  tabularx,
  mathtools,
  caption,
  subcaption,
  color
}

% Extra colors
\usepackage[x11names,table]{xcolor}

% To cross references and metadata
\usepackage[]{hyperref}
\hypersetup{%
  pdfencoding=auto,
  pdftitle={Degree Final Project},
  pdfauthor={Javier Fernández // jfernandil@alumnos.unex.es},
  pdfsubject={},
  pdfkeywords={Project, Dendrometry, LoRa, Arduino, IoT},
  colorlinks = false,
  linkbordercolor = DarkGoldenrod2,
  urlbordercolor= DarkGoldenrod2,
  %linkcolor = black,
  %urlcolor  = black,
  breaklinks
}

% To make the frame around UEx seal
\usepackage{mdframed}

% To include References and TOC in the TOC
\usepackage[]{tocbibind}

% To avoid 'underfull box' warning caused because bibliography line-break
\usepackage{etoolbox}
\apptocmd{\thebibliography}{\raggedright}{}{}

% To use \begin{markdown}\end{markdown} environment and be able to type markdown mode
% IMPORTANT: do not use indent inside \begin{markdown}\end{markdown} environment in the code! 
% if you do lualatex is not translating the markdown code.
\usepackage{markdown}

% To get proper skips between paragraphs when you left a blank line in the code! (so I don't need to use `\\` 
% and I'm not getting warnings all the time) 
% IMPORTANT:
% here is an interesting article speaking about paragraphs an indentation, there are two types of paragraphs; 
% - classic: using indentation to distinguish between one and another
% - modern: dispenses with indents and instead distinguish paragraphs using a blank line between them
% Commenting/uncommenting the following package you can choose between two types.
% more info here: https://tex.stackexchange.com/a/196941/83189
% However this package also has options, which can be
% `indent` (to enable indentation)
% `skip=.5\baselineskip` (by default)
% `parfill` (to impose a minimum space at the end of the last line of a paragraph =30pt by default)
% \usepackage[skip=\baselineskip]{parskip}
% Another different option
% https://tex.stackexchange.com/a/371986/83189
\parskip=.75\baselineskip
\parindent=0pt

% To make quotes (I need renewcommand for mkbegdispquote and mkenddispquote to enable quote marks at the
% end and beginning to displayquote environment. More info: https://tex.stackexchange.com/questions/101809/how-to-force-csquotes-to-show-the-quotations-marks-in-block-quotes)
\usepackage[style=english]{csquotes}
\renewcommand{\mkbegdispquote}[2]{\openautoquote}
\renewcommand{\mkenddispquote}[2]{\closeautoquote#1#2}
\renewcommand{\mkcitation}[1]{ #1}

% To print units
\usepackage{siunitx}

% Enumerate lists
\usepackage{enumitem}
% To fix spaces between paragraphs, items and top or bottom fo enumerate/itemize environments I need the following
% see page 3 in https://osl.ugr.es/CTAN/macros/latex/contrib/enumitem/enumitem.pdf
\setlist[itemize]{%
  parsep=.75\baselineskip,
  topsep=0pt,
  partopsep=0pt,
  itemsep=0pt
}
\setlist[enumerate]{%
  parsep=.75\baselineskip,
  topsep=0pt,
  partopsep=0pt,
  itemsep=0pt
}

% To backrefs in the footnotes
\usepackage{footnotebackref}

% To improve headers an footers (I'm using to set page number at right)
\usepackage{fancyhdr}
% Turn on the style
\pagestyle{fancy}
\fancyhf{} % Start with clearing everything in the header and footer
\renewcommand{\headrulewidth}{0pt} % Removes the horizontal rule in header by default
\fancyfoot[R]{\thepage} % Set the right side of the footer to be the page number

% To insert json snippets
\usepackage{listings}
\renewcommand{\lstlistingname}{Ejemplo}

% \colorlet{punct}{red!60!black}
% \definecolor{background}{HTML}{EEEEEE}
% \definecolor{delim}{RGB}{20,105,176}
% \colorlet{numb}{magenta!60!black}
\definecolor{punct}{HTML}{0064d3}
\colorlet{background}{gray!20}
\definecolor{delim}{HTML}{ff0000}
\definecolor{numb}{HTML}{0f9d58}

\lstdefinelanguage{json}{
  % aboveskip=\baselineskip,
  % xleftmargin=2em,
  basicstyle=\scriptsize\ttfamily,
  numbers=left,
  numberstyle=\scriptsize,
  stepnumber=1,
  numbersep=8pt,
  showstringspaces=false,
  breaklines=true,
  frame=lines,
  backgroundcolor=\color{background},
  literate=
   *{0}{{{\color{numb}0}}}{1}
    {1}{{{\color{numb}1}}}{1}
    {2}{{{\color{numb}2}}}{1}
    {3}{{{\color{numb}3}}}{1}
    {4}{{{\color{numb}4}}}{1}
    {5}{{{\color{numb}5}}}{1}
    {6}{{{\color{numb}6}}}{1}
    {7}{{{\color{numb}7}}}{1}
    {8}{{{\color{numb}8}}}{1}
    {9}{{{\color{numb}9}}}{1}
    {:}{{{\color{punct}{:}}}}{1}
    {,}{{{\color{punct}{,}}}}{1}
    {"}{{{\color{punct}{"}}}}{1}
    {\{}{{{\color{delim}{\{}}}}{1}
    {\}}{{{\color{delim}{\}}}}}{1}
    {[}{{{\color{delim}{[}}}}{1}
    {]}{{{\color{delim}{]}}}}{1},
}

\newcommand\YAMLcolonstyle{\color{red}\mdseries}
\newcommand\YAMLkeystyle{\color{black}\mdseries}
\newcommand\YAMLvaluestyle{\color{blue}\mdseries}
% here is a macro expanding to the name of the language
% (handy if you decide to change it further down the road)
\lstdefinelanguage{yaml}{
  keywords={true,false,null,y,n},
  keywordstyle=\color{darkgray}\bfseries,
  basicstyle=\YAMLkeystyle\scriptsize\ttfamily,
  numbers=left,
  numberstyle=\scriptsize,
  numbersep=8pt,
  breaklines=true,
  frame=lines,
  backgroundcolor=\color{background},
  sensitive=false,
  comment=[l]{\#},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\YAMLvaluestyle\ttfamily,
  moredelim=[l][\color{orange}]{\&},
  moredelim=[l][\color{magenta}]{*},
  moredelim=**[il][\YAMLcolonstyle{:}\YAMLvaluestyle]{:},   % switch to value style at :
  morestring=[b]',
  morestring=[b]",
  literate = {---}{{\ProcessThreeDashes}}3
             {>}{{\textcolor{red}\textgreater}}1     
             {|}{{\textcolor{red}\textbar}}1 
             {\ -\ }{{\mdseries\ -\ }}3,
}

% To highlight text with automated line break or 'Overfull \hbox' messages
\usepackage{soul}
\soulregister\enquote7

% biblatex package
\usepackage[block=ragged,style=ieee]{biblatex}
\addbibresource{Paper.bib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%    .d8888b.                                                              888
%   d88P  Y88b                                                             888
%   888    888                                                             888
%   888         .d88b.  88888b.d88b.  88888b.d88b.   8888b.  88888b.   .d88888 .d8888b
%   888        d88""88b 888 "888 "88b 888 "888 "88b     "88b 888 "88b d88" 888 88K
%   888    888 888  888 888  888  888 888  888  888 .d888888 888  888 888  888 "Y8888b.
%   Y88b  d88P Y88..88P 888  888  888 888  888  888 888  888 888  888 Y88b 888      X88
%    "Y8888P"   "Y88P"  888  888  888 888  888  888 "Y888888 888  888  "Y88888  88888P'

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Provides the \doubt{} command
\newcommand{\doubt}[1] {\textbf{\color{red}#1}}

% Provides the \cmd{} command; using soul package and previous \letcolor
% definition for listings package. Read page 3 of soul.pdf or 
% https://tex.stackexchange.com/a/231647/83189
% I need that to get a correct and automated line break when the command I want
% to highlight is too long and is placed at the end of the line.
\sethlcolor{background}
\newcommand{\cmd}[1] {\hl{\texttt{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   88888888888 d8b 888    888
%       888     Y8P 888    888
%       888         888    888
%       888     888 888888 888  .d88b.
%       888     888 888    888 d8P  Y8b
%       888     888 888    888 88888888
%       888     888 Y88b.  888 Y8b.
%       888     888  "Y888 888  "Y8888

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{\headingfont%
  University of Extremadura\\
  Faculty of Science\\[1cm]
  \textbf{%
  	Physics degree\\
	  Degree Final Project\\[3cm]
	  \flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm]
  }
}

% I've used that \parbox to flush right the author, but I must have use this package https://osl.ugr.es/CTAN/macros/latex/contrib/titling/titling.pdf
\author{%
  \parbox{.9\textwidth}{
    \begin{flushright}
      Javier Fernández Aparicio\\
      % $\blacktriangledown$\\
      \href{mailto:jfernandil@alumnos.unex.es}{jfernandil@alumnos.unex.es}\\[\dimexpr\baselineskip + 1cm]
      July 2020
    \end{flushright}
  }
}
\date{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%   888888b.                     d8b               8888888b.
%   888  "88b                    Y8P               888  "Y88b
%   888  .88P                                      888    888
%   8888888K.   .d88b.   .d88b.  888 88888b.       888    888  .d88b.   .d8888b
%   888  "Y88b d8P  Y8b d88P"88b 888 888 "88b      888    888 d88""88b d88P"
%   888    888 88888888 888  888 888 888  888      888    888 888  888 888
%   888   d88P Y8b.     Y88b 888 888 888  888      888  .d88P Y88..88P Y88b.
%   8888888P"   "Y8888   "Y88888 888 888  888      8888888P"   "Y88P"   "Y8888P
%                            888
%                       Y8b d88P
%                        "Y88P"

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% To manage hyphenation (words break at end of the line), in this case is to prevent "Tellink" breakage.
\hyphenation{Tellink}

\begin{document}
\noindent % to prevent the indentation before first minipage and be able to expand minipages
\begin{minipage}[t]{.28\textwidth}
    \centering
      \begin{mdframed}[innerbottommargin=490pt, innertopmargin=40pt, linewidth=1pt]
        \includegraphics[width=\textwidth]{../../pictures/Seal/marca-uex-2-color.png}
      \end{mdframed}
\end{minipage}
\hskip 30pt minus 5pt % It is a dynamic space (glue) that starts on 30pt and is 5pt shrinkable
\begin{minipage}[t]{.6\textwidth}
  \centering
  \vspace{1cm}
  \maketitle
\end{minipage}
\thispagestyle{empty}
\newpage
\begingroup
  \thispagestyle{empty}
  \null%
\endgroup
\newpage
% \pagenumbering{arabic}% Arabic page numbers (and reset to 1)
\setcounter{page}{1}
\tableofcontents

\newpage

\section{Abstract}
This document gives a detailed description of this project, which is focused on researching possibles low-cost alternatives for wireless dendrometry systems. Currently there exist a lot of expensive and professional systems in the market, that's because this project is intended to reduce costs and increase the versatility, scalability and accessibility.

In order to reach these objetives the project will be supported over free software such as FIWARE\cite{Fiware} or free hardware such as Arduino\cite{Arduino} and RaspberryPi\cite{Raspberrypi}.

\printbibliography
\end{document}

### 2020-05-22 docs/paper/LaTex/Paper.tex
Includes testing folder and few changes in Paper.tex

@@
     Degree Final Project\\[3cm]
-	  \flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm]
+    \flushleft{\large{Developement of a FIWARE-based application for tree species monitoring (dendrometry)}}\\[3cm]
+    %\flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm]
   }
@@
 
-\section{Abstract}
+\section*{Abstract}
+\addcontentsline{toc}{section}{Abstract}%
 This document gives a detailed description of this project, which is focused on researching possibles low-cost alternatives for wireless dendrometry systems.
@@
 
+\section{Introduction}
+This project arises itself from a direct interaction with professionals inside forestal sector.
+The original idea was to give technical coverage for particular necessities which professionals in this sector had to face off with.
+At this point is easy to notice this solution will need to be a distributed solution, due high samples dispersion.
+
 \printbibliography

### 2020-05-30 docs/paper/LaTex/Paper.tex
Modifies bibliography structure to prevent URL breakage
Completing main docu

@@
 % biblatex package
-\usepackage[block=ragged,style=ieee]{biblatex}
+% 'block=nbpar' produces a new paragraph for every block (printing the URLs in a new line and preventing breaking short URLs)
+% page 50 in https://osl.ugr.es/CTAN/macros/latex/contrib/biblatex/doc/biblatex.pdf
+\usepackage[block=par,style=ieee]{biblatex}
 \addbibresource{Paper.bib}
@@
 % Provides the \doubt{} command
-\newcommand{\doubt}[1] {\textbf{\color{red}#1}}
+\newcommand{\doubt}[1] {\textbf{\color{Red3}#1}}
+% Provides \lang{} command to mark language doubts
+\newcommand{\lang}[1] {\textbf{\color{Tomato1}#1}}
 
@@
 At this point is easy to notice this solution will need to be a distributed solution, due high samples dispersion.
+\lang{As can be seen}, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts \doubt{between $157$-$170$ specimens per hectare}\cite{ForestStandVol} (depending on the used model).
+So according to this required resolution it will be needed a big wireless network of distributed devices.
+
+This is more or less, the definition of the IoT (Internet of Things) concept; according to \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
+However, also according to \cite{IoTOverview} there is not an official definition for IoT concept, but
+
+\begin{displayquote}[\cite{IoTObjetives}]
+  based on the traditional information carriers including the Internet, telecommunication  network  and  so  on,  Internet  of Things (IoT) is a network that interconnects ordinary physical  objects  with  the identifiable addresses  so that provides intelligent services.
+\end{displayquote}
 

### 2020-06-01 docs/paper/LaTex/Paper.tex
Completing documentation

@@
 % Paper geometry
-\usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
-% \usepackage[showframe,a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
+% \usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
+\usepackage[showframe,a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
 
@@
 
+% biblatex package
+% 'block=nbpar' produces a new paragraph for every block (printing the URLs in a new line and preventing breaking short URLs)
+% page 50 in https://osl.ugr.es/CTAN/macros/latex/contrib/biblatex/doc/biblatex.pdf
+\usepackage[block=par,style=ieee,backend=biber]{biblatex}
+\addbibresource{Paper.bib}
+\setcounter{biburllcpenalty}{7000}
+\setcounter{biburlucpenalty}{7000}
+\setcounter{biburlnumpenalty}{7000}
+
 % Basic packages (they don't need extra configurations)
@@
   subcaption,
-  color
+  color,
+  url
 }
@@
 
-% biblatex package
-% 'block=nbpar' produces a new paragraph for every block (printing the URLs in a new line and preventing breaking short URLs)
-% page 50 in https://osl.ugr.es/CTAN/macros/latex/contrib/biblatex/doc/biblatex.pdf
-\usepackage[block=par,style=ieee]{biblatex}
-\addbibresource{Paper.bib}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 At this point is easy to notice this solution will need to be a distributed solution, due high samples dispersion.
-\lang{As can be seen}, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts \doubt{between $157$-$170$ specimens per hectare}\cite{ForestStandVol} (depending on the used model).
-So according to this required resolution it will be needed a big wireless network of distributed devices.
+\lang{As can be seen}, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts \doubt{between $157$-$170$ indviduals per hectare}\cite{ForestStandVol} (depending on the used model).
+So according to this and sample size determination theories, to get a great resolution could be necessary a big size for samples and the necessity of a big wireless network of distributed devices, since each device will correspond with an individual.
 
@@
 
+This, at least, covers a little part what this project is intended to do: \enquote{Interconnect ordinary physical objects with the identifiable adresses} to provide intelligent services.
+These physical objects are in this case ordinary dendrometers.
+
+Over the years there have existed analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way.
+This could turn out bothering because the big size for this statistical population, \lang{as it was exposed before}.
+So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.
+
+\subsection{Dendrometry, a formal definition}
+The GEMET (General Multilingual Environmental Thesaurus) adopts the definition for \textit{dendrometry} from \cite{DictNaturalRes}
+
+\begin{displayquote}
+  The measuring of the diameter of standing trees from the ground with a dendrometer that can also be used to measure tree heights.
+\end{displayquote}
+
+This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point could be interesting to extend this project to include also a sensor to heights measurement.\footnote{This project is already considered extensive enough.}
+
+There are a lot of comercial dendrometers in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently as has been already said, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
+So this is intended to get a cheaper system and make it accessible to everyone who wants \lang{to monitorize} one or more trees growth.
+
+\subsection{Arduino, multipurpose microcontroller}
+It's not difficult to justify the use of such an interesting platform as Arduino.
+The adaptability is one of its strengths, therefore; it is able to acquire and process certain data coming from a set of sensors and manage it to send this via any plugged wireless network interface.
+
+Due circumstances exposed in introduction section is needed an accessible and multipurpose platform \lang{to be the basis for the device design itself}, this is to say the core part of the dendrometer is going to be an Arduino microcontroller.
+
 \printbibliography

### 2020-06-02 docs/paper/LaTex/Paper.tex
Completing docs

@@
 
-There are a lot of comercial dendrometers in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently as has been already said, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
+There are a lot of comercial dendrometry systems in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently \lang{¿like/as?} has been already said, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
 So this is intended to get a cheaper system and make it accessible to everyone who wants \lang{to monitorize} one or more trees growth.
+
+There are plenty types of dendrometers but according to \cite{DendroResearch} \enquote{It is possible to define two broad categories of dendrometer: those that contact the stem and those that do not}.
+This project is focused on the former kind, so for this project is being developed a \enquote{contact dendrometer} based on a linear potentiometer.
 
@@
 
-Due circumstances exposed in introduction section is needed an accessible and multipurpose platform \lang{to be the basis for the device design itself}, this is to say the core part of the dendrometer is going to be an Arduino microcontroller.
+Due circumstances exposed in introduction section is needed an accessible and multipurpose platform \lang{to be the basis for the device design itself}, this is to say the core part of the dendrometer is an Arduino microcontroller.
+
+Since the idea is to produce a low-cost device in order to distribute a high number of them, it must be a simple design; that's because it consist only of three parts,
+
+\begin{itemize}
+  \item \href{https://es.rs-online.com/web/p/products/0317780/}{\doubt{Linear potentiometer:}} which is the sensor itself due is directly in contact with the stem.
+In order \doubt{to improve data acquisition it will be necessary to use a High Input Impedance Amplifier.}
+  \item Arduino microcontroller: this is the core part for the device, it will be responsible for acquire linear potentiometer data and send it to a gateway through LoRa interface. 
+  \item LoRa interface: similar to other existing wireless interfaces  
+\end{itemize}
 

### 2020-06-03 docs/paper/LaTex/Paper.tex
Completing docs

@@
 \subsection{Dendrometry, a formal definition}
-The GEMET (General Multilingual Environmental Thesaurus) adopts the definition for \textit{dendrometry} from \cite{DictNaturalRes}
+The GEMET (General Multilingual Environmental Thesaurus) adopts the definition for \textit{dendrometry} from \cite{DictNaturalRes}:
 
@@
 
-There are a lot of comercial dendrometry systems in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently \lang{¿like/as?} has been already said, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
+A lot of comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently \lang{¿like/as?} has been already said, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
 So this is intended to get a cheaper system and make it accessible to everyone who wants \lang{to monitorize} one or more trees growth.
@@
   \item Arduino microcontroller: this is the core part for the device, it will be responsible for acquire linear potentiometer data and send it to a gateway through LoRa interface. 
-  \item LoRa interface: similar to other existing wireless interfaces  
+  \item LoRa interface: similar to other existing wireless interfaces, it is necessary to forward the sensor data to a concentrator (gateway).
+Usually and due its complexity these kind of interfaces are integrated circuits which are mounted on a PCB in order to obtain a pluggable card/shield.
 \end{itemize}
+
+\subsection{LoRa (Long Range)}
+LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite{LoRaGeneral}.
+This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
+So this low availability along to their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
+
+In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN)
+
+\begin{displayquote}[\cite{LoRaGeneral}]
+  Colloquially  speaking,  an LPWAN is supposed to be to the IoT what WiFi was to consumer networking: offering radio coverage over a (very) large area by way of base stations and adapting transmission rates, transmission power, modulation, duty cycles, etc., such that end-devices incur a very low energy consumption due to their being connected.
+\end{displayquote}
+
+It is important to note that \lang{when talking about \enquote{low-power consumption}, in many cases it is actually meaning} battery-powered devices, for example.
+
+By other hand, Lora can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
+The former one (the physical layer), is a proprietary technology developed by Semtech.
 

### 2020-06-04 docs/paper/LaTex/Paper.tex
Completing docs

@@
 
-By other hand, Lora can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
+By other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
 The former one (the physical layer), is a proprietary technology developed by Semtech.
+So this does mean this layer is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer,  It is intended to sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite{LoRaGeneral}.
+This particularly means that LoRaWAN protocol is perfect for the purpose of this project.
+
+\subsection{RaspberryPi, a powerful microcomputer}
 

### 2020-06-05 docs/paper/LaTex/Paper.tex
Completing docs

@@
 % Paper geometry
-% \usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
-\usepackage[showframe,a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
+\usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
+% \usepackage[showframe,a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
 
@@
 \subsection{RaspberryPi, a powerful microcomputer}
+As in the case of Arduino, RaspberryPi provides a powerful platform, however in the case of a RaspberryPi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, RaspberryPi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a RaspberryPi to support a whole operating system; that's because this devices are usually called \textit{single-board computers}.
+
+A few of interesting hardware specs are for instance, a \textit{Broadcom BCM2837B0, Cortex-A53 (ARMv8) 64-bit SoC @ 1.4GHz} as its CPU, \textit{1GB LPDDR2 SDRAM}, \textit{Gigabit Ethernet over USB 2.0 (maximum throughput 300 Mbps)} or even an \textit{Extended 40-pin GPIO header} among others.\cite{RPiSpecs}
+
+The role for RaspberryPi in this project is to act like a gateway, receiving all the data from different nodes.
+So in order to perform this task correctly it is necessary to provide a LoRa interface, which is also embebed (as in Arduino) in a separate pluggable shield/card/hat.
+
+\subsection{Dragino shields/hats}
 

### 2020-06-06 docs/paper/LaTex/Paper.tex
Completing docs

@@
 \subsection{Dragino shields/hats}
+These shields/expansion cards are necessary because those devices (Arduino and RaspberryPi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages.
+Both shields are based on the SX1276/SX1278 transceiver.
+However the RaspberryPi hat also has a L80 GPS interface (Base on MTK MT3339), meanwhile Arduino has not.
+
+This project is using the following models:
+
+\begin{itemize}
+  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, so this means this hat is designed to be a node-class radio, not a gateway.}
+  \item \textit{LoRa Shield for Arduino}\cite{DraginoArdShield}, which is plugged through analog and digital pins.
+\end{itemize}
+
+\section{Background}
 

### 2020-06-06 docs/paper/LaTex/Paper.tex
Fixing links and magic comments
- Magic comments for VScode has been clarified

- \phantomsection was added to Abstract section in order
to get its TOC link working

- TTN forum cite url was improved and also has been
introduced a question about this cite usage itself

@@
 % !TEX program = lualatex
+
+% `-shell-scape` argument doesn't work in Windows 10 lualatex
 %% !TEX options = -synctex=1 -interaction=nonstopmode -shell-scape "%DOC%"
-% `-shell-scape` argument doesn't work in Windows 10 lualatex
 % !TEX options = -synctex=1 -interaction=nonstopmode "%DOC%"
@@
 
-\section*{Abstract}
+\phantomsection\section*{Abstract}
 \addcontentsline{toc}{section}{Abstract}%
@@
 \begin{itemize}
-  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, so this means this hat is designed to be a node-class radio, not a gateway.}
+  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack},\doubt{(I'm not sure about this citation (It's a forum))} so this means this hat is designed to be a node-class radio, not a gateway.}
   \item \textit{LoRa Shield for Arduino}\cite{DraginoArdShield}, which is plugged through analog and digital pins.

### 2020-06-08 docs/paper/LaTex/Paper.tex
Completing docs

@@
 \usepackage[style=english]{csquotes}
-\renewcommand{\mkbegdispquote}[2]{\openautoquote}
-\renewcommand{\mkenddispquote}[2]{\closeautoquote#1#2}
-\renewcommand{\mkcitation}[1]{ #1}
+% \renewcommand{\mkbegdispquote}[2]{\openautoquote}
+% \renewcommand{\mkenddispquote}[2]{\closeautoquote#1#2}
+% \renewcommand{\mkcitation}[1]{ #1}
+
+% Switching to quoting package (for block quotes, I still need csquotes for \enquote{})
+\usepackage[begintext=``, endtext='']{quoting}
+\quotingsetup{vskip=5pt}
 
@@
 This is more or less, the definition of the IoT (Internet of Things) concept; according to \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
-However, also according to \cite{IoTOverview} there is not an official definition for IoT concept, but
+However, also according to \cite{IoTOverview} there is not an official definition for IoT concept, but according to \cite{IoTObjetives}
 
-\begin{displayquote}[\cite{IoTObjetives}]
+\begin{quoting}
   based on the traditional information carriers including the Internet, telecommunication  network  and  so  on,  Internet  of Things (IoT) is a network that interconnects ordinary physical  objects  with  the identifiable addresses  so that provides intelligent services.
-\end{displayquote}
+\end{quoting}
 
@@
 
-\begin{displayquote}
+\begin{quoting}
   The measuring of the diameter of standing trees from the ground with a dendrometer that can also be used to measure tree heights.
-\end{displayquote}
+\end{quoting}
 
@@
 
-\subsection{Arduino, multipurpose microcontroller}
+\subsection{Hardware}
+
+\subsubsection{Arduino, multipurpose microcontroller}
 It's not difficult to justify the use of such an interesting platform as Arduino.
@@
 
-\subsection{LoRa (Long Range)}
+\subsubsection{LoRa (Long Range)}
 LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite{LoRaGeneral}.
@@
 
-In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN)
+In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), so as claimed by \cite{LoRaGeneral}
 
-\begin{displayquote}[\cite{LoRaGeneral}]
+\begin{quoting}
   Colloquially  speaking,  an LPWAN is supposed to be to the IoT what WiFi was to consumer networking: offering radio coverage over a (very) large area by way of base stations and adapting transmission rates, transmission power, modulation, duty cycles, etc., such that end-devices incur a very low energy consumption due to their being connected.
-\end{displayquote}
+\end{quoting}
 
@@
 
-\subsection{RaspberryPi, a powerful microcomputer}
+\subsubsection{RaspberryPi, a powerful microcomputer}
 As in the case of Arduino, RaspberryPi provides a powerful platform, however in the case of a RaspberryPi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, RaspberryPi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a RaspberryPi to support a whole operating system; that's because this devices are usually called \textit{single-board computers}.
@@
 
-\subsection{Dragino shields/hats}
+\subsubsection{Dragino shields/hats}
 These shields/expansion cards are necessary because those devices (Arduino and RaspberryPi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages.
@@
 
+\subsection{Software}
+
+\subsubsection{FIWARE}
+It is defined as \enquote{The Open Source Platform for Our Smart Digital Future}\cite{Fiware}; in a deeper sense is \enquote{an open source initiative defending an universal set of standards for context data management}\cite{Fiware}, which basically means that \textit{FIWARE} is an open source platform where develop IoT and smart solutions.
+The platform provides a number of software modules called \textit{Generic Enablers}; among all those \textit{Generic Enablers} there is one particularly important called \textit{Orion Context Broker}.
+In fact, for a solution to be considered as \textit{Powered by FIWARE} it must to use the \textit{Orion Context Broker} at least.
+
+This \textit{Context data} is just a way to name any kind of data coming from any kind of sensor.
+\textit{Orion Context Broker} is designed to manage this context data through concepts such as \textit{subscriptions} or \textit{entities}, to name but a few; however, all these \textit{Generics Enablers} communicate with each other using the \textit{FIWARE NGSI RESTful API}\cite{NGSI}.\footnote{\textit{RESTful} term comes from the software architectural style \textit{REST}, which stands for \textit{Representational State Transfer}}
+
 \section{Background}

### 2020-06-09 docs/paper/LaTex/Paper.tex
Modifying docs, schemes and taking notes

@@
 
+\section{Technical description}
+Like in the Introduction the most proper way to present this technical description is differencing between software and hardware parts.
+However, is still interesting to present a simple and general diagram for whole designed solution:
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=.9\textwidth]{../../schemes/main_scheme_tbg.png}
+  \caption{General view of this solution.}
+  \label{fig:GenView}
+\end{figure}
+
+\subsection{Hardware}
+
+\subsubsection{Nodes}
+Nodes are the parts of the system in direct contact with trees themselves (one node per each tree).
+Most important features in those devices are: low-cost, low-powered, wireless communication and small size.
+The following diagram shows a diagram of its most relevant parts:
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=.9\textwidth]{../../schemes/node_tbg.png}
+  \caption{Diagram for a detailed view of a node.}
+  \label{fig:NodeDiag}
+\end{figure}
+
+This diagram shows three different parts:
+
+\begin{itemize}
+  \item A linear potentiometer, ideally an \href{https://docs.rs-online.com/37bf/0900766b814f0bd0.pdf}{RS Pro Conductive Polymer
+  Potentiometer for AutomotiveApplications}, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, \doubt{this project will implement a regular rotary potentiometer.}
+  \item \doubt{A signal conditioning stage, which is needed in order to improve the quality of the voltage signal produced by the potentiometer.}
+  \item Arduino + Dragino LoRa Shield; of course it is necessary to schedule data transmission and acquisition, as well as transmit them using the LoRa physical layer and LoRaWAN protocol.
+For this purpose an Arduino-like microcontroller is perfect.
+\end{itemize}
+
 \printbibliography

### 2020-06-10 docs/paper/LaTex/Paper.tex
Completing docs

@@
 
+% To add an extra level in sections
+\usepackage{titlesec}
+\titleclass{\subsubsubsection}{straight}[\subsection]
+\newcounter{subsubsubsection}[subsubsection]
+\renewcommand\thesubsubsubsection{\thesubsubsection.\arabic{subsubsubsection}}
+\renewcommand\theparagraph{\thesubsubsubsection.\arabic{paragraph}} % optional; useful if paragraphs are to be numbered
+\titleformat{\subsubsubsection}
+  {\normalfont\normalsize\bfseries}{\thesubsubsubsection}{1em}{}
+\titlespacing*{\subsubsubsection}
+{0pt}{3.25ex plus 1ex minus .2ex}{1.5ex plus .2ex}
+\makeatletter
+\renewcommand\paragraph{\@startsection{paragraph}{5}{\z@}%
+  {3.25ex \@plus1ex \@minus.2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries}}
+\renewcommand\subparagraph{\@startsection{subparagraph}{6}{\parindent}%
+  {3.25ex \@plus1ex \@minus .2ex}%
+  {-1em}%
+  {\normalfont\normalsize\bfseries}}
+\def\toclevel@subsubsubsection{4}
+\def\toclevel@paragraph{5}
+\def\toclevel@paragraph{6}
+\def\l@subsubsubsection{\@dottedtocline{4}{7em}{4em}}
+\def\l@paragraph{\@dottedtocline{5}{10em}{5em}}
+\def\l@subparagraph{\@dottedtocline{6}{14em}{6em}}
+\makeatother
+\setcounter{secnumdepth}{4}
+\setcounter{tocdepth}{4}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \item A linear potentiometer, ideally an \href{https://docs.rs-online.com/37bf/0900766b814f0bd0.pdf}{RS Pro Conductive Polymer
-  Potentiometer for AutomotiveApplications}, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, \doubt{this project will implement a regular rotary potentiometer.}
+  Potentiometer for Automotive Applications}, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, \doubt{this project will implement a regular rotary potentiometer.}
   \item \doubt{A signal conditioning stage, which is needed in order to improve the quality of the voltage signal produced by the potentiometer.}
@@
 
+\subsubsubsection{Linear Potentiometer}
+As said, ideally this potentiometer should be a linear potentiometer with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it's recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that's because the figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
+
+\subsubsubsection{Signal conditioning}
+This stage is focused on improve the signal acquisition, according to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option.
+\doubt{Here is shown a possible schematics for this stage:}
+
+\begin{figure}[htp]
+  \centering
+    \includegraphics[width=.9\textwidth]{../../schemes/Signal_Conditioning.png}
+  \caption{Signal conditioning stage.}
+  \label{fig:SignalCond}
+\end{figure}
+
+\subsubsection{gateway}
+Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role.
+These must be generic devices due LoRaWAN protocol versatility.
+LoRaWAN is a cloud-based medium access control (MAC) layer protocol, but actually acts as a network layer protocol for managing communication between gateways and nodes, similar to a routing protocol.
+So it is possible for any device which implements hardware for a LoRa physical layer, to act as a gateway.
+
+Nevertheless, there are important considerations about the said above,
+
+\subsection{Software}
+
 \printbibliography

### 2020-06-11 docs/paper/LaTex/Paper.tex
Checking bibliography references and completing docs

@@
 
-This is more or less, the definition of the IoT (Internet of Things) concept; according to \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
-However, also according to \cite{IoTOverview} there is not an official definition for IoT concept, but according to \cite{IoTObjetives}
+This is more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
+However, also according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept, but according to \cite[p.~2(920)]{IoTObjetives}
 
@@
 Over the years there have existed analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way.
-This could turn out bothering because the big size for this statistical population, \lang{as it was exposed before}.
+This could turn out bothering because the big size for this statistical population, \lang{¿as/like?
+it was exposed ¿before/above?}.
 So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.
@@
 \subsubsection{LoRa (Long Range)}
-LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite{LoRaGeneral}.
+LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite[]{LoRaGeneral}.
 This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
@@
 
-In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), so as claimed by \cite{LoRaGeneral}
+In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), so as claimed by \cite[]{LoRaGeneral}
 
@@
 The former one (the physical layer), is a proprietary technology developed by Semtech.
-So this does mean this layer is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer,  It is intended to sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite{LoRaGeneral}.
+So this does mean this layer is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer,  It is intended to sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite[p.~9]{LoRaGeneral}.
 This particularly means that LoRaWAN protocol is perfect for the purpose of this project.
@@
 
-\subsubsection{Dragino shields/hats}
+\subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
 These shields/expansion cards are necessary because those devices (Arduino and RaspberryPi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages.
@@
 
-\subsubsection{gateway}
+\subsubsection{Gateway}
 Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role.
@@
 
-Nevertheless, there are important considerations about the said above,
+Nevertheless, there are important considerations about the said above, for example, nodes are not actually associated with an specific gateway.
+Instead, data transmitted by a node is typically received by multiple gateways.
+Each gateway will forward the received packets from the end-node to de cloud-based network server.
+Besides, this project is using a RaspberryPi 3B+ with a Dragino Hat which mounts a SX1276 LoRa \textbf{transceiver}\cite{SX1276}; this is so important because means, according to Semtech, this transceiver is not intended to play a gateway role, but a end-node role.
+
+A transceiver, by definition, is a device that is able to both, transmit and receive (in fact, the word itself is a mix between both, \textbf{trans}miter and re\textbf{ceiver}) that is what a node must be able to do; i.e.
+transmit the sensor data (context) and receive data to perform operations with its actuators.
+
+Nevertheless, LoRaWAN specification varies from region to region \enquote{based on
+the different regional spectrum allocations and regulatory requirements}\cite[p.~12-13]{LoRaWANspec}
 

### 2020-06-12 docs/paper/LaTex/Paper.tex
Customizes section titles and completes docs

@@
 
-\subsubsection{LoRa (Long Range)}
+% I need to use that \texorpdfstring{}{} for the left-side ToC in PDF viewers, because the ® cannot be rendered in there.
+\subsubsection{\texorpdfstring{LoRa\textsuperscript{\textregistered} (Lo{\normalfont\sffamily ng} Ra{\normalfont\sffamily nge})}{LoRa® (Long Range)}}
 LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite[]{LoRaGeneral}.
 This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
-So this low availability along to their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
+So this low availability (of resources) along to their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
 
@@
 This particularly means that LoRaWAN protocol is perfect for the purpose of this project.
+
+In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances:
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=.9\textwidth]{../../pictures/LoRa_Why_Range.png}
+  \caption{Bandwidth vs range plot for two of the most used telecommunication systems and LoRa.}
+  \label{}
+\end{figure}
 
@@
 
+\newpage
 \section{Background}
 
+\newpage
 \section{Technical description}
-Like in the Introduction the most proper way to present this technical description is differencing between software and hardware parts.
+Like in the Introduction, the most proper way to present this technical description is to difference between software and hardware parts.
 However, is still interesting to present a simple and general diagram for whole designed solution:
@@
 Nevertheless, LoRaWAN specification varies from region to region \enquote{based on
-the different regional spectrum allocations and regulatory requirements}\cite[p.~12-13]{LoRaWANspec}
+the different regional spectrum allocations and regulatory requirements}\cite[p.~12]{LoRaWANspec}. In fact, for Europe, and again as reported by \cite[p.~13]{LoRaWANspec}
+
+\begin{quoting}
+LoRaWAN defines ten channels, eight of which are multi data rate from 250bps to
+5.5 kbps, a single high data rate LoRa channel at 11kbps, and a single FSK channel
+at 50kbps.
+\end{quoting}
+
+Here is the important point.
+This Dragino HAT for the RaspberryPi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is not LoRaWAN compliant}.
+The main reason the SX1276 transceiver is not suitable to work as a gateway is that \doubt{it is actually a single channel transceiver}.
+Despite all of this, there still exist the possibility of use it as a gateway, because it is technically possible.
+
+Dragino foresees this and provides a \href{https://github.com/dragino/dual_chan_pkt_fwd}{dual channel controller}.
+The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.
 

### 2020-06-13 docs/paper/LaTex/Paper.tex
Fixing typos and completing docs

@@
 
-In order to reach these objetives the project will be supported over free software such as FIWARE\cite{Fiware} or free hardware such as Arduino\cite{Arduino} and RaspberryPi\cite{Raspberrypi}.
+In order to reach these objetives the project will be supported over free software such as FIWARE\cite{Fiware} or free hardware such as Arduino\cite{Arduino} and Raspberry Pi\cite{Raspberrypi}.
 
@@
 
-\subsubsection{RaspberryPi, a powerful microcomputer}
-As in the case of Arduino, RaspberryPi provides a powerful platform, however in the case of a RaspberryPi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, RaspberryPi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a RaspberryPi to support a whole operating system; that's because this devices are usually called \textit{single-board computers}.
+\subsubsection{Raspberry Pi, a powerful microcomputer}
+As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, Raspberry Pi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's because this devices are usually called \textit{single-board computers}.
 
@@
 
-The role for RaspberryPi in this project is to act like a gateway, receiving all the data from different nodes.
+The role for Raspberry Pi in this project is to act like a gateway, receiving all the data from different nodes.
 So in order to perform this task correctly it is necessary to provide a LoRa interface, which is also embebed (as in Arduino) in a separate pluggable shield/card/hat.
@@
 \subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
-These shields/expansion cards are necessary because those devices (Arduino and RaspberryPi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages.
+These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages.
 Both shields are based on the SX1276/SX1278 transceiver.
-However the RaspberryPi hat also has a L80 GPS interface (Base on MTK MT3339), meanwhile Arduino has not.
+However the Raspberry Pi hat also has a L80 GPS interface (Base on MTK MT3339), meanwhile Arduino has not.
 
@@
 \subsubsection{Gateway}
+
+\subsubsubsection{Raspberry Pi 3 B+}
+This model of Raspberry Pi is able to run almost every GNU/Linux distribution ported to ARM architecture, Raspberry Pi OS (previously called Raspbian)\cite{RaspberryPiOs} is in fact a Debian port to ARM architecture.
+After a little researching, it is possible to conclude that there is not other operative system which improves the performance of Raspberry Pi OS in a Raspberry Pi.
+Due this former argument this project is going to use Raspberry Pi OS.
+
+One of the most interesting features of Raspberry Pi is precisely that OS is installed and run from a microSD card, so the hardware is loading the operative system from this microSD card to the RAM directly, which improves notably the system load times, and increments its portability.
+
+\subsubsubsection{Dragino HAT}
 Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role.
@@
 Each gateway will forward the received packets from the end-node to de cloud-based network server.
-Besides, this project is using a RaspberryPi 3B+ with a Dragino Hat which mounts a SX1276 LoRa \textbf{transceiver}\cite{SX1276}; this is so important because means, according to Semtech, this transceiver is not intended to play a gateway role, but a end-node role.
+Besides, this project is using a Raspberry Pi 3B+ with a Dragino Hat which mounts a SX1276 LoRa \textbf{transceiver}\cite{SX1276}; this is so important because means, according to Semtech, this transceiver is not intended to play a gateway role, but a end-node role.
 
@@
 Here is the important point.
-This Dragino HAT for the RaspberryPi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is not LoRaWAN compliant}.
+This Dragino HAT for the Raspberry Pi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is not LoRaWAN compliant}.
 The main reason the SX1276 transceiver is not suitable to work as a gateway is that \doubt{it is actually a single channel transceiver}.

### 2020-06-13 docs/paper/LaTex/Paper.tex
Test few commands about lengths in docs and completes them

@@
   mathtools,
-  caption,
-  subcaption,
   color,
@@
 }
+
+% Improve captions in figures
+\usepackage[margin=2.5cm]{caption}
+\usepackage{subcaption}
+% \setlength{\textfloatsep}{1\baselineskip plus 2\baselineskip minus 2\baselineskip}
+% \setlength{\intextsep}{1\baselineskip plus 2\baselineskip minus 2\baselineskip}
+% \setlength{\textfloatsep}{1\baselineskip}
+\setlength{\intextsep}{.5\baselineskip}
 
@@
 
-There are plenty types of dendrometers but according to \cite{DendroResearch} \enquote{It is possible to define two broad categories of dendrometer: those that contact the stem and those that do not}.
+There are quite a few types of dendrometers but according to \cite{DendroResearch} \enquote{It is possible to define two broad categories of dendrometer: those that contact the stem and those that do not}.
 This project is focused on the former kind, so for this project is being developed a \enquote{contact dendrometer} based on a linear potentiometer.
@@
   \caption{Bandwidth vs range plot for two of the most used telecommunication systems and LoRa.}
-  \label{}
+  \label{fig:LoraComparison}
 \end{figure}
@@
 \subsubsubsection{Raspberry Pi 3 B+}
-This model of Raspberry Pi is able to run almost every GNU/Linux distribution ported to ARM architecture, Raspberry Pi OS (previously called Raspbian)\cite{RaspberryPiOs} is in fact a Debian port to ARM architecture.
+This model of Raspberry Pi is able to run almost every GNU/Linux distribution ported to ARM architecture, Raspberry Pi OS, formerly known as Raspbian,\cite{RaspberryPiOs} is in fact a Debian port to ARM architecture.
 After a little researching, it is possible to conclude that there is not other operative system which improves the performance of Raspberry Pi OS in a Raspberry Pi.
@@
 One of the most interesting features of Raspberry Pi is precisely that OS is installed and run from a microSD card, so the hardware is loading the operative system from this microSD card to the RAM directly, which improves notably the system load times, and increments its portability.
+
+Raspberry Pi foundation provides the \href{https://www.raspberrypi.org/downloads/noobs/}{\textit{Raspberry Pi Imager}} to perform the installation in a microSD card.\footnote{There are many ways to perform the Raspberry Pi OS installation in a microSD card, this document leaves it to readers to use their preferred method.
+However, this document also considers the \textit{Raspberry Pi Imager} method as the best one.} This tool allows the reader to choose between three different versions of Raspberry Pi OS.
+Recommended, Lite and Full version.
+Lite version is the same than Recommended version but without graphical user interface (GUI or Desktop Environment), meanwhile Full version is the same than Recommended version but with a few extra applications.
+
+This project is using the Recommended version for Raspberry Pi OS, nevertheless, the project does not require absolutely the desktop environment, so to maximize the available free space in the SD card, is better to install the Lite version instead of the Recommended version.
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=.7\textwidth]{../../pictures/rpi_imager.png}
+  \caption{Raspberry Pi Imager showing different options to install Raspberry Pi OS in a microSD card.}
+  \label{fig:RpiImager}
+\end{figure}
+
+Even so, independently of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-conf} application.\cite{RaspiConf}
 
@@
 
+\phantomsection
+\addcontentsline{toc}{section}{References}
 \printbibliography

### 2020-06-13 docs/paper/LaTex/Paper.tex
Fixes spaces between figures and text
Completing docs

@@
 % \setlength{\textfloatsep}{1\baselineskip}
-\setlength{\intextsep}{.5\baselineskip}
+\setlength{\intextsep}{1\baselineskip}
 
@@
 \usepackage{listings}
-\renewcommand{\lstlistingname}{Ejemplo}
+\renewcommand{\lstlistingname}{Example}
 
@@
 
-\newpage
-\section{Background}
-
+% \newpage
+% \section{Background}
 \newpage
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../../schemes/main_scheme_tbg.png}
+  \includegraphics[width=.73\textwidth]{../../schemes/main_scheme_tbg.png}
   \caption{General view of this solution.}
@@
 
-Even so, independently of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-conf} application.\cite{RaspiConf}
+Even so, independently of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-conf} application \cite{RaspiConf}.
+
+\doubt{WIFI (COMPLETE)}
+
+Another configuration that must be done is the Secure Shell (\cmd{ssh}) service; which according to Raspberry Pi documentation can be done from terminal using \cmd{systemd} \cite{RaspiSSH}.
+
+So generating a pair of keys and configuring properly the ssh access; it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions \cmd{700}, after this the reader should be able to connect through ssh service.
+
+\begin{lstlisting}[%
+    % float=thp,
+    language=json,
+    firstnumber=1,
+    caption={Creating a pair of ssh keys.},
+    label=ej:CreateEntity
+    ]
+    $ ssh-keygen
+
+    Generating public/private rsa key pair.
+    Enter file in which to save the key (/home/wyre/.ssh/id_rsa): 
+    Enter passphrase (empty for no passphrase): 
+    Enter same passphrase again: 
+    Your identification has been saved in /home/wyre/.ssh/id_rsa
+    Your public key has been saved in /home/wyre/.ssh/id_rsa.pub
+    The key fingerprint is:
+    SHA256:vkNRk/Fo8iGGo0pYlwb2L3vf3TgOfm11MmZW+BipXgQ wyre@DESKTOP-AFG84JJ
+    The key's randomart image is:
+    +---[RSA 3072]----+
+    |  o       .o     |
+    | . o . .  +oE    |
+    |  . = o +.+... o |
+    | o o o o.= .  = .|
+    |. . o . S..  o = |
+    | . . o ..   . O +|
+    |  . . ... .. * +.|
+    |     . ..+ o+oo  |
+    |        o.oo+o.  |
+    +----[SHA256]-----+
+\end{lstlisting}
+
+Once is available the remote control for Raspberry Pi through ssh, is possible to manage it completely from the command line, even upgrade the system and compile the required controller to get properly working the LoRa transceiver.
 
@@
 Dragino foresees this and provides a \href{https://github.com/dragino/dual_chan_pkt_fwd}{dual channel controller}.
-The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.
+The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.\footnote{This of course means Raspberry}
 

### 2020-06-14 docs/paper/LaTex/Paper.tex
Completing docs and testing vertical spaces in lstlisting environments

@@
 \renewcommand{\lstlistingname}{Example}
+% \lstset{%
+  % aboveskip=1\baselineskip plus 1\baselineskip minus 2\baselineskip,
+  % belowskip=1\baselineskip plus 1\baselineskip minus 2\baselineskip
+% }
+% \lstset{aboveskip=-18pt plus 2pt, belowskip=-18pt plus 2pt}
 
@@
 
-\lstdefinelanguage{json}{
+\lstdefinelanguage{json}{%
   % aboveskip=\baselineskip,
@@
 % (handy if you decide to change it further down the road)
-\lstdefinelanguage{yaml}{
+\lstdefinelanguage{yaml}{%
   keywords={true,false,null,y,n},
@@
 \setcounter{tocdepth}{4}
+
+% To include random paragraphs with \blindtext
+\usepackage{blindtext}
 
@@
 
-Even so, independently of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-conf} application \cite{RaspiConf}.
+Even so, independently of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-config} application \cite{RaspiConf}.
 
-\doubt{WIFI (COMPLETE)}
+One of the first and most important configurations is the internet access; this can be done via two different ways, using the ethernet port (which does not require any extra configuration, just tu plug the cable) or using the WiFi interface, which can be configured using also \cmd{raspi-config} \cite{RaspiWIFIcli}.
+Due to its versatility, there are multiple setup that could fulfil the requirements of this projects:
+
+\begin{itemize}
+  \item \textbf{PoE} (Power-over-Ethernet), this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} which is using the GPIO connector, so this prevents using it with the LoRa HAT.
+Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}.
+Besides, this method only works with up to 100m cables, so it would be necessary to have a power source and internet access point within a 100m radius.
+  \item Simple \textbf{ethernet} and mains electricity, this appears to be the best method among these shown
+\end{itemize}
 
@@
 
-So generating a pair of keys and configuring properly the ssh access; it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions \cmd{700}, after this the reader should be able to connect through ssh service.
+So generating a pair of keys and configuring properly the ssh access; it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
 
 \begin{lstlisting}[%
-    % float=thp,
+    % float=ht,
     language=json,

### 2020-06-14 docs/paper/LaTex/Paper.tex
Completing docs

@@
 This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
-So this low availability (of resources) along to their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
+So this low availability (of resources) along with their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
 
@@
 \begin{itemize}
-  \item \textbf{PoE} (Power-over-Ethernet), this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} which is using the GPIO connector, so this prevents using it with the LoRa HAT.
+  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} which is using the GPIO connector, so this makes impossible to use it along the LoRa HAT.
 Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}.
 Besides, this method only works with up to 100m cables, so it would be necessary to have a power source and internet access point within a 100m radius.
-  \item Simple \textbf{ethernet} and mains electricity, this appears to be the best method among these shown
+  \item \textbf{WiFi} and battery powered; this setup implies also the existence of a relatively close power source, because the access point used to provide that wireless network connection should work also with mains power.
+Probably the WiFi radio is not enough to make it the difference between this setup and the next one suggested.
+  \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one exposed, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
+  \item \textbf{GSM module} and battery or mains powered; this document consider this the ideal setup.
+This setup doesn't require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone.
+This would be the ideal way because it minimizes the resources needed, which is crucial in an environment where those probably won't be available ---in the middle of the forest.
+
+  Furthermore, this is apparently possible through two different options:
+    \begin{itemize}
+      \item \href{https://www.itead.cc/wiki/RPI_SIM800_GSM/GPRS_ADD-ON_V2.0}{Itead Raspberry Pi GSM Board (SIM800)}.
+This is not the most interesting way, because though GPIO pins comes through this does mean the HAT supports stacking, and even if that does, doesn't mean every other HAT would.
+      \item \href{https://tutorials-raspberrypi.com/raspberry-pi-gsm-module-mobile-internet-lte-3g-umts/}{USB GSM module}.
+This is apparently the most promising option.
+These kind of GSM modules are in general compliant with GNU/Linux systems and using a regular SIM card, they would be eventually able to provide an internet access point.
+Of course, this also will to increment the size of the whole device.
     \end{itemize}
+\end{itemize}
+
+It is important to note the necessity of a stable internet connection; \doubt{COMPLETE}
 
@@
     $ ssh-keygen
-
     Generating public/private rsa key pair.

### 2020-06-14 docs/paper/LaTex/Paper.tex
Completes technical info

@@
 
-It is important to note the necessity of a stable internet connection; \doubt{COMPLETE}
+It is important to note the necessity of a relatively stable internet connection; those gateways will forward the trafic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.
 
-Another configuration that must be done is the Secure Shell (\cmd{ssh}) service; which according to Raspberry Pi documentation can be done from terminal using \cmd{systemd} \cite{RaspiSSH}.
+Another configuration which must be done is the Secure Shell (\cmd{ssh}) service; which according to Raspberry Pi documentation can be done from terminal using \cmd{systemd} \cite{RaspiSSH}.
 
@@
 Once is available the remote control for Raspberry Pi through ssh, is possible to manage it completely from the command line, even upgrade the system and compile the required controller to get properly working the LoRa transceiver.
+At this point and depending on the type of access point used to provide internet service to the Raspberry, the reader will be able even to remote control the Raspberry over internet.
+For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh.}\footnote{In the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not.
+Anyway, as said the internet access point could be variable and it is not determined by this project.\\ The most important thing is to work confortably with the Raspberr Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a frist place.}
 
-\subsubsubsection{Dragino HAT}
+\subsubsubsection{Dragino GPS and LoRa HAT}
 Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role.
-These must be generic devices due LoRaWAN protocol versatility.
+These usually are generic devices due LoRaWAN protocol versatility.
 LoRaWAN is a cloud-based medium access control (MAC) layer protocol, but actually acts as a network layer protocol for managing communication between gateways and nodes, similar to a routing protocol.

### 2020-06-15 docs/paper/LaTex/Paper.tex
Completes technical description

@@
 Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}.
-Besides, this method only works with up to 100m cables, so it would be necessary to have a power source and internet access point within a 100m radius.
+Besides, this method only works with up to 100m cables, so it would be necessary to have a power source and internet access point within a 100m radius.\footnote{The \href{https://github.com/hallard/LoRasPI}{LoRasPI HAT} is more suitable for this option because does not cover whole GPIO connector, leaving free the GPIO pins used in \cite{RaspiPoEHack}.}
   \item \textbf{WiFi} and battery powered; this setup implies also the existence of a relatively close power source, because the access point used to provide that wireless network connection should work also with mains power.
@@
       \item \href{https://www.itead.cc/wiki/RPI_SIM800_GSM/GPRS_ADD-ON_V2.0}{Itead Raspberry Pi GSM Board (SIM800)}.
-This is not the most interesting way, because though GPIO pins comes through this does mean the HAT supports stacking, and even if that does, doesn't mean every other HAT would.
+This is not the most interesting way, because though GPIO pins comes through this does mean the HAT supports \href{http://www.pi-in-the-sky.com/index.php?id=stacking-guide}{stacking}, and even if that does, doesn't mean every other HAT would.
       \item \href{https://tutorials-raspberrypi.com/raspberry-pi-gsm-module-mobile-internet-lte-3g-umts/}{USB GSM module}.
@@
 
-It is important to note the necessity of a relatively stable internet connection; those gateways will forward the trafic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.
+It is important to note the necessity of a relatively stable internet connection; those gateways will forward the traffic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.
 
@@
 For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh.}\footnote{In the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not.
-Anyway, as said the internet access point could be variable and it is not determined by this project.\\ The most important thing is to work confortably with the Raspberr Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a frist place.}
+Anyway, as said the internet access point could be variable and it is not determined by this project.\\ The most important thing is to work comfortably with the Raspberry Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a first place.}
 
@@
 
-Dragino foresees this and provides a \href{https://github.com/dragino/dual_chan_pkt_fwd}{dual channel controller}.
-The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.\footnote{This of course means Raspberry}
+Dragino foresees this and provides a dual channel controller \cite{Dragino_DualChannelController_Rpi}.
+The most important thing about this dual channel controller, is not the possibility to use the transceiver in a dual channel mode, but to use it as a gateway in the physical sense.
+
+In order to get the HAT operative, the reader will need to enable the SPI (Serial Peripheral Interface). This can be done also with \cmd{raspi-config}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../../pictures/SPI_raspi-config.png}
+  \caption{SSH session to enable the SPI kernel module in Raspberry Pi using \cmd{raspi-config}. Inside \enquote{5 Interfacing options}}
+  \label{fig:EnablingSPI}
+\end{figure}
+
+Then it would be necessary to install the GPIO access library from Raspberry repositories, \cmd{sudo apt install wiringpi}.
+After this, the reader will be in position to clone the controller from the github repo \cite{Dragino_DualChannelController_Rpi}
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=json,
+  firstnumber=1,
+  caption={Instructions to compile and install as a system service},
+  label=ej:compiling_dual_channel_pkg
+  ]
+  cd ~
+  git clone https://github.com/dragino/dual_chan_pkt_fwd
+  cd dual_chan_pkt_fwd
+  make
+\end{lstlisting}
+
+Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=json,
+  firstnumber=1,
+  caption={defined pins in \cmd{global\_config.json}},
+  label=ej:global_configJSON
+  ]
+  "pin_nss": 6,
+  "pin_dio0": 7,
+  "pin_rst": 0
+\end{lstlisting}
+
+I order to obtain the gateway ID the reader will need to perform the following command to run for the first time the \cmd{dual\_chan\_pkt\_fwd} binary, i.e.
+once the compilation process has produced the \cmd{dual\_chan\_pkt\_fwd} is necessary to run it directly to check the terminal output and find for the gateway ID.
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=json,
+  firstnumber=1,
+  caption={Running for the first time the LoRa HAT controller.},
+  label=ej:Running_DualChann_Controller
+  ]
+  $ sudo ./dual_chan_pkt_fwd
+  server: .address = router.eu.staging.thethings.network; .port = 1700; .enable = 1
+  server: .address = router.eu.thethings.network; .port = 1700; .enable = 0
+  Gateway Configuration
+    your name (a@b.c)
+    Dual channel pkt forwarder
+    Latitude=0.00000000
+    Longitude=0.00000000
+    Altitude=10
+    Interface: eth0
+  Trying to detect module CE0 with NSS=6 DIO0=7 Reset=3 Led1=unused
+  SX1276 detected on CE0, starting.
+  Trying to detect module CE1 with NSS=6 DIO0=7 Reset=3 Led1=unused
+  SX1276 detected on CE1, starting.
+  Gateway ID: b8:27:eb:ff:ff:1b:14:9b
+  Listening at SF7 on 868.100000 Mhz.
+  Listening at SF7 on 868.100000 Mhz.
+  -----------------------------------
+  stat update: 2020-06-15 10:53:04 GMT no packet received yet 
+\end{lstlisting}
+
+After this last instruction (\cmd{sudo make install}) the
 

### 2020-06-15 docs/paper/LaTex/Paper.tex
Completes paper.tex

@@
 \fancyhf{} % Start with clearing everything in the header and footer
-\renewcommand{\headrulewidth}{0pt} % Removes the horizontal rule in header by default
+\renewcommand{\headrulewidth}{1pt} % Removes the horizontal rule in header by default
 \fancyfoot[R]{\thepage} % Set the right side of the footer to be the page number
+\fancyhead[R]{\rightmark}
+\setlength{\headheight}{14pt}% ...at least 51.60004pt
 
@@
 Once is available the remote control for Raspberry Pi through ssh, is possible to manage it completely from the command line, even upgrade the system and compile the required controller to get properly working the LoRa transceiver.
+
 At this point and depending on the type of access point used to provide internet service to the Raspberry, the reader will be able even to remote control the Raspberry over internet.
-For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh.}\footnote{In the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not.
-Anyway, as said the internet access point could be variable and it is not determined by this project.\\ The most important thing is to work comfortably with the Raspberry Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a first place.}
+For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh} ---in the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not.
+Anyway, as said the internet access point could be variable and it is not determined by this project.\footnote{The most important thing is to work comfortably with the Raspberry Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a first place.}
 

### 2020-06-15 docs/LaTex/Paper.tex
Sorting files again

--- 2020-06-15 docs/paper/LaTex/Paper.tex
@@
       \begin{mdframed}[innerbottommargin=490pt, innertopmargin=40pt, linewidth=1pt]
-        \includegraphics[width=\textwidth]{../../pictures/Seal/marca-uex-2-color.png}
+        \includegraphics[width=\textwidth]{../pictures/Seal/marca-uex-2-color.png}
       \end{mdframed}
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../../pictures/LoRa_Why_Range.png}
+  \includegraphics[width=.9\textwidth]{../pictures/LoRa_Why_Range.png}
   \caption{Bandwidth vs range plot for two of the most used telecommunication systems and LoRa.}
@@
   \centering
-  \includegraphics[width=.73\textwidth]{../../schemes/main_scheme_tbg.png}
+  \includegraphics[width=.73\textwidth]{../schemes/main_scheme_tbg.png}
   \caption{General view of this solution.}
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../../schemes/node_tbg.png}
+  \includegraphics[width=.9\textwidth]{../schemes/node_tbg.png}
   \caption{Diagram for a detailed view of a node.}
@@
   \centering
-    \includegraphics[width=.9\textwidth]{../../schemes/Signal_Conditioning.png}
+    \includegraphics[width=.9\textwidth]{../schemes/Signal_Conditioning.png}
   \caption{Signal conditioning stage.}
@@
   \centering
-  \includegraphics[width=.7\textwidth]{../../pictures/rpi_imager.png}
+  \includegraphics[width=.7\textwidth]{../pictures/rpi_imager.png}
   \caption{Raspberry Pi Imager showing different options to install Raspberry Pi OS in a microSD card.}
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../../pictures/SPI_raspi-config.png}
+  \includegraphics[width=.9\textwidth]{../pictures/SPI_raspi-config.png}
   \caption{SSH session to enable the SPI kernel module in Raspberry Pi using \cmd{raspi-config}. Inside \enquote{5 Interfacing options}}

### 2020-06-15 docs/LaTex/Paper.tex
Minor fixes

@@
 
-% Paper geometry
+% Document geometry
 \usepackage[a4paper,left=3cm,right=2.5cm,vmargin=2.5cm]{geometry}
@@
 % To produce a spanish automated output, for instance, TOC is called 'Índice' and
-% References is called 'Referencias'
+% References is called 'Referencias' (in this case I prefer english)
 \usepackage[english]{babel}
@@
 Then it would be necessary to install the GPIO access library from Raspberry repositories, \cmd{sudo apt install wiringpi}.
-After this, the reader will be in position to clone the controller from the github repo \cite{Dragino_DualChannelController_Rpi}
+After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi}
 
@@
 
-Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be
+Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be the following as stated by \cite{Dragino_DualChannelController_Rpi}
 
@@
 I order to obtain the gateway ID the reader will need to perform the following command to run for the first time the \cmd{dual\_chan\_pkt\_fwd} binary, i.e.
-once the compilation process has produced the \cmd{dual\_chan\_pkt\_fwd} is necessary to run it directly to check the terminal output and find for the gateway ID.
+once the compilation process has produced the \cmd{dual\_chan\_pkt\_fwd} binary is necessary to run it directly to check the terminal output and find for the gateway ID.
 

### 2020-06-15 docs/LaTex/Paper.tex
Minor fixes

@@
 At this point is easy to notice this solution will need to be a distributed solution, due high samples dispersion.
-\lang{As can be seen}, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts \doubt{between $157$-$170$ indviduals per hectare}\cite{ForestStandVol} (depending on the used model).
+As indicated, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts \doubt{between $157$-$170$ indviduals per hectare}\cite{ForestStandVol} (depending on the used model).
 So according to this and sample size determination theories, to get a great resolution could be necessary a big size for samples and the necessity of a big wireless network of distributed devices, since each device will correspond with an individual.
@@
 
-It is important to note that \lang{when talking about \enquote{low-power consumption}, in many cases it is actually meaning} battery-powered devices, for example.
+It is important to note that when talking about \enquote{low-power consumption}, in many cases it is actually meaning battery-powered devices, for example.
 

### 2020-06-16 docs/LaTex/Paper.tex
Completes docs and minor fixes

@@
 Over the years there have existed analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way.
-This could turn out bothering because the big size for this statistical population, \lang{¿as/like?
-it was exposed ¿before/above?}.
+This could turn out bothering because the big size for this statistical population, as previously stated.
 So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.
@@
 
-A lot of comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently \lang{¿like/as?} has been already said, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
+A lot of comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently as stated, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
 So this is intended to get a cheaper system and make it accessible to everyone who wants \lang{to monitorize} one or more trees growth.
@@
 
-\subsection{Hardware}
+\subsection{Physical Layer}
 
@@
 
-Due circumstances exposed in introduction section is needed an accessible and multipurpose platform \lang{to be the basis for the device design itself}, this is to say the core part of the dendrometer is an Arduino microcontroller.
+Due circumstances exposed in introduction section is needed an accessible and multipurpose platform to be the basis for the device design itself, this is to say the core part of the dendrometer is an Arduino microcontroller.
 
@@
 
-It is important to note that when talking about \enquote{low-power consumption}, in many cases it is actually meaning battery-powered devices, for example.
+It is important to note that when talking about \enquote{low-power consumption}, in many cases it is actually meaning battery-powered devices.
 
@@
 
-\subsection{Software}
+\subsection{Abstract layer}
 
@@
 
-After this last instruction (\cmd{sudo make install}) the
+So at the line 15 in the previous example \ref{ej:Running_DualChann_Controller}, the reader can see the Gateway ID, which wil be needed to connect this gateway to The Things Network stack.
+This ID is unique and it depends on the hardware.
+
+After to have obtained the gateway ID (\cmd{b8:27:eb:ff:ff:1b:14:9b} for this Dragino HAT), the reader can proceed then to install the system service by perform \cmd{sudo make install}.
+As the content of \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/Makefile#L22}{\cmd{Makefile}} shows at the line 22, \cmd{sudo make install} will copy the system service \cmd{dual\_chan\_pkt\_fwd.service} to the path \cmd{/lib/systemd/system/} and it will enable it in order to start the service ---the service will start the controller as the reader can see at line 9 in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{\cmd{dual\_chan\_pkt\_fwd.service}}--- at every system startup.\footnote{These kind of services are the way in which \cmd{systemd} manages the daemons running in background.}
+
+The reader must be aware of the path where is located the \cmd{dual\_chan\_pkt\_fwd} binary, this is so important because the service \cmd{dual\_chan\_pkt\_fwd.service} is using the absolute path to run the \cmd{dual\_chan\_pkt\_fwd} binary; So this could result in an error if the binary is not properly located or service is not properly modified.
+
+Once the service is installed and all these point have been checked, the reader will be able to manage the service to
+
+\begin{itemize}
+  \item start it with \cmd{sudo systemctl start dual\_chan\_pkt\_fwd.service}
+  \item stop it with \cmd{sudo systemctl stop dual\_chan\_pkt\_fwd.service}
+  \item check its status with \cmd{systemctl status -l dual\_chan\_pkt\_fwd.service}
+  \item check the journal with \cmd{journalctl -u dual\_chan\_pkt\_fwd.service}
+\end{itemize}
+
+These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status
 

### 2020-06-17 docs/LaTex/Paper.tex
Finises gateway subsection

@@
 
-So generating a pair of keys and configuring properly the ssh access; it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
+So generating a pair of keys and configuring properly the ssh access;\footnote{The reader will probably need to figure out the Raspberry IP in their local network, by other hand, inside a local network will not be necessary specify any port for ssh, though it usually is the 22 by default.} it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
 
@@
 
-Then it would be necessary to install the GPIO access library from Raspberry repositories, \cmd{sudo apt install wiringpi}.
+Then it would be necessary to install the GPIO access library from Raspberry repositories, so performing \cmd{sudo\, apt\, install\, wiringpi}.
 After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi}
@@
 
-These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status
+These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status; in the following picture it can be seen the service active (running)
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.82\textwidth]{../pictures/pkt_fwd-service.png}
+  \caption{systemd showing the \cmd{dual\_chan\_pkt\_fwd.service} status; it is active and running.}
+  \label{fig:dual_chan_pkt_fwd-status}
+\end{figure}
+
+As can be seen, this service was modified in order to reach properly the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even setup as suggested with \cmd{pi} user.}
+
+After this the Raspberry Pi is ready to forward LoRa packets to the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json#L33}{specified server at line 33 in \cmd{global\_conf.json}}.
+Besides, the reader will be able to manage the Raspberry Pi through ssh service, so the Raspberry only precises any kind of internet access point and electrical power supply.
 

### 2020-06-18 testing/Arduino/dendro/payload_decoder.js
Includes payload decoder and fixes for arduino code

@@ -1,0 +1,56 @@
+// TTN Decoder for TTN ABP Feather EU868 dendrometer
+// Link: 
+function Decoder(bytes, port) {
+    // Decode an uplink message from a buffer
+    // (array) of bytes to an object of fields.
+    var decoded = {};
+    
+    // Potmeter
+    rawPot = bytes[0] + bytes[1] * 256;
+    decoded.potVal = sflt162f(rawPot) * 1023;
+    
+    return decoded;
+  }
+  
+  function sflt162f(rawSflt16)
+      {
+      // rawSflt16 is the 2-byte number decoded from wherever;
+      // it's in range 0..0xFFFF
+      // bit 15 is the sign bit
+      // bits 14..11 are the exponent
+      // bits 10..0 are the the mantissa. Unlike IEEE format, 
+      // 	the msb is transmitted; this means that numbers
+      //	might not be normalized, but makes coding for
+      //	underflow easier.
+      // As with IEEE format, negative zero is possible, so
+      // we special-case that in hopes that JavaScript will
+      // also cooperate.
+      //
+      // The result is a number in the open interval (-1.0, 1.0);
+      // 
+      
+      // throw away high bits for repeatability.
+      rawSflt16 &= 0xFFFF;
+  
+      // special case minus zero:
+      if (rawSflt16 == 0x8000)
+          return -0.0;
+  
+      // extract the sign.
+      var sSign = ((rawSflt16 & 0x8000) !== 0) ? -1 : 1;
+      
+      // extract the exponent
+      var exp1 = (rawSflt16 >> 11) & 0xF;
+  
+      // extract the "mantissa" (the fractional part)
+      var mant1 = (rawSflt16 & 0x7FF) / 2048.0;
+  
+      // convert back to a floating point number. We hope 
+      // that Math.pow(2, k) is handled efficiently by
+      // the JS interpreter! If this is time critical code,
+      // you can replace by a suitable shift and divide.
+      var f_unscaled = sSign * mant1 * Math.pow(2, exp1 - 15);
+  
+      return f_unscaled;
+      }
+  

### 2020-06-18 testing/Arduino/dendro/payload_decoder.js
Minor fixes in decoder

@@ -1,4 +1,4 @@
-// TTN Decoder for TTN ABP Feather EU868 dendrometer
+// TTN Decoder for TTN ABP dendrometer EU868
 // Link: 
 function Decoder(bytes, port) {
     // Decode an uplink message from a buffer

### 2020-06-19 src/arduino/dendro/payload_decoder.js
Adds arduino src

@@ -1,0 +1,56 @@
+// TTN Decoder for TTN ABP dendrometer EU868
+// Link: 
+function Decoder(bytes, port) {
+    // Decode an uplink message from a buffer
+    // (array) of bytes to an object of fields.
+    var decoded = {};
+    
+    // Potmeter
+    rawPot = bytes[0] + bytes[1] * 256;
+    decoded.potVal = sflt162f(rawPot) * 1023;
+    
+    return decoded;
+  }
+  
+  function sflt162f(rawSflt16)
+      {
+      // rawSflt16 is the 2-byte number decoded from wherever;
+      // it's in range 0..0xFFFF
+      // bit 15 is the sign bit
+      // bits 14..11 are the exponent
+      // bits 10..0 are the the mantissa. Unlike IEEE format, 
+      // 	the msb is transmitted; this means that numbers
+      //	might not be normalized, but makes coding for
+      //	underflow easier.
+      // As with IEEE format, negative zero is possible, so
+      // we special-case that in hopes that JavaScript will
+      // also cooperate.
+      //
+      // The result is a number in the open interval (-1.0, 1.0);
+      // 
+      
+      // throw away high bits for repeatability.
+      rawSflt16 &= 0xFFFF;
+  
+      // special case minus zero:
+      if (rawSflt16 == 0x8000)
+          return -0.0;
+  
+      // extract the sign.
+      var sSign = ((rawSflt16 & 0x8000) !== 0) ? -1 : 1;
+      
+      // extract the exponent
+      var exp1 = (rawSflt16 >> 11) & 0xF;
+  
+      // extract the "mantissa" (the fractional part)
+      var mant1 = (rawSflt16 & 0x7FF) / 2048.0;
+  
+      // convert back to a floating point number. We hope 
+      // that Math.pow(2, k) is handled efficiently by
+      // the JS interpreter! If this is time critical code,
+      // you can replace by a suitable shift and divide.
+      var f_unscaled = sSign * mant1 * Math.pow(2, exp1 - 15);
+  
+      return f_unscaled;
+      }
+  

### 2020-06-19 docs/LaTex/Paper.tex
Completes docs and includes a new testing version for dendro.ino

@@
 \begin{itemize}
-  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack},\doubt{(I'm not sure about this citation (It's a forum))} so this means this hat is designed to be a node-class radio, not a gateway.}
+  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, \doubt{(I'm not sure about this cite (It's a forum))} so this means this hat is designed to be a node-class radio, not a gateway.}
   \item \textit{LoRa Shield for Arduino}\cite{DraginoArdShield}, which is plugged through analog and digital pins.
@@
 \end{figure}
+
+\subsubsubsection{Arduino sketch}
+The sketch to take advantage of LoRa Shield is mainly based on the LoRaMAC-in-C library (\cmd{lmic})\footnote{Dragino has a \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered as less accurate than the MCCI version, which is a fork itself from the first port of LMIC library by Matthijs Kooijman.} implemented by \href{https://mcci.com/}{MCCI} whose technical documentation can be found in \cite{MCCI_lmic}
+
+\doubt{}
 

### 2020-06-20 docs/LaTex/Paper.tex
Minor fixes
Minor changes

@@
 \subsubsubsection{Arduino sketch}
-The sketch to take advantage of LoRa Shield is mainly based on the LoRaMAC-in-C library (\cmd{lmic})\footnote{Dragino has a \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered as less accurate than the MCCI version, which is a fork itself from the first port of LMIC library by Matthijs Kooijman.} implemented by \href{https://mcci.com/}{MCCI} whose technical documentation can be found in \cite{MCCI_lmic}
-
-\doubt{}
+In order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has a \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered as less accurate than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}
 

### 2020-06-20 docs/LaTex/Paper.tex
Completes docs and minor fixes

--- 2020-06-19 docs/LaTex/Paper.tex
@@
 Like in the Introduction, the most proper way to present this technical description is to difference between software and hardware parts.
-However, is still interesting to present a simple and general diagram for whole designed solution:
+However, is still interesting to present a simple and general diagram for whole designed solution.
 
@@
 Most important features in those devices are: low-cost, low-powered, wireless communication and small size.
-The following diagram shows a diagram of its most relevant parts:
+The figure \ref{fig:NodeDiag} shows a diagram of its most relevant parts:
 
@@
 This stage is focused on improve the signal acquisition, according to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option.
-\doubt{Here is shown a possible schematics for this stage:}
+\doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage}
 
@@
 \subsubsubsection{Arduino sketch}
-The sketch to take advantage of LoRa Shield is mainly based on the LoRaMAC-in-C library (\cmd{lmic})\footnote{Dragino has a \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered as less accurate than the MCCI version, which is a fork itself from the first port of LMIC library by Matthijs Kooijman.} implemented by \href{https://mcci.com/}{MCCI} whose technical documentation can be found in \cite{MCCI_lmic}
+In a first place, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino}{the sketch} can be found in \cmd{src/dendro/dendro.ino} path.
 
-\doubt{}
+By other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has a \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
+The \cmd{arduino-lmic} library is, as its description in the README.md says, a port from the IBM LoRaWAN MAC in C, but \enquote{slightly modified to run in the Arduino environment allowing using the SX1272, SX1276 transceivers and compatible modules} \cite{MCCI_lmic}.
+
+There is an \textit{Arduino-like wrapper} library called \href{https://github.com/mcci-catena/arduino-lorawan}{\cmd{arduino-lorawan}} also developed by MCCI; this library provides apparently a higher abstraction level and it could be helpful in a future to keep growing this project, however, for now this project is implemented using the mentioned \cmd{arduino-lmic}.
+
+Basically the sketch is based on the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino}{ttn-abp example}, as the reader can see in the first comment block; that example sends a LoRaWAN valid package with the payload \enquote{Hello World!}.
 
@@
   \centering
-  \includegraphics[width=.7\textwidth]{../pictures/rpi_imager.png}
+  \includegraphics[width=.9\textwidth]{../pictures/rpi_imager.png}
   \caption{Raspberry Pi Imager showing different options to install Raspberry Pi OS in a microSD card.}
@@
 
-In order to get the HAT operative, the reader will need to enable the SPI (Serial Peripheral Interface). This can be done also with \cmd{raspi-config}
+In order to get the HAT operative, the reader will need to enable the SPI (Serial Peripheral Interface). This can be done also with \cmd{raspi-config} like shows the figure \ref{fig:EnablingSPI}
 
@@
 
-These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status; in the following picture it can be seen the service active (running)
+These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status; in the picture \ref{fig:dual_chan_pkt_fwd-status} it can be seen the service active (running)
 
@@
   \centering
-  \includegraphics[width=.82\textwidth]{../pictures/pkt_fwd-service.png}
+  \includegraphics[width=.9\textwidth]{../pictures/pkt_fwd-service.png}
   \caption{systemd showing the \cmd{dual\_chan\_pkt\_fwd.service} status; it is active and running.}

### 2020-06-20 docs/LaTex/Paper.tex
Completes docs

@@
 In a first place, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino}{the sketch} can be found in \cmd{src/dendro/dendro.ino} path.
+This is the sketch that must be uploaded to arduino microcontroller.
+MEGA 2560 in this case.
 
@@
 The \cmd{arduino-lmic} library is, as its description in the README.md says, a port from the IBM LoRaWAN MAC in C, but \enquote{slightly modified to run in the Arduino environment allowing using the SX1272, SX1276 transceivers and compatible modules} \cite{MCCI_lmic}.
+The original IBM LoRaWAN MAC in C library was developed for microcontrollers in general, their examples and HAL ---Hardware Abstraction Layer or rather Hardware Annotation Library--- were only for STM32 MCUs (but bare-metal, so no Arduino on top).
 
@@
 Basically the sketch is based on the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino}{ttn-abp example}, as the reader can see in the first comment block; that example sends a LoRaWAN valid package with the payload \enquote{Hello World!}.
+Besides it uses the ABP (Activation-by-personalisation) method ---in opposition to OTAA (Over-the-air-activation) method---, depending on the used method it is necessary to include different security keys provided by The Things Network stack in the code.
+
+As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point
 

### 2020-06-21 docs/LaTex/Paper.tex
Fixes bad pin definition in .ino sketch and completes docs

--- 2020-06-20 docs/LaTex/Paper.tex
@@
              {\ -\ }{{\mdseries\ -\ }}3,
+}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{dred}{rgb}{0.545,0,0}
+\definecolor{dblue}{rgb}{0,0,0.545}
+% \definecolor{lgrey}{rgb}{0.9,0.9,0.9}
+% \definecolor{gray}{rgb}{0.4,0.4,0.4}
+\definecolor{darkblue}{rgb}{0.0,0.0,0.6}
+\lstdefinelanguage{cpp}{
+      backgroundcolor=\color{background},  
+      basicstyle=\color{black}\mdseries\scriptsize\ttfamily,   
+      % breakatwhitespace=false,       
+      breaklines=true,               
+      % captionpos=b,                   
+      commentstyle=\color{dkgreen},   
+      deletekeywords={...},          
+      escapeinside={\%*}{*)},                  
+      frame=lines,                  
+      language=C++,                
+      keywordstyle=\color{purple},  
+      morekeywords={BRIEFDescriptorConfig,string,TiXmlNode,DetectorDescriptorConfigContainer,istringstream,cerr,exit}, 
+      identifierstyle=\color{black},
+      stringstyle=\color{blue},      
+      numbers=left,                 
+      numbersep=8pt,                  
+      numberstyle=\scriptsize, 
+      rulecolor=\color{black},        
+      showspaces=false,               
+      showstringspaces=false,        
+      showtabs=false,                
+      stepnumber=1,                   
+      tabsize=5,                     
+      title=\lstname,                 
     }
@@
 Basically the sketch is based on the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino}{ttn-abp example}, as the reader can see in the first comment block; that example sends a LoRaWAN valid package with the payload \enquote{Hello World!}.
+Besides it uses the ABP (Activation-by-personalization) method ---in opposition to OTAA (Over-the-air-activation) method---, depending on the used method it is necessary to include different security keys provided by The Things Network stack in the code.
+\cite{TTN_Security}
+
+As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
+This is to enable specific functions in the library to the Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must looks like in te example \ref{ej:lmic_project_config.h}
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=cpp,
+  firstnumber=1,
+  caption={Configuring lmic library radio.},
+  label=ej:lmic_project_config.h
+  ]
+  // project-specific definitions
+  #define CFG_eu868 1
+  //#define CFG_us915 1
+  //#define CFG_au915 1
+  //#define CFG_as923 1
+  // #define LMIC_COUNTRY_CODE LMIC_COUNTRY_CODE_JP	/* for as923-JP */
+  //#define CFG_kr920 1
+  //#define CFG_in866 1
+  #define CFG_sx1276_radio 1
+  //#define LMIC_USE_INTERRUPTS
+\end{lstlisting}
+
+After this, the sketch may use specific functions to perform radio transmissions in Europe region.
+This in fact, will has an impact in the \cmd{void setup()\{\}} function ---in the example sketch ttn-abp--- so lines between \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L232}{line 232} and line 254 will be executed when the library is configured for Europe region.
+
+Once the library has been properly configured, the second most important thing is set the pin mapping for the chip
 

### 2020-06-21 docs/LaTex/Paper.tex
Completes docs

@@
 As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
-This is to enable specific functions in the library to the Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must looks like in te example \ref{ej:lmic_project_config.h}
+This is to enable specific functions in the library for Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must looks like in te example \ref{ej:lmic_project_config.h}
 
@@
 
-Once the library has been properly configured, the second most important thing is set the pin mapping for the chip
+Once the library has been properly configured, the second most important thing is set the pin mapping for the chip, according to the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} in the \cmd{README.md}, \enquote{a variety of configurations are possible}.
+Then, the most appropriate way to make the pin mapping is following the Dragino indications, so following the figure \ref{fig:Ard_LoRa_Shield_pin_mapping} ---which can also be found in \href{http://wiki.dragino.com/index.php?title=Lora_Shield#Pin_Mapping_and_Unused_Pins}{Dragino documentation}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/LoRa_Shield_Pin_Mapping.png}
+  \caption{Arduino LoRa Shield pin mapping.}
+  \label{fig:Ard_LoRa_Shield_pin_mapping}
+\end{figure}
+
+The reader can see the LoRa Chip reset pin is \cmd{D9} and as DIO pins ---Digital Input/Output--- the board has assigned \cmd{D2} to DIO0, \cmd{D6} to DIO1, \cmd{D7} to DIO2.
+Besides, a brand new board will only have loaded the \cmd{R9} resistor, this means the LoRa CS pin will be the \cmd{D10}.
+This leads to set the pin mapping shown in example \ref{ej:pin_mapping}
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=cpp,
+  firstnumber=76,
+  caption={Pin mapping for this Dragino LoRa Shield.},
+  label=ej:pin_mapping
+  ]
+  // Pin mapping
+  const lmic_pinmap lmic_pins = {
+    .nss = 10,
+    .rxtx = LMIC_UNUSED_PIN,
+    .rst = 9,
+    .dio = {2, 6, 7},
+  };
+\end{lstlisting}
+
+By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as can be read also in the Pin Mapping section of the library \cmd{README.md}.
+As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.
 
@@
 \subsubsubsection{Raspberry Pi 3 B+}
-This model of Raspberry Pi is able to run almost every GNU/Linux distribution ported to ARM architecture, Raspberry Pi OS, formerly known as Raspbian,\cite{RaspberryPiOs} is in fact a Debian port to ARM architecture.
+This model of Raspberry Pi is able to run almost every GNU/Linux distribution ported to ARM architecture, Raspberry Pi OS, formerly known as Raspbian, \cite{RaspberryPiOs} is in fact a Debian port to ARM architecture.
 After a little researching, it is possible to conclude that there is not other operative system which improves the performance of Raspberry Pi OS in a Raspberry Pi.

### 2020-06-21 docs/LaTex/Paper.tex
Completes docs

@@
 
+Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L45}{45}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L49}{49} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L54}{54} must be obtained from TTN console \doubt{AS WILL BE EXPLAINED IN NEXT SECTIONS}
+
 \subsubsection{Gateway}

### 2020-06-22 docs/LaTex/Paper.tex
Completes docs and minor fixes

@@
 
-By other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has a \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
+By other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
 The \cmd{arduino-lmic} library is, as its description in the README.md says, a port from the IBM LoRaWAN MAC in C, but \enquote{slightly modified to run in the Arduino environment allowing using the SX1272, SX1276 transceivers and compatible modules} \cite{MCCI_lmic}.
@@
 Basically the sketch is based on the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino}{ttn-abp example}, as the reader can see in the first comment block; that example sends a LoRaWAN valid package with the payload \enquote{Hello World!}.
-Besides it uses the ABP (Activation-by-personalization) method ---in opposition to OTAA (Over-the-air-activation) method---, depending on the used method it is necessary to include different security keys provided by The Things Network stack in the code.
+Besides it uses the ABP (Activation-by-personalization) method ---in opposition to OTAA (Over-the-air-activation) method---, depending on the used method it is necessary to include different security keys provided by The Things Network stack (TTN stack) in the code.
 \cite{TTN_Security}
@@
 Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L45}{45}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L49}{49} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L54}{54} must be obtained from TTN console \doubt{AS WILL BE EXPLAINED IN NEXT SECTIONS}
+
+The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L69}{69}, as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=cpp,
+  firstnumber=64,
+  caption={Variables for the potentiometer signal acquisition.},
+  label=ej:potmeter_vars
+  ]
+  // payload to send to TTN gateway
+  static uint8_t payload[3];
+  static osjob_t sendjob;
+
+  // Potentiometer pins
+  int potPin = A2;               // Potmeter pin
+  double potVal = 0;             // Potmeter value
+\end{lstlisting}
+
+Finally, the reader can find the core of this sketch inside \cmd{void do\_send(osjob\_t* j)\{\}} function, at line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L175}{175}; here the reader can find the following lines of code
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=cpp,
+  firstnumber=180,
+  caption={Prpareing the payload.},
+  label=ej:prep_payload
+  ]
+  // Prepare upstream data transmission at the next possible time.        
+  // Read the analog value of the potmeter (0-1023)
+  potVal = analogRead(potPin);
+
+  // Write the value to the serial monitor
+  Serial.println(potVal);
+
+  // adjust for the f2sflt16 range (-1 to 1)
+  potVal = potVal/1023;
+
+  // Write AGAIN the value to the serial monitor
+  Serial.println(potVal);
+
+  // float -> int
+  // note: this uses the sflt16 datum (https://github.com/mcci-catena/arduino-lmic#sflt16)
+  uint16_t payloadPot = LMIC_f2sflt16(potVal);
+
+  Serial.println(payloadPot);
+
+  // int -> bytes
+  byte potLow = lowByte(payloadPot);
+  byte potHigh = highByte(payloadPot);
+
+  // place the bytes into the payload
+  payload[0] = potLow;
+  payload[1] = potHigh;
+
+  // prepare upstream data transmission at the next possible time.
+  // transmit on port 1 (the first parameter); you can use any value from 1 to 223 (others are reserved).
+  // don't request an ack (the last parameter, if not zero, requests an ack from the network).
+  // Remember, acks consume a lot of network resources; don't ask for an ack unless you really need it.
+  LMIC_setTxData2(1, payload, sizeof(payload)-1, 0);
+  Serial.println(F("Packet queued"));
+\end{lstlisting}
+
+So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
+The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.
+This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{pauloadPot} 16-bit integer is split into two bytes.
 

### 2020-06-22 docs/LaTex/Paper.tex
Completes docs

@@
 A lot of comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently as stated, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
-So this is intended to get a cheaper system and make it accessible to everyone who wants \lang{to monitorize} one or more trees growth.
+So this is intended to get a cheaper system and make it accessible to everyone who wants to monitor one or more trees growth.
 
@@
 So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
-The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.
-This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{pauloadPot} 16-bit integer is split into two bytes.
+The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.\footnote{A javascript decoder is also provided by the \href{https://github.com/mcci-catena/arduino-lmic\#uflt16-javascript-decoder}{\cmd{README.md}} but \doubt{it will be described in next sections.}}
+
+This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions.
+Finally these bytes are stored in the first two elements of initial \cmd{payload[3]} string declared at line 65, as the reader can see in example \ref{ej:potmeter_vars}, and the payload is prepared to send at the next possible time using the function \cmd{LMIC\_setTxData2()}.
+
+It is important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup() \{\}} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=cpp,
+  firstnumber=262,
+  caption={Channels setup.},
+  label=ej:chan_setup
+  ]
+  LMIC_setupChannel(0, 868100000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(1, 868300000, DR_RANGE_MAP(DR_SF12, DR_SF7B), BAND_CENTI);  // g-band
+  LMIC_setupChannel(2, 868500000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(3, 867100000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(4, 867300000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(5, 867500000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(6, 867700000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(7, 867900000, DR_RANGE_MAP(DR_SF12, DR_SF7),  BAND_CENTI);  // g-band
+  LMIC_setupChannel(8, 868800000, DR_RANGE_MAP(DR_FSK,  DR_FSK),  BAND_MILLI);  // g2-band
+\end{lstlisting}
+
+Because of explained above, the gateway will just receive one package per each $\approx$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening the channel 0, the microcontroller is transmitting at this channel once in 9---
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=cpp,
+  firstnumber=72,
+  caption={Transmission interval.},
+  label=ej:tx_interval
+  ]
+  // Schedule TX every this many seconds (might become longer due to duty
+  // cycle limitations).
+  const unsigned TX_INTERVAL = 60;
+\end{lstlisting}
+
+Of course this parameter can be adjusted, but for the final purpose of this project 60 seconds are even too many, in order to monitor changes in stem diameter of a tree, a more appropriate value would be 3600 seconds (1 hour) or even greater.
 
@@
 
-Here is the important point.
-This Dragino HAT for the Raspberry Pi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is not LoRaWAN compliant}.
-The main reason the SX1276 transceiver is not suitable to work as a gateway is that \doubt{it is actually a single channel transceiver}.
+\hypertarget{Raspi_HAT}{Here} is the important point.
+This Dragino HAT for the Raspberry Pi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is LoRaWAN compatible but isn't LoRaWAN compliant}.
+The main reason the SX1276 transceiver is not suitable to work as a gateway is that \textbf{it is actually a single channel transceiver}.
 Despite all of this, there still exist the possibility of use it as a gateway, because it is technically possible.
@@
 
+\newpage
 \phantomsection

### 2020-06-24 testing/lmic-decoder/payload_decoder.js
Testing CayenneLPP encoder

@@ -1,5 +1,4 @@
 // TTN Decoder for TTN ABP dendrometer EU868
-// Link: 
 function Decoder(bytes, port) {
     // Decode an uplink message from a buffer
     // (array) of bytes to an object of fields.
@@ -7,7 +6,9 @@ function Decoder(bytes, port) {
     
     // Potmeter
     rawPot = bytes[0] + bytes[1] * 256;
-    decoded.potVal = sflt162f(rawPot) * 1023;
+    //decoded.potVal = sflt162f(rawPot) * 1023;
+    //decoded.potVal = parseFloat((sflt162f(rawPot) * 1023).toPrecision(2));
+    decoded.potVal = parseFloat((sflt162f(rawPot) * 1023).toFixed(2));
     
     return decoded;
   }
@@ -53,4 +54,3 @@ function Decoder(bytes, port) {
   
       return f_unscaled;
       }
-  

### 2020-06-25 docs/LaTex/Paper.tex
Testing CayenneLPP encoding
Modifies docs, adds Postman environment and QuantumLeap docker files

@@
   language=cpp,
-  firstnumber=76,
+  firstnumber=77,
   caption={Pin mapping for this Dragino LoRa Shield.},
@@
   language=cpp,
-  firstnumber=64,
+  firstnumber=69,
   caption={Variables for the potentiometer signal acquisition.},
@@
   ]
-  // payload to send to TTN gateway
-  static uint8_t payload[3];
-  static osjob_t sendjob;
-
   // Potentiometer pins
@@
   language=cpp,
-  firstnumber=180,
+  firstnumber=181,
   caption={Prpareing the payload.},
@@
   // Prepare upstream data transmission at the next possible time.
+
   // Read the analog value of the potmeter (0-1023)
@@
   Serial.println(potVal);
-
-  // adjust for the f2sflt16 range (-1 to 1)
-  potVal = potVal/1023;
-
-  // Write AGAIN the value to the serial monitor
-  Serial.println(potVal);
-
-  // float -> int
-  // note: this uses the sflt16 datum (https://github.com/mcci-catena/arduino-lmic#sflt16)
-  uint16_t payloadPot = LMIC_f2sflt16(potVal);
-
-  Serial.println(payloadPot);
-
-  // int -> bytes
-  byte potLow = lowByte(payloadPot);
-  byte potHigh = highByte(payloadPot);
-
-  // place the bytes into the payload
-  payload[0] = potLow;
-  payload[1] = potHigh;
 
@@
   // Remember, acks consume a lot of network resources; don't ask for an ack unless you really need it.
-  LMIC_setTxData2(1, payload, sizeof(payload)-1, 0);
+
+  lpp.reset();
+  lpp.addAnalogInput(1, potVal);
+
+  LMIC_setTxData2(1, lpp.getBuffer(), lpp.getSize(), 0);
   Serial.println(F("Packet queued"));
@@
 
-So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
+So, as can be seen, in order to prepare the payload firstly the analog pin is stored in \cmd{potVal} variable, after this the function \cmd{lpp.reset()} is ued to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the  \doubt{REDO}
+%So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
 The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.\footnote{A javascript decoder is also provided by the \href{https://github.com/mcci-catena/arduino-lmic\#uflt16-javascript-decoder}{\cmd{README.md}} but \doubt{it will be described in next sections.}}
@@
   language=cpp,
-  firstnumber=262,
+  firstnumber=248,
   caption={Channels setup.},
@@
   language=cpp,
-  firstnumber=72,
+  firstnumber=73,
   caption={Transmission interval.},

### 2020-06-25 docs/LaTex/Paper.tex
Sorting files and folders

@@
 
+  // Reset CayenneLPP buffer
+  lpp.reset();
+
+  // Add the measured value to CayenneLPP buffer
+  lpp.addAnalogInput(1, potVal);
+
   // prepare upstream data transmission at the next possible time.
@@
   // Remember, acks consume a lot of network resources; don't ask for an ack unless you really need it.
-
-  lpp.reset();
-  lpp.addAnalogInput(1, potVal);
-
   LMIC_setTxData2(1, lpp.getBuffer(), lpp.getSize(), 0);
@@
 
-So, as can be seen, in order to prepare the payload firstly the analog pin is stored in \cmd{potVal} variable, after this the function \cmd{lpp.reset()} is ued to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the  \doubt{REDO}
+So, as can be seen, in order to prepare the payload firstly the analog pin is stored in \cmd{potVal} variable, after this the function \cmd{lpp.reset()} is ued to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time.
 %So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
@@
   language=cpp,
-  firstnumber=248,
+  firstnumber=250,
   caption={Channels setup.},

### 2020-06-25 docs/LaTex/Paper.tex
Minor fixes

@@
 
-Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L45}{45}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L49}{49} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L54}{54} must be obtained from TTN console \doubt{AS WILL BE EXPLAINED IN NEXT SECTIONS}
-
-The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L69}{69}, as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
+Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console \doubt{AS WILL BE EXPLAINED IN NEXT SECTIONS}.
+The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
 
@@
 
-Finally, the reader can find the core of this sketch inside \cmd{void do\_send(osjob\_t* j)\{\}} function, at line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L175}{175}; here the reader can find the following lines of code
+Finally, the reader can find the core of this sketch inside \cmd{void do\_send(osjob\_t* j)\{\}} function, at line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L176}{176}; here the reader can find the following lines of code
 
@@
 
-This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions.
+These \cmd{lpp.*()} functions are part of the CayenneLPP library, which is included at line 
+%This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions.
 Finally these bytes are stored in the first two elements of initial \cmd{payload[3]} string declared at line 65, as the reader can see in example \ref{ej:potmeter_vars}, and the payload is prepared to send at the next possible time using the function \cmd{LMIC\_setTxData2()}.

### 2020-06-25 src/ttn-stack/payload_decoder.js
Removes old javascript decoder
It isn't needed anymore because the project
has switched to CayenneLPP to FIWARE compatibility

@@ -1,56 +1,0 @@
-// TTN Decoder for TTN ABP dendrometer EU868
-// Link: 
-function Decoder(bytes, port) {
-    // Decode an uplink message from a buffer
-    // (array) of bytes to an object of fields.
-    var decoded = {};
-    
-    // Potmeter
-    rawPot = bytes[0] + bytes[1] * 256;
-    decoded.potVal = sflt162f(rawPot) * 1023;
-    
-    return decoded;
-  }
-  
-  function sflt162f(rawSflt16)
-      {
-      // rawSflt16 is the 2-byte number decoded from wherever;
-      // it's in range 0..0xFFFF
-      // bit 15 is the sign bit
-      // bits 14..11 are the exponent
-      // bits 10..0 are the the mantissa. Unlike IEEE format, 
-      // 	the msb is transmitted; this means that numbers
-      //	might not be normalized, but makes coding for
-      //	underflow easier.
-      // As with IEEE format, negative zero is possible, so
-      // we special-case that in hopes that JavaScript will
-      // also cooperate.
-      //
-      // The result is a number in the open interval (-1.0, 1.0);
-      // 
-      
-      // throw away high bits for repeatability.
-      rawSflt16 &= 0xFFFF;
-  
-      // special case minus zero:
-      if (rawSflt16 == 0x8000)
-          return -0.0;
-  
-      // extract the sign.
-      var sSign = ((rawSflt16 & 0x8000) !== 0) ? -1 : 1;
-      
-      // extract the exponent
-      var exp1 = (rawSflt16 >> 11) & 0xF;
-  
-      // extract the "mantissa" (the fractional part)
-      var mant1 = (rawSflt16 & 0x7FF) / 2048.0;
-  
-      // convert back to a floating point number. We hope 
-      // that Math.pow(2, k) is handled efficiently by
-      // the JS interpreter! If this is time critical code,
-      // you can replace by a suitable shift and divide.
-      var f_unscaled = sSign * mant1 * Math.pow(2, exp1 - 15);
-  
-      return f_unscaled;
-      }
-  

### 2020-06-25 docs/LaTex/Paper.tex
Completes docs

@@
 
-As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
+As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L30}{lines 30 and 31} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
 This is to enable specific functions in the library for Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must looks like in te example \ref{ej:lmic_project_config.h}

### 2020-06-26 docs/LaTex/Paper.tex
Completes docs

@@
 However, is still interesting to present a simple and general diagram for whole designed solution.
+So as the reader can see in the figure \ref{fig:GenView} is shown the general diagram.
+
+Firstly the TTN stack send through MQTT protocol the payload to FIWARE IoTagent-LoRaWAN generic enabler.
+Secondly, FIWARE generic enablers process that payload and store the contained information (also called \textit{context data}) along with a timestamp in the persisting data base (CrateDB), after this, the context data can be visualized in a web view provided by Grafena.
+
+At this moment this solution requires to send manually the HTTP requests in order to communicate with FIWARE generic enablers manually.
+The ideal situation would be to develop a desktop or web application capable of provide any kind of graphical user interface to easily setup the two most important requirements to get the solution working as intended ---Create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
 
@@
   \centering
-  \includegraphics[width=.73\textwidth]{../schemes/main_scheme_tbg.png}
+  \includegraphics[width=\textwidth]{../schemes/main_scheme_tbg.png}
   \caption{General view of this solution.}
@@
 This stage is focused on improve the signal acquisition, according to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option.
-\doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage}
+\doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage.}
+
+It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead a variable resistor, this is because taking advantage of the Arduino power source, particularly the 5 \si{\volt} and GND pins.
 
@@
 
-By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as can be read also in the Pin Mapping section of the library \cmd{README.md}.
+By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
 As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.
@@
   potVal = analogRead(potPin);
+  potVal = 100.0 * potVal / 1023;
 
@@
 
-So, as can be seen, in order to prepare the payload firstly the analog pin is stored in \cmd{potVal} variable, after this the function \cmd{lpp.reset()} is ued to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time.
+So, as can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time.
 %So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
@@
 
-These \cmd{lpp.*()} functions are part of the CayenneLPP library, which is included at line 
+These \cmd{lpp.*()} functions are part of the CayenneLPP library, in particular, the \cmd{lpp.addAnalogInput()} function encodes the raw value as a number encoded in 2 bytes (16 bits signed) with a unit of 0.01, in 16 bits signed is possible to represent numbers from -32768 until 32767, so with a unit of 0.01, that amounts to a range from -327.68 to +327.67; that is because the sketch includes the line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L185}{185}, where the raw value \cmd{potVal} is scaled to a [0, 100.0] range.
+This range will fit in the range that can be encoded in CayenneLPP.
 %This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions.
@@
 
-It is important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup() \{\}} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
+CayenneLPP is so importante because FIWARE will work with this data model, so any other third party encoding will not be admitted/decoded by the FIWARE LoRaWAN agent.
+
+It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup() \{\}} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
 
@@
   language=cpp,
-  firstnumber=250,
+  firstnumber=251,
   caption={Channels setup.},
@@
 
-Then it would be necessary to install the GPIO access library from Raspberry repositories, so performing \cmd{sudo\, apt\, install\, wiringpi}.
-After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi}
+Then it would be necessary to install the GPIO access library from Raspberry repositories, so performing \cmd{sudo\, apt\, install\, wiringpi} the package manager should install \cmd{wiringpi} providing all necessary libraries.
+
+After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi} and compile it performing the command shown in example \ref{ej:compiling_dual_channel_pkg} ---Like will be explained below, it is also important to note where \cmd{git} will clone the repository and where the reader will compile it, because the location of the \cmd{dual\_chan\_pkt\_fwd} binary is so critical for that systemd service (also included in the GitHub repository) \cmd{dual\_chan\_pkt\_fwd.service} works properly.
 
@@
 
+The core of the software part is deployed using Docker containers \cite{Docker_container}. These docker containers are deploying essentially the following components
+
+\begin{itemize}
+  \item \textbf{IoTagent-LoRaWAN}, which is in charge of forward the traffic from TTN stack to Orion Context Broker.
+  \item \textbf{Orion Context Broker}, this is the \textbf{core generic enabler}, in fact as indicated, this is the only one required component to the solution be considered as \enquote{powered by FIWARE}.
+Besides this generic enabler requires a \textbf{Mongo database}.
+  \item \textbf{QuantumLeap}, in order to make the context data persisting, this generic enabler is needed.
+Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the information will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persisting.
+\end{itemize}
+
 \newpage

### 2020-06-26 docs/LaTex/Paper.tex
Completes docs

@@
 
+% Improves underline creating the command \myuline{}
+% https://alexwlchan.net/2017/10/latex-underlines/
+\usepackage{contour}
+\usepackage{ulem}
+\renewcommand{\ULdepth}{1.8pt}
+\contourlength{0.8pt}
+\newcommand{\myuline}[1]{%
+  \uline{\phantom{#1}}%
+  \llap{\contour{white}{#1}}%
+}
+
 % To add an extra level in sections
@@
 
-So, as can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time.
+So, as can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
 %So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
@@
 
-These \cmd{lpp.*()} functions are part of the CayenneLPP library, in particular, the \cmd{lpp.addAnalogInput()} function encodes the raw value as a number encoded in 2 bytes (16 bits signed) with a unit of 0.01, in 16 bits signed is possible to represent numbers from -32768 until 32767, so with a unit of 0.01, that amounts to a range from -327.68 to +327.67; that is because the sketch includes the line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L185}{185}, where the raw value \cmd{potVal} is scaled to a [0, 100.0] range.
+These \cmd{lpp.*()} functions are part of the CayenneLPP library \cite{CayenneLPP_lib} ---which is also dependent of ArduinoJson library \cite{ArduinoJson_lib}; in particular, the \cmd{lpp.addAnalogInput()} function encodes the raw value as a number encoded in 2 bytes (16 bits signed) with a unit of 0.01.
+In 16 bits signed is possible to represent numbers from -32768 until 32767, so with a unit of 0.01, that amounts to a range from -327.68 to +327.67; that is because the sketch includes the line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L185}{185}, where the raw value \cmd{potVal} is scaled to a [0, 100.0] range.
 This range will fit in the range that can be encoded in CayenneLPP.
@@
 
-CayenneLPP is so importante because FIWARE will work with this data model, so any other third party encoding will not be admitted/decoded by the FIWARE LoRaWAN agent.
-
-It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup() \{\}} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
+CayenneLPP is so important because FIWARE will work with this data model, so any other third party encoding will not be admitted/decoded by the FIWARE LoRaWAN agent.
+
+It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup()} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
 
@@
   \item \textbf{IoTagent-LoRaWAN}, which is in charge of forward the traffic from TTN stack to Orion Context Broker.
-  \item \textbf{Orion Context Broker}, this is the \textbf{core generic enabler}, in fact as indicated, this is the only one required component to the solution be considered as \enquote{powered by FIWARE}.
+  \item \textbf{Orion Context Broker}, this is the \myuline{core generic enabler}, in fact as indicated, this is the only one required component to the solution can be considered as \enquote{powered by FIWARE}.
 Besides this generic enabler requires a \textbf{Mongo database}.
-  \item \textbf{QuantumLeap}, in order to make the context data persisting, this generic enabler is needed.
-Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the information will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persisting.
+  \item \textbf{QuantumLeap}, in order to make the context data persistent, this generic enabler is needed.
+Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the information will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persistent.
 \end{itemize}

### 2020-06-26 docs/LaTex/Paper.tex
Minor fixes
Completes docs

@@
 % https://alexwlchan.net/2017/10/latex-underlines/
+% `normalem` option is to prevent underlines for book,
+% articles, etc. titles in references.
 \usepackage{contour}
-\usepackage{ulem}
+\usepackage[normalem]{ulem}
 \renewcommand{\ULdepth}{1.8pt}
@@
 
-Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console \doubt{AS WILL BE EXPLAINED IN NEXT SECTIONS}.
+Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{pleas, refer to that subsection for further details}.
 The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
@@
   \item \textbf{Orion Context Broker}, this is the \myuline{core generic enabler}, in fact as indicated, this is the only one required component to the solution can be considered as \enquote{powered by FIWARE}.
-Besides this generic enabler requires a \textbf{Mongo database}.
+Besides this generic enabler requires a \textbf{Mongo database} to store entities and subscriptions among other things.
   \item \textbf{QuantumLeap}, in order to make the context data persistent, this generic enabler is needed.
-Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the information will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persistent.
+Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the context data will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persistent.
+Also a \textbf{Redis database} is used for geocoding purposes, however this project will not deal with it ---this is because of this database is not even being included in the figure \ref{fig:GenView}.
+  \item \textbf{Grafana}, this container is intended to provide a web interface to visualize and manage the retrieved context data.
+It stores its configuration and customization values in an embebed sqlite3 database; so there is not necessity for a separate container.
 \end{itemize}
+
+These bold highlighted are the deployed containers that will be detailed later, many of them are not even part of FIWARE platform, though.
+First this document will expose how to setup the gateway, create an application and add devices inside The Things Network stack.
+
+\subsubsection{The Things Network Stack}\label{sssec:TTN_Stack}
+This is a platform for the LoRaWAN protocol where is possible to manage gateways, devices and applications.
+It provides a centralized way to manage these kind of things regard in security too, generating several keys to tight the end-devices to the chosen server cloud solution (FIWARE in this case).
+There are three fundamental process the reader must follow to include nodes and gateways in the platform, those are luckily pretty well documented.
+
+\begin{itemize}
+  \item Register a gateway; this is the process where \cmd{Gateway ID} shown in example \ref{ej:Running_DualChann_Controller} is needed; this process is, as indicated, well documented in \cite{Gateway_reg}.
+  \item Add an application; this application will provide a set of security keys as well as a zone where manage different devices (these are basically the end-nodes i.e.
+the arduino microcontrollers).
+Is also documented in \cite{Add_application}.
+  \item Register a device within added application; these devices will be ultimately the microcontrollers. Again the reader can find the relative documents in \cite{Device_reg}. 
+\end{itemize}
+
+\newpage
+\section{Conclusions}
 

### 2020-06-27 docs/LaTex/Paper.tex
Completes docs

@@
 
-Because of explained above, the gateway will just receive one package per each $\approx$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening the channel 0, the microcontroller is transmitting at this channel once in 9---
+Because of explained above, the gateway will just receive one package per each $\sim$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening the channel 0, the microcontroller is transmitting at this channel once in 9---
 
@@
 
+Once these steps are completed, the TTN web interface will provide an overview with all data regarding to the device like figure \ref{fig:dev_overview} shows.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/TTN_dev_overview.png}
+  \caption{A device overview.}
+  \label{fig:dev_overview}
+\end{figure}
+
+Another important consideration inside TTN platform is to check which decoder is chosen for the receives payloads, this can be check in
+
 \newpage

### 2020-06-27 docs/LaTex/Paper.tex
Completes docs

@@
 % I need to use that \texorpdfstring{}{} for the left-side ToC in PDF viewers, because the ® cannot be rendered in there.
-\subsubsection{\texorpdfstring{LoRa\textsuperscript{\textregistered} (Lo{\normalfont\sffamily ng} Ra{\normalfont\sffamily nge})}{LoRa® (Long Range)}}
+\subsubsection{\texorpdfstring{LoRa\textsuperscript{\textregistered} (Lo{\normalfont\sffamily ng} Ra{\normalfont\sffamily nge})}{LoRa (Long Range)}}
 LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite[]{LoRaGeneral}.
@@
 \begin{itemize}
-  \item Register a gateway; this is the process where \cmd{Gateway ID} shown in example \ref{ej:Running_DualChann_Controller} is needed; this process is, as indicated, well documented in \cite{Gateway_reg}.
-  \item Add an application; this application will provide a set of security keys as well as a zone where manage different devices (these are basically the end-nodes i.e.
+  \item \textbf{Register a gateway}; this is the process where \cmd{Gateway ID} shown in example \ref{ej:Running_DualChann_Controller} is needed; this process is, as indicated, well documented in \cite{Gateway_reg}.
+  \item \textbf{Add an application}; this application will provide a set of security keys as well as a zone where manage different devices (these are basically the end-nodes i.e.
 the arduino microcontrollers).
 Is also documented in \cite{Add_application}.
-  \item Register a device within added application; these devices will be ultimately the microcontrollers. Again the reader can find the relative documents in \cite{Device_reg}. 
+  \item \textbf{Register a device} within added application; these devices will be ultimately the microcontrollers. Again the reader can find the relative documents in \cite{Device_reg}. 
 \end{itemize}
@@
 
-Another important consideration inside TTN platform is to check which decoder is chosen for the receives payloads, this can be check in
+Another important consideration inside TTN platform is to check which decoder is chosen for the received payloads, this can be checked by clicking on \cmd{Payload Formats} tab inside Application Overview.
+As can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so importan to choose the \cmd{CayenneLPP} decoder, because as indicated, the payload is being encoded in the end-nodes using the CayenneLPP standard.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/TTN_Payload_Formats.png}
+  \caption{Choosing CayenneLPP as payload decoder}
+  \label{fig:CayenneLPP_decoder}
+\end{figure}
+
+Once this has been done, the microcontrollers has also been programmed and the gateway configured, the reader should be able to read the sensor data in the \cmd{Data} tab like figure \ref{fig:sens_data_TTN}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/TTN_sensors_data.png}
+  \caption{Following sensors data in TTN.}
+  \label{fig:sens_data_TTN}
+\end{figure}
+
+Then it should be possible to retrieve this data using the FIWARE IoTagent-LoRaWAN.
+The setup of this different FIWARE generic enablers will be explained in the next subsection, but essentially this IoTagent will use, as indicated previously, all those application and devices identifiers in order to manage the incoming context data from different microcontrollers (also called devices or end-nodes).
+
+\subsubsection{FIWARE}
+As shown in figure \ref{fig:GenView}, this solution is being implemented using particularly three FIWARE generic enablers, the most important one, of course, \textbf{Orion Context Broker}, also a LoRaWAN agent called \textbf{IoTagent-LoRaWAN} and finally to make data persistent, \textbf{QuantumLeap}.
+
+The workflow is like picture \ref{fig:GenView} shows, first the IoTagent receives the context data from TTN, this IoTagent sends this context data to the Orion Context Broker and this latter notifies QuantumLeap the variations on the context data, thus QuantumLeap can proceed to store them into Crate database.
+
+However this process does not happen out of the box, i.e.
+it is necessary to properly setup these generic enablers to get this workflow.
+This can be done following the FIWARE API documentation.
+These configuration can be donde using HTTP requests due FIWARE essentially provides a RESTful API \cite{FIWARE_RESTful}, these HTTP requests can be performed using multiple clients like \cmd{curl} or \cmd{Postman} among others.
+This project in fact also provides a \href{https://github.com/WyRe/lora-arduino-dendrometer/tree/master/src/postman}{\cmd{Postman} collection and environment} that can be imported to manage them using a confortable graphical interface.
+
+Anyway and after of this little introduction, the most important thing is to deploy all FIWARE generic enablers and its dependencies in a secure and stable way.
+This is because this project uses \cmd{docker} and \cmd{docker-compose} in order to ensure these modules deployment and execution.
+Again this project provides the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/fiware/docker-compose.yml}{\cmd{docker-compose.yml}} file to ensure a correct deployment.
+
+\begin{itemize}
+  \item In a first place the reader will need to create a new device for the \textbf{IoTagent-LoRaWAN}, this
+\end{itemize}
 

### 2020-06-27 docs/LaTex/Paper.tex
Completes docs and adds a first version for a python GUI

@@
 Another important consideration inside TTN platform is to check which decoder is chosen for the received payloads, this can be checked by clicking on \cmd{Payload Formats} tab inside Application Overview.
-As can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so importan to choose the \cmd{CayenneLPP} decoder, because as indicated, the payload is being encoded in the end-nodes using the CayenneLPP standard.
+As can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so importan to choose the \cmd{CayenneLPP} decoder, because as indicated, \myuline{the payload is being encoded in the end-nodes using the CayenneLPP standard}.
 
@@
 
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/docker_containers_deployed.png}  
+  \caption{\cmd{tmux} showing docker containers status and logs.}
+  \label{fig:docker_containers}
+\end{figure}
+
+So cloning the project and navigating to \cmd{src/fiware/} folder in the terminal the reader will must perform \cmd{docker-compose up -d} to deploy all containers and keep detached from the logs output.
+Once this has been done, performing \cmd{docker ps -a} will show the status of deployed containers, as figure \ref{fig:docker_containers} shows.
+In a similar way, to check the logs the reader could perform \cmd{docker-compose logs -f <NAME>} where the argument \cmd{<NAME>} is optional and when is not provided \cmd{docker-compose} will show logs for all containers; this is also shown in picture \ref{fig:docker_containers}.
+
+When these containers are properly deployed, it should be possible to send the HTTP requests, however, these must follow a precise order;
+
 \begin{itemize}
-  \item In a first place the reader will need to create a new device for the \textbf{IoTagent-LoRaWAN}, this
+  \item In a first place the reader will need to create a new device for the \textbf{IoTagent-LoRaWAN}, this can be done performing a \cmd{POST} HTTP request whose body will include the necessary data to retrieve the TTN stack context data.
+
+  \begin{lstlisting}[%
+    % float=ht,
+    language=json,
+    firstnumber=1,
+    caption={Device provisioning for IoTagent-LoRaWAN},
+    label=ej:provisioning_devices
+    ]
+    curl localhost:4061/iot/devices -s -S -H 'Content-Type: application/json' -H 'Accept: application/json' -H 'fiware-service: atosioe' -H 'fiware-servicepath: /lorattn' -d @- <<EOF
+    {
+      "devices": [
+        {
+          "device_id": "node_2",
+          "entity_name": "LORA-N-2",
+          "entity_type": "LoraDevice",
+          "timezone": "Europe/Madrid",
+          "attributes": [
+            {
+              "object_id": "analog_in_1",
+              "name": "analog_in_1",
+              "type": "Number"
+            }
+          ],
+          "internal_attributes": {
+            "lorawan": {
+              "application_server": {
+                "host": "eu.thethings.network",
+                "username": "dendrometer",
+                "password": "ttn-account-v2.173lH8wwDiIRC8E2JgM9ScXyuNRlPpMefpazS0TIhnU",
+                "provider": "TTN"
+              },
+              "dev_eui": "00EC51D264A6F8FD",
+              "app_eui": "70B3D57ED0030B5D",
+              "application_id": "dendrometer",
+              "application_key": "3736B21218C4443F639FAF4114B723A1"
+            }
+          }
+        }
+      ]
+    }
+    EOF
+
+    EOF
+  \end{lstlisting}
 \end{itemize}

### 2020-06-27 src/api-gui/api-gui.py

@@ -1,0 +1,222 @@
+import requests
+import tkinter as tk
+import json
+
+
+def use_api(url, method=requests.get, header=None, body=None):
+    body = json.dumps(body) if body else None
+
+    respuesta = method(url, headers=header, data=body)
+    print(url, method, respuesta.status_code, respuesta.text)
+
+    return respuesta
+
+
+# meter en env
+fiware_service = 'atosioe'
+fiware_servicepath = '/lorattn'
+TTN_app_id = 'dendrometer'
+TTN_app_pw = 'ttn-account-v2.173lH8wwDiIRC8E2JgM9ScXyuNRlPpMefpazS0TIhnU'
+TTN_dev_eui = '00EC51D264A6F8FD'
+TTN_app_eui = '70B3D57ED0030B5D'
+TTN_app_skey = '7AF6B9A29EAF893F66C0F4360BA8DD5B'
+
+
+header_iot_device = {
+    "Content-Type": "application/json",
+    "Accept": "application/json",
+    "fiware-service": f"{fiware_service}",
+    "fiware-servicepath": f"{fiware_servicepath}"
+}
+
+header_orion = {
+        "fiware-service": f"{fiware_service}",
+        "fiware-servicepath": f"{fiware_servicepath}"
+}
+
+
+class Application(tk.Frame):
+    entities = list()
+    default_entity = 'LORA-N-0'
+
+    devices = list()
+    default_device = 'node_0'
+
+    subscriptions = list()
+
+    def __init__(self):
+        self.gui = tk.Tk()
+        self.gui.title('Fiware Client')
+        super().__init__(self.gui)
+        self.gui.geometry("350x400")
+        self.pack()
+        main_frame = tk.Frame(self.gui)
+        main_frame.pack(fill=tk.BOTH, expand=tk.YES)
+
+        self.label_orion_ver = tk.Label(main_frame, text='Orion version: ' + use_api('http://localhost:1026/version')
+                                        .json()['orion']['version'])
+        self.label_iot_ver = tk.Label(main_frame, text='IoT version: ' + use_api('http://localhost:4061/iot/about')
+                                      .json()['libVersion'])
+        self.label_quantumleap_ver = tk.Label(main_frame,
+                                              text='QuantumLeap version: ' +
+                                                    use_api('http://localhost:8668/v2/version').json()['version'])
+
+        self.label_orion_ver.place(x=10, y=10)
+        self.label_iot_ver.place(x=10, y=30)
[…]

### 2020-06-27 docs/LaTex/Paper.tex
Completes docs and Postman collection

@@
         {
-          "device_id": "node_2",
-          "entity_name": "LORA-N-2",
+          "device_id": "node_0",
+          "entity_name": "LORA-N-0",
           "entity_type": "LoraDevice",
@@
     EOF
-
-    EOF
   \end{lstlisting}

### 2020-06-27 docs/LaTex/Paper.tex
Completes docs and improves Postman Collection

@@
 \begin{itemize}
-  \item In a first place the reader will need to create a new device for the \textbf{IoTagent-LoRaWAN}, this can be done performing a \cmd{POST} HTTP request whose body will include the necessary data to retrieve the TTN stack context data.
+  \item In a first place the reader will need to provision a new device for the \textbf{IoTagent-LoRaWAN}, this can be done performing a \cmd{POST} HTTP request whose body will include the necessary data to retrieve the TTN stack context data.
 
@@
     ]
-    curl localhost:4061/iot/devices -s -S -H 'Content-Type: application/json' -H 'Accept: application/json' -H 'fiware-service: atosioe' -H 'fiware-servicepath: /lorattn' -d @- <<EOF
+    $ curl localhost:4061/iot/devices -s -S -H 'Content-Type: application/json' -H 'Accept: application/json' -H 'fiware-service: atosioe' -H 'fiware-servicepath: /lorattn' -d @- <<EOF
     {
@@
               "application_id": "dendrometer",
-              "application_key": "3736B21218C4443F639FAF4114B723A1"
+              "application_key": "3736B21218C4443F639FAF4114B723A1",
+              "data_model": "cayennelpp"
             }
@@
   \end{lstlisting}
+
+  As example \ref{ej:provisioning_devices} shows, is mandatory to provide a \cmd{"device\_id"}, which would be desirable to match with de TTN device ID, an \cmd{"entity\_name"} in order to make Orion Context Broker create that entity and also an \cmd{"username"} and \cmd{"application\_id"} that must match with the application ID in TTN.
+Besides It will be needed the \cmd{"password"}, which can be obtained from the Application Overview in TTN website and the \cmd{"dev\_eui"}, \cmd{"app\_eui"} and \cmd{"application\_key"}, all of these ---relative to the created device--- are available into TTN dashboard.
+This \cmd{POST} request is included in the \cmd{Postman} collection provided by this project repository; of course \myuline{the reader will need to replace the keys in the environment for their own}.
+
+  When a new device is created, \textbf{IoTagent-LoRaWAN} causes that \textbf{Orion Context Broker} creates a new entity. In fact this can be checked by performing a \cmd{GET} HTTP request,
+
+  \begin{lstlisting}[%
+    % float=ht,
+    language=json,
+    firstnumber=1,
+    caption={Querying OrionCB for entities list.},
+    label=ej:OrionCB_entities_Get
+    ]
+    $ curl localhost:1026/v2/entities \                                                                       
+       -H 'fiware-service: atosioe' \
+       -H 'fiware-servicepath: /lorattn' | python -mjson.tool
+      % Total    % Received % Xferd  Average Speed   Time    Time     Time  Current
+                                     Dload  Upload   Total   Spent    Left  Speed
+    100   248  100   248    0     0   4275      0 --:--:-- --:--:-- --:--:--  4275
+    [
+        {
+            "id": "LORA-N-0",
+            "type": "LoraDevice",
+            "TimeInstant": {
+                "type": "DateTime",
+                "value": "2020-06-27T15:58:48.00Z",
+                "metadata": {}
+            },
+            "analog_in_1": {
+                "type": "Number",
+                "value": 55.03,
+                "metadata": {
+                    "TimeInstant": {
+                        "type": "DateTime",
+                        "value": "2020-06-27T15:58:48.00Z"
+                    }
+                }
+            }
+        }
+    ]
+  \end{lstlisting}
+
+  Example \ref{ej:OrionCB_entities_Get} shows the response of that \cmd{GET} request when a new entity is created by Orion Context Broker.
+
+  \item Secondly, after have created the device and therefore the entity, the next thing to do is to create a subscription in order to Orion Context Broker notifies somewhere ---QuantumLeap in this case.
+In a similar way, to create this subscription the reader can perform a \cmd{POST} request against Orion Context Broker, so as example \ref{ej:OrionCB_subscription} shows,
+
+  \begin{lstlisting}[%
+    % float=ht,
+    language=json,
+    firstnumber=1,
+    caption={Creating a subscription to notify QuantumLeap.},
+    label=ej:OrionCB_subscription
+    ]
+    $ curl localhost:1026/v2/subscriptions -s -S -H 'Content-Type: application/json' -H 'fiware-service: atosioe' -H 'fiware-servicepath: /lorattn' -d @- <<EOF
+    {
+      "description": "A subscription to get info about LORA-N-0",
+      "subject": {
+        "entities": [
+          {
+            "id": "LORA-N-0",
+            "type": "LoraDevice"
+          }
+        ],
+        "condition": {
+          "attrs": [
+            "analog_in_1"
+          ]
+        }
+      },
+      "notification": {
+        "http": {
+          "url": "http://quantumleap:8668/v2/notify"
+        },
+        "attrs": [
+          "analog_in_1"
+        ],
+        "metadata": ["dateCreated", "dateModified"]
+      },
+      "throttling": 5
+    }
+    EOF
+  \end{lstlisting}
+
+  As the reader can see, for this subscription has been setup the \cmd{"url": "http://quantumleap:8668/v2/notify"} i.e.
+QuantumLeap will be notified at any change in the entity with \cmd{"id": "LORA-N-0"}; also it's a good practice to check that all went as expected, so sending a \cmd{GET} request the response should be like example \ref{ej:OrionCB_subscriptions_get}
+
+  \begin{lstlisting}[%
+    % float=ht,
+    language=json,
+    firstnumber=1,
+    caption={Querying for all existent subscriptions in OrionCB.},
+    label=ej:OrionCB_subscriptions_get
+    ]
+    $ curl localhost:1026/v2/subscriptions \
+    -H 'fiware-service: atosioe' \
+    -H 'fiware-servicepath: /lorattn' | python -mjson.tool
+      % Total    % Received % Xferd  Average Speed   Time    Time     Time  Current
+                                     Dload  Upload   Total   Spent    Left  Speed
+    100   535  100   535    0     0   9067      0 --:--:-- --:--:-- --:--:--  9067
+    [
+        {
+            "id": "5ef772be8153f85ef485b5fd",
+            "description": "A subscription to get info about LORA-N-0",
+            "status": "active",
+            "subject": {
+                "entities": [
+                    {
+                        "id": "LORA-N-0",
+                        "type": "LoraDevice"
+                    }
+                ],
+                "condition": {
+                    "attrs": [
+                        "analog_in_1"
+                    ]
+                }
+            },
+            "notification": {
+                "timesSent": 17,
+                "lastNotification": "2020-06-27T16:40:56.00Z",
+                "attrs": [
+                    "analog_in_1"
+                ],
+                "onlyChangedAttrs": false,
+                "attrsFormat": "normalized",
+                "http": {
+                    "url": "http://quantumleap:8668/v2/notify"
+                },
+                "metadata": [
+                    "dateCreated",
+                    "dateModified"
+                ],
+                "lastSuccess": "2020-06-27T16:40:56.00Z",
+                "lastSuccessCode": 200
+            },
+            "throttling": 5
+        }
+    ]
+
+  \end{lstlisting}
+
 \end{itemize}
+
+Once these two steps have been performed successfully, the context data should be routed to the Crate database where will be persistent.
 

### 2020-06-27 src/api-gui/api-gui.py
Renames some folders and sorts folders and files

@@ -6,21 +6,24 @@ import json
 def use_api(url, method=requests.get, header=None, body=None):
     body = json.dumps(body) if body else None
 
-    respuesta = method(url, headers=header, data=body)
-    print(url, method, respuesta.status_code, respuesta.text)
+    response = method(url, headers=header, data=body)
+    print(url, method, response.status_code, response.text)
 
-    return respuesta
+    return response
 
 
-# meter en env
+# These are customizable patterns, actually
 fiware_service = 'atosioe'
 fiware_servicepath = '/lorattn'
+
+# These are application unique
 TTN_app_id = 'dendrometer'
 TTN_app_pw = 'ttn-account-v2.173lH8wwDiIRC8E2JgM9ScXyuNRlPpMefpazS0TIhnU'
-TTN_dev_eui = '00EC51D264A6F8FD'
 TTN_app_eui = '70B3D57ED0030B5D'
-TTN_app_skey = '7AF6B9A29EAF893F66C0F4360BA8DD5B'
 
+# These are device unique
+TTN_app_skey = '7AF6B9A29EAF893F66C0F4360BA8DD5B'
+TTN_dev_eui = '00EC51D264A6F8FD'
 
 header_iot_device = {
     "Content-Type": "application/json",
@@ -30,8 +33,8 @@ header_iot_device = {
 }
 
 header_orion = {
-        "fiware-service": f"{fiware_service}",
-        "fiware-servicepath": f"{fiware_servicepath}"
+    "fiware-service": f"{fiware_service}",
+    "fiware-servicepath": f"{fiware_servicepath}"
 }
 
 
@@ -48,7 +51,7 @@ class Application(tk.Frame):
         self.gui = tk.Tk()
         self.gui.title('Fiware Client')
         super().__init__(self.gui)
-        self.gui.geometry("350x400")
+        self.gui.geometry("640x480")
         self.pack()
         main_frame = tk.Frame(self.gui)
         main_frame.pack(fill=tk.BOTH, expand=tk.YES)
@@ -59,7 +62,7 @@ class Application(tk.Frame):
                                       .json()['libVersion'])
         self.label_quantumleap_ver = tk.Label(main_frame,
                                               text='QuantumLeap version: ' +
-                                                    use_api('http://localhost:8668/v2/version').json()['version'])
+                                                   use_api('http://localhost:8668/v2/version').json()['version'])
 
         self.label_orion_ver.place(x=10, y=10)
         self.label_iot_ver.place(x=10, y=30)
[…]

### 2020-06-27 docs/LaTex/Paper.tex
Improves docs and docker-compose
A separate volume for grafena is being created from now
to make the grafana configs persistent
Completes docs

@@
 At this moment this solution requires to send manually the HTTP requests in order to communicate with FIWARE generic enablers manually.
-The ideal situation would be to develop a desktop or web application capable of provide any kind of graphical user interface to easily setup the two most important requirements to get the solution working as intended ---Create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
+The ideal situation would be to develop a desktop or web application capable of provide any kind of graphical user interface to easily set up the two most important requirements to get the solution working as intended ---Create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
 
@@
 
-It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup()} function the channels are setup as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
+It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup()} function the channels are set up as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
 
@@
   \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one exposed, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
-  \item \textbf{GSM module} and battery or mains powered; this document consider this the ideal setup.
+  \item \textbf{GSM module} and battery or mains powered; this document consider this as the ideal setup.
 This setup doesn't require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone.
@@
 
-As can be seen, this service was modified in order to reach properly the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even setup as suggested with \cmd{pi} user.}
+As can be seen, this service was modified in order to reach properly the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even set up as suggested with \cmd{pi} user.}
 
@@
 These bold highlighted are the deployed containers that will be detailed later, many of them are not even part of FIWARE platform, though.
-First this document will expose how to setup the gateway, create an application and add devices inside The Things Network stack.
+First this document will expose how to set up the gateway, create an application and add devices inside The Things Network stack.
 
@@
 However this process does not happen out of the box, i.e.
-it is necessary to properly setup these generic enablers to get this workflow.
+it is necessary to properly set up these generic enablers to get this workflow.
 This can be done following the FIWARE API documentation.
@@
 
-  As the reader can see, for this subscription has been setup the \cmd{"url": "http://quantumleap:8668/v2/notify"} i.e.
+  As the reader can see, for this subscription has been set up the \cmd{"url": "http://quantumleap:8668/v2/notify"} i.e.
 QuantumLeap will be notified at any change in the entity with \cmd{"id": "LORA-N-0"}; also it's a good practice to check that all went as expected, so sending a \cmd{GET} request the response should be like example \ref{ej:OrionCB_subscriptions_get}
@@
 
+\subsubsection{Grafana}
+Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; due it's actually capable to acquire the data from multiple sources.
+In this project will be used in order to monitor the context data along the time.
+
+This web application can be accessed via \cmd{http://localhost:3000} in the web browser ---the file \cmd{docker-compose.yml} is already setup to expose that port and allow the user to connect Grafana in that way.
+The default credentials for the first access are \cmd{username: admin} and \cmd{password: admin}.
+Once the reader has accessed for a first time is possible to modify those credentials or even create new users with different privileges.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/Grafana_home.png}
+  \caption{Grafana home view.}
+  \label{fig:grafana_home}
+\end{figure}
+
+AS the central panel in the home view indicates, there are two fundamental task that must be performed to get a detailed view of our context data; these are to add a data source and set up a new dashboard.
+
+To add a new data source the reader can use the left sidebar or click where says \enquote{Add your first dada source} in the home view.
+CrateDB is PostgreSQL compatible, so it must be chosen a PostgreSQL database and fill the form with the data shown in figure \ref{fig:Grafana_add_datasource}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/Grafana_add_datasource.png}
+  \caption{Adding CrateDB as the data source}
+  \label{fig:Grafana_add_datasource}
+\end{figure}
+
+\enquote{Host} must be \cmd{crate:5432} because is the tcp port that grafana uses to queries, this can be seen also in figure \ref{fig:docker_containers} another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/Grafana_panel_setup.png}
+  \caption{Setting up a panel.}
+  \label{fig:panel_setup}
+\end{figure}
+
+To get the panel drawing the CrateDB stored context data, the reader must specify a query to retrieve the data from the database.
+This can be done in the \enquote{Query} tab below the graph shown in picture \ref{fig:panel_setup}, here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should looks as follows in example \ref{ej:CrateDB_query}
+
+\begin{lstlisting}[%
+  % float=ht,
+  language=json,
+  firstnumber=1,
+  caption={Query to retrieve the stored context data.},
+  label=ej:CrateDB_query
+  ]
+    SELECT
+    time_index AS "time",
+    entity_id,
+    analog_in_1
+  FROM mtatosioe.etloradevice
+  ORDER BY 1
+\end{lstlisting}
+
+The table name \cmd{mtatosioe.etloradevice} can be found inside CrateDB admin panel, which can be accessed in \cmd{http://localhost:4200/\#!/tables} from the web browser.
+So after clicking on \enquote{Apply} button on top right corner the reader should have something similar to figure \ref{fig:final_panel}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/Grafana_final_panel.png}
+  \caption{Dendrometers panel}
+  \label{fig:final_panel}
+\end{figure}
+
 \newpage

### 2020-06-27 docs/LaTex/Paper.tex
Testing python ideas

@@
 
+Grafana has a wide set of options intended to monitor different kind of systems, it can represent data in different type of graphics and even allows export to csv format, this can be useful for a more deep analysis using other technologies like, for instance, R, a well-known language widely used for statistical analysis.
+
+Of course, this context data could be acquired directly from CrateDB by means of some custom application developed specifically to this task, however this is beyond of the purpose of this project.
+Maybe in a future a feature like this could be implemented.
+
 \newpage

### 2020-06-27 testing/python/api-gui_v2/api-gui.py

@@ -1,0 +1,225 @@
+import requests
+import tkinter as tk
+import json
+
+
+def use_api(url, method=requests.get, header=None, body=None):
+    body = json.dumps(body) if body else None
+
+    response = method(url, headers=header, data=body)
+    print(url, method, response.status_code, response.text)
+
+    return response
+
+
+# These are customizable patterns, actually
+fiware_service = 'atosioe'
+fiware_servicepath = '/lorattn'
+
+# These are application unique
+TTN_app_id = 'dendrometer'
+TTN_app_pw = 'ttn-account-v2.173lH8wwDiIRC8E2JgM9ScXyuNRlPpMefpazS0TIhnU'
+TTN_app_eui = '70B3D57ED0030B5D'
+
+# These are device unique
+TTN_app_skey = '7AF6B9A29EAF893F66C0F4360BA8DD5B'
+TTN_dev_eui = '00EC51D264A6F8FD'
+
+header_iot_device = {
+    "Content-Type": "application/json",
+    "Accept": "application/json",
+    "fiware-service": f"{fiware_service}",
+    "fiware-servicepath": f"{fiware_servicepath}"
+}
+
+header_orion = {
+    "fiware-service": f"{fiware_service}",
+    "fiware-servicepath": f"{fiware_servicepath}"
+}
+
+
+class Application(tk.Frame):
+    entities = list()
+    default_entity = 'LORA-N-0'
+
+    devices = list()
+    default_device = 'node_0'
+
+    subscriptions = list()
+
+    def __init__(self):
+        self.gui = tk.Tk()
+        self.gui.title('Fiware Client')
+        super().__init__(self.gui)
+        self.gui.geometry("640x480")
+        self.pack()
+        main_frame = tk.Frame(self.gui)
+        main_frame.pack(fill=tk.BOTH, expand=tk.YES)
+
+        self.label_orion_ver = tk.Label(main_frame, text='Orion version: ' + use_api('http://localhost:1026/version')
+                                        .json()['orion']['version'])
+        self.label_iot_ver = tk.Label(main_frame, text='IoT version: ' + use_api('http://localhost:4061/iot/about')
+                                      .json()['libVersion'])
+        self.label_quantumleap_ver = tk.Label(main_frame,
+                                              text='QuantumLeap version: ' +
+                                                   use_api('http://localhost:8668/v2/version').json()['version'])
+
+        self.label_orion_ver.place(x=10, y=10)
+        self.label_iot_ver.place(x=10, y=30)
[…]

### 2020-06-27 testing/python/tkinter_tests/tabs/tabs.py

@@ -1,0 +1,18 @@
+import tkinter as tk
+from tkinter import ttk
+
+root = tk.Tk()
+root.title("Tab Widget")
+tabControl = ttk.Notebook(root)
+
+tab1 = ttk.Frame(tabControl)
+tab2 = ttk.Frame(tabControl)
+
+tabControl.add(tab1, text='Tab 1')
+tabControl.add(tab2, text='Tab 2')
+tabControl.pack(expand=1, fill="both")
+
+ttk.Label(tab1, text="Welcome to GeeksForGeeks").grid(column=0, row=0, padx=30, pady=30)
+ttk.Label(tab2, text="Lets dive into the world of computers").grid(column=0, row=0, padx=30, pady=30)
+
+root.mainloop()

### 2020-06-28 testing/python/api-gui_v2/api-gui.py
Improves api-guy.py

@@ -1,29 +1,13 @@
 import requests
 import tkinter as tk
+from tkinter import ttk
+from string import hexdigits
+
 import json
-
-
-def use_api(url, method=requests.get, header=None, body=None):
-    body = json.dumps(body) if body else None
-
-    response = method(url, headers=header, data=body)
-    print(url, method, response.status_code, response.text)
-
-    return response
-
 
 # These are customizable patterns, actually
 fiware_service = 'atosioe'
 fiware_servicepath = '/lorattn'
-
-# These are application unique
-TTN_app_id = 'dendrometer'
-TTN_app_pw = 'ttn-account-v2.173lH8wwDiIRC8E2JgM9ScXyuNRlPpMefpazS0TIhnU'
-TTN_app_eui = '70B3D57ED0030B5D'
-
-# These are device unique
-TTN_app_skey = '7AF6B9A29EAF893F66C0F4360BA8DD5B'
-TTN_dev_eui = '00EC51D264A6F8FD'
 
 header_iot_device = {
     "Content-Type": "application/json",
@@ -38,6 +22,19 @@ header_orion = {
 }
 
 
+def use_api(url, method=requests.get, header=None, body=None):
+    body = json.dumps(body) if body else None
+
+    response = method(url, headers=header, data=body)
+    # print(url, method, response.status_code, response.text)
+
+    return response
+
+
+def is_hex(string) -> bool:
+    return all([char in hexdigits for char in string])
+
+
 class Application(tk.Frame):
     entities = list()
     default_entity = 'LORA-N-0'
@@ -47,56 +44,169 @@ class Application(tk.Frame):
 
     subscriptions = list()
 
+    # These are application unique
+    TTN_app_id = None
+    TTN_app_pw = None
+    TTN_app_eui = None
+
+    # These are device unique
+    TTN_app_skey = None
+    TTN_dev_eui = None
+
     def __init__(self):
+        # TKINTER constructor
         self.gui = tk.Tk()
         self.gui.title('Fiware Client')
         super().__init__(self.gui)
-        self.gui.geometry("640x480")
+        self.gui.minsize(640, 480)
+        # self.gui.geometry("640x480")
         self.pack()
-        main_frame = tk.Frame(self.gui)
-        main_frame.pack(fill=tk.BOTH, expand=tk.YES)
 
-        self.label_orion_ver = tk.Label(main_frame, text='Orion version: ' + use_api('http://localhost:1026/version')
[…]

### 2020-06-28 testing/python/api-gui_v2/api-gui.py
Improves api-gui.py

@@ -78,7 +78,7 @@ class Application(tk.Frame):
         self.set_about_widgets()
         self.set_applications_keys_widgets()
 
-        #
+        # Initial get request
         self.api_get_devices_and_entities()
 
     def set_about_widgets(self) -> None:
@@ -106,8 +106,8 @@ class Application(tk.Frame):
 
         # Inputs
         self.entry_ttn_app_id = tk.Entry(self.tab_app_keys)
-        self.entry_ttn_app_pw = tk.Entry(self.tab_app_keys)
-        self.entry_ttn_app_eui = tk.Entry(self.tab_app_keys)
+        self.entry_ttn_app_pw = tk.Entry(self.tab_app_keys, width=60)
+        self.entry_ttn_app_eui = tk.Entry(self.tab_app_keys, width=17)
         self.entry_ttn_app_id.place(x=10, y=30)
         self.entry_ttn_app_pw.place(x=10, y=70)
         self.entry_ttn_app_eui.place(x=10, y=110)
@@ -315,20 +315,21 @@ class Application(tk.Frame):
         self.api_get_devices_and_entities()
 
     @staticmethod
-    def get_iot_body() -> dict:
-        if len(Application.devices) == 0:
-            device = Application.default_device
-        else:
-            last_device = Application.devices[-1]
-            last_id = int(last_device.split('_')[-1])
-            device = last_device.split('_')[0] + '_' + str(last_id + 1)
+    def generate_next_device_name():
 
-        if len(Application.entities) == 0:
-            entity = Application.default_entity
-        else:
-            last_entity = Application.entities[-1]
-            last_id = int(last_entity.split('-')[-1])
-            entity = '-'.join(last_entity.split('-')[:-1]) + '-' + str(last_id + 1)
+        if not Application.devices:
+            return Application.default_device, Application.default_entity
+        ids = [int(device.split('_')[-1]) for device in Application.devices]
+        for i in range(0, max(ids) + 1):
+            if i not in ids:
+                return f'node_{i}', f'LORA-N-{i}'
+
+        next_id = max(ids) + 1
+        return f'node_{next_id}', f'LORA-N-{next_id}'
+
+    @staticmethod
+    def get_iot_body():
+        device, entity = Application.generate_next_device_name()
 
         return {
             "devices": [

### 2020-06-28 testing/python/api-gui_v0.2/api-gui.py
Sorting api-gui versions

@@ -2,7 +2,6 @@ import requests
 import tkinter as tk
 from tkinter import ttk
 from string import hexdigits
-
 import json
 
 # These are customizable patterns, actually
@@ -62,7 +61,7 @@ class Application(tk.Frame):
         # self.gui.geometry("640x480")
         self.pack()
 
-        # TABS
+        # Tabs
         self.notebook = ttk.Notebook(self.gui)  # Tab controller
         self.tab_app_keys = ttk.Frame(self.notebook)
         self.tab_devices_management = ttk.Frame(self.notebook)
@@ -70,7 +69,7 @@ class Application(tk.Frame):
         self.notebook.add(self.tab_app_keys, text='Applications Keys')
         self.notebook.add(self.tab_devices_management, text='Devices Management')
         self.notebook.add(self.tab_about, text='About')
-        self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
+        # self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
         self.notebook.pack(expand=1, fill="both")
 
         # Widgets
@@ -80,6 +79,34 @@ class Application(tk.Frame):
 
         # Initial get request
         self.api_get_devices_and_entities()
+
+        # Labels
+        self.label_ttn_app_id = tk.Label()
+        self.label_ttn_app_pw = tk.Label()
+        self.label_ttn_app_eui = tk.Label()
+        self.label_app_skey = tk.Label()
+        self.label_dev_eui = tk.Label()
+        self.label_title_devices = tk.Label()
+        self.label_title_entities = tk.Label()
+        self.label_title_subscriptions = tk.Label()
+        self.label_devices = tk.Label()
+        self.label_entities = tk.Label()
+        self.label_subscriptions = tk.Label()
+
+        # Inputs
+        self.entry_ttn_app_id = tk.Entry()
+        self.entry_ttn_app_pw = tk.Entry()
+        self.entry_ttn_app_eui = tk.Entry()
+        self.entry_app_skey = tk.Entry()
+        self.entry_dev_eui = tk.Entry()
+
+        # Buttons
+        self.set_app_keys = tk.Button()
+        self.button_create_device = tk.Button()
+        self.button_delete_device = tk.Button()
+
+        # Listbox
+        self.listbox_devices = tk.Listbox()
 
     def set_about_widgets(self) -> None:
         self.label_orion_ver = tk.Label(self.tab_about,

### 2020-06-28 testing/python/api-gui_v0.3/api-gui.py

@@ -1,0 +1,394 @@
+import requests
+import tkinter as tk
+from tkinter import ttk
+from string import hexdigits
+import json
+
+# These are customizable patterns, actually
+fiware_service = 'atosioe'
+fiware_servicepath = '/lorattn'
+
+header_iot_device = {
+    "Content-Type": "application/json",
+    "Accept": "application/json",
+    "fiware-service": f"{fiware_service}",
+    "fiware-servicepath": f"{fiware_servicepath}"
+}
+
+header_orion = {
+    "fiware-service": f"{fiware_service}",
+    "fiware-servicepath": f"{fiware_servicepath}"
+}
+
+
+def use_api(url, method=requests.get, header=None, body=None):
+    body = json.dumps(body) if body else None
+
+    response = method(url, headers=header, data=body)
+    # print(url, method, response.status_code, response.text)
+
+    return response
+
+
+def is_hex(string) -> bool:
+    return all([char in hexdigits for char in string])
+
+
+class Application(tk.Frame):
+    entities = list()
+    default_entity = 'LORA-N-0'
+
+    devices = list()
+    default_device = 'node_0'
+
+    subscriptions = list()
+
+    # These are application unique
+    TTN_app_id = None
+    TTN_app_pw = None
+    TTN_app_eui = None
+
+    # These are device unique
+    TTN_app_skey = None
+    TTN_dev_eui = None
+
+    def __init__(self):
+        # TKINTER constructor
+        self.gui = tk.Tk()
+        self.gui.title('Fiware Client')
+        super().__init__(self.gui)
+        self.gui.minsize(640, 480)
+        # self.gui.geometry("640x480")
+        self.pack()
+
+        # Tabs
+        self.notebook = ttk.Notebook(self.gui)  # Tab controller
+        self.tab_app_keys = ttk.Frame(self.notebook)
+        self.tab_devices_management = ttk.Frame(self.notebook)
+        self.tab_about = ttk.Frame(self.notebook)
+        self.notebook.add(self.tab_app_keys, text='Applications Keys')
+        self.notebook.add(self.tab_devices_management, text='Devices Management')
+        self.notebook.add(self.tab_about, text='About')
+        # self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
+        self.notebook.pack(expand=1, fill="both")
+
+        # Widgets
+        self.set_devices_management_widgets()
+        self.set_about_widgets()
+        self.set_applications_keys_widgets()
+
[…]

### 2020-06-28 testing/python/api-gui_v0.2/api-gui_v0.2.py
Improves api-gui_v0.3

@@ -69,7 +69,7 @@ class Application(tk.Frame):
         self.notebook.add(self.tab_app_keys, text='Applications Keys')
         self.notebook.add(self.tab_devices_management, text='Devices Management')
         self.notebook.add(self.tab_about, text='About')
-        # self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
+        self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
         self.notebook.pack(expand=1, fill="both")
 
         # Widgets

### 2020-06-28 testing/python/api-gui_v0.3/api-gui_v0.3.py

@@ -22,6 +22,14 @@ header_orion = {
 
 
 def use_api(url, method=requests.get, header=None, body=None):
+    """
+    Generic function to attack the FIWARE API
+    :param url:
+    :param method:
+    :param header:
+    :param body:
+    :return:
+    """
     body = json.dumps(body) if body else None
 
     response = method(url, headers=header, data=body)
@@ -31,12 +39,18 @@ def use_api(url, method=requests.get, header=None, body=None):
 
 
 def is_hex(string) -> bool:
+    """
+    Function to check whether a string is hexadecimal
+    :param string:
+    :return:
+    """
     return all([char in hexdigits for char in string])
 
 
 class Application(tk.Frame):
     entities = list()
-    default_entity = 'LORA-N-0'
+    # default_entity = 'LORA-N-0'
+    default_entity = 'node_0'
 
     devices = list()
     default_device = 'node_0'
@@ -92,6 +106,9 @@ class Application(tk.Frame):
         self.label_devices = tk.Label()
         self.label_entities = tk.Label()
         self.label_subscriptions = tk.Label()
+        self.label_orion_ver = tk.Label()
+        self.label_iot_ver = tk.Label()
+        self.label_quantumleap_ver = tk.Label()
 
         # Inputs
         self.entry_ttn_app_id = tk.Entry()
@@ -109,6 +126,10 @@ class Application(tk.Frame):
         self.listbox_devices = tk.Listbox()
 
     def set_about_widgets(self) -> None:
+        """
+        Sets up the About tab where versions are shown
+        :return:
+        """
         self.label_orion_ver = tk.Label(self.tab_about,
                                         text='Orion version: ' + use_api('http://localhost:1026/version')
                                         .json()['orion']['version'])
@@ -122,7 +143,10 @@ class Application(tk.Frame):
         self.label_quantumleap_ver.place(x=10, y=50)
 
     def set_applications_keys_widgets(self) -> None:
-
+        """
+        Set labals, entries and buttons for application keys tab
+        :return:
+        """
         # Labels
         self.label_ttn_app_id = tk.Label(self.tab_app_keys, text='TTN Application ID')
         self.label_ttn_app_pw = tk.Label(self.tab_app_keys, text='TTN Application Access Key')
@@ -131,7 +155,7 @@ class Application(tk.Frame):
         self.label_ttn_app_pw.place(x=10, y=50)
[…]

### 2020-06-28 testing/python/api-gui_v0.3/api-gui_v0.3.py
Includes an entry for device and entity ID in api-gui.py

@@ -48,12 +48,14 @@ def is_hex(string) -> bool:
 
 
 class Application(tk.Frame):
-    entities = list()
+    # entities = list()
     # default_entity = 'LORA-N-0'
-    default_entity = 'node_0'
 
-    devices = list()
-    default_device = 'node_0'
+    # devices = list()
+    # default_device = 'node_0'
+
+    devices = entities = list()
+    default_device = default_entity = 'node_0'
 
     subscriptions = list()
 
@@ -83,7 +85,7 @@ class Application(tk.Frame):
         self.notebook.add(self.tab_app_keys, text='Applications Keys')
         self.notebook.add(self.tab_devices_management, text='Devices Management')
         self.notebook.add(self.tab_about, text='About')
-        # self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
+        self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
         self.notebook.pack(expand=1, fill="both")
 
         # Widgets
@@ -94,36 +96,38 @@ class Application(tk.Frame):
         # Initial get request
         self.api_get_devices_and_entities()
 
-        # Labels
-        self.label_ttn_app_id = tk.Label()
-        self.label_ttn_app_pw = tk.Label()
-        self.label_ttn_app_eui = tk.Label()
-        self.label_app_skey = tk.Label()
-        self.label_dev_eui = tk.Label()
-        self.label_title_devices = tk.Label()
-        self.label_title_entities = tk.Label()
-        self.label_title_subscriptions = tk.Label()
-        self.label_devices = tk.Label()
-        self.label_entities = tk.Label()
-        self.label_subscriptions = tk.Label()
-        self.label_orion_ver = tk.Label()
-        self.label_iot_ver = tk.Label()
-        self.label_quantumleap_ver = tk.Label()
-
-        # Inputs
-        self.entry_ttn_app_id = tk.Entry()
-        self.entry_ttn_app_pw = tk.Entry()
-        self.entry_ttn_app_eui = tk.Entry()
-        self.entry_app_skey = tk.Entry()
-        self.entry_dev_eui = tk.Entry()
-
-        # Buttons
-        self.set_app_keys = tk.Button()
-        self.button_create_device = tk.Button()
-        self.button_delete_device = tk.Button()
-
-        # Listbox
-        self.listbox_devices = tk.Listbox()
+        # # Labels
[…]

### 2020-06-28 src/api-gui/api-gui.py
Improves and documents api-gui.py

@@ -1,29 +1,14 @@
+#!/usr/bin/python
+
 import requests
 import tkinter as tk
+from tkinter import ttk
+from string import hexdigits
 import json
-
-
-def use_api(url, method=requests.get, header=None, body=None):
-    body = json.dumps(body) if body else None
-
-    response = method(url, headers=header, data=body)
-    print(url, method, response.status_code, response.text)
-
-    return response
-
 
 # These are customizable patterns, actually
 fiware_service = 'atosioe'
 fiware_servicepath = '/lorattn'
-
-# These are application unique
-TTN_app_id = 'dendrometer'
-TTN_app_pw = 'ttn-account-v2.173lH8wwDiIRC8E2JgM9ScXyuNRlPpMefpazS0TIhnU'
-TTN_app_eui = '70B3D57ED0030B5D'
-
-# These are device unique
-TTN_app_skey = '7AF6B9A29EAF893F66C0F4360BA8DD5B'
-TTN_dev_eui = '00EC51D264A6F8FD'
 
 header_iot_device = {
     "Content-Type": "application/json",
@@ -38,65 +23,256 @@ header_orion = {
 }
 
 
-class Application(tk.Frame):
-    entities = list()
-    default_entity = 'LORA-N-0'
+def use_api(url, method=requests.get, header=None, body=None):
+    """
+    Generic function to attack the FIWARE API
+    :param url:
+    :param method:
+    :param header:
+    :param body:
+    :return:
+    """
+    body = json.dumps(body) if body else None
 
-    devices = list()
-    default_device = 'node_0'
+    response = method(url, headers=header, data=body)
+    # print(url, method, response.status_code, response.text)
+
+    return response
+
+
+def is_hex(string) -> bool:
+    """
+    Function to check whether a string is hexadecimal
+    :param string:
+    :return:
+    """
+    return all([char in hexdigits for char in string])
+
+
+class Application(tk.Frame):
+
+    devices = entities = list()
+    default_device = default_entity = 'node_0'
 
     subscriptions = list()
 
+    # These are application unique
+    TTN_app_id = None
+    TTN_app_pw = None
+    TTN_app_eui = None
+
+    # These are device unique
+    TTN_app_skey = None
+    TTN_dev_eui = None
+
     def __init__(self):
+        # TKINTER constructor
         self.gui = tk.Tk()
         self.gui.title('Fiware Client')
         super().__init__(self.gui)
-        self.gui.geometry("640x480")
[…]

### 2020-06-28 testing/python/api-gui_v0.3/api-gui_v0.3.py

@@ -1,3 +1,5 @@
+#!/usr/bin/python
+
 import requests
 import tkinter as tk
 from tkinter import ttk
@@ -48,11 +50,6 @@ def is_hex(string) -> bool:
 
 
 class Application(tk.Frame):
-    # entities = list()
-    # default_entity = 'LORA-N-0'
-
-    # devices = list()
-    # default_device = 'node_0'
 
     devices = entities = list()
     default_device = default_entity = 'node_0'
@@ -88,6 +85,39 @@ class Application(tk.Frame):
         self.notebook.tab(1, state='disabled')  # disabled by default until the app data is introduced
         self.notebook.pack(expand=1, fill="both")
 
+        # Labels
+        self.label_dev_ent_id = None
+        self.label_ttn_app_id = None
+        self.label_ttn_app_pw = None
+        self.label_ttn_app_eui = None
+        self.label_app_skey = None
+        self.label_dev_eui = None
+        self.label_title_devices = None
+        self.label_title_entities = None
+        self.label_title_subscriptions = None
+        self.label_devices = None
+        self.label_entities = None
+        self.label_subscriptions = None
+        self.label_orion_ver = None
+        self.label_iot_ver = None
+        self.label_quantumleap_ver = None
+
+        # Inputs
+        self.entry_dev_ent_id = None
+        self.entry_ttn_app_id = None
+        self.entry_ttn_app_pw = None
+        self.entry_ttn_app_eui = None
+        self.entry_app_skey = None
+        self.entry_dev_eui = None
+
+        # Buttons
+        self.set_app_keys = None
+        self.button_create_device = None
+        self.button_delete_device = None
+
+        # Listbox
+        self.listbox_devices = None
+
         # Widgets
         self.set_devices_management_widgets()
         self.set_about_widgets()
@@ -95,39 +125,6 @@ class Application(tk.Frame):
 
         # Initial get request
         self.api_get_devices_and_entities()
-
-        # # Labels
-        # self.label_dev_ent_id = tk.Label()
-        # self.label_ttn_app_id = tk.Label()
-        # self.label_ttn_app_pw = tk.Label()
-        # self.label_ttn_app_eui = tk.Label()
-        # self.label_app_skey = tk.Label()
-        # self.label_dev_eui = tk.Label()
-        # self.label_title_devices = tk.Label()
-        # self.label_title_entities = tk.Label()
-        # self.label_title_subscriptions = tk.Label()
-        # self.label_devices = tk.Label()
-        # self.label_entities = tk.Label()
[…]

### 2020-06-28 docs/LaTex/Paper.tex
Completes docs

@@
 
+\subsubsection{FIWARE API Graphical User Interface}
+Finally, to make more friendly the FIWARE interaction this project also provide a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
+This graphical interface has been developed using the well-known tkinter library in python, it requires also a few python modules that are specified in \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/pip_requirements.txt}{\cmd{/src/api-gui/pip\_requirements.txt}}, in fact the user should be able to install them by performing in a terminal the following command \cmd{pip install -r pip\_requirements.txt}.
+
+When the requirements are satisfied and all docker containers are running properly\footnote{In a first place the GUI will retrieve the versions for Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap, so if they are not running the GUI wont open.} the GUI can be run performing \cmd{./api-gui.py}
+
 \newpage

### 2020-06-28 docs/LaTex/Paper.tex
Completes docs

--- 2020-06-27 docs/LaTex/Paper.tex
@@
 
-\subsubsection{FIWARE}
+\subsubsection{FIWARE}\label{sssec:FIWARE}
 As shown in figure \ref{fig:GenView}, this solution is being implemented using particularly three FIWARE generic enablers, the most important one, of course, \textbf{Orion Context Broker}, also a LoRaWAN agent called \textbf{IoTagent-LoRaWAN} and finally to make data persistent, \textbf{QuantumLeap}.
@@
 
+\subsubsection{FIWARE API Graphical User Interface}
+Finally, to make more friendly the FIWARE interaction exposed in \hyperref[sssec:FIWARE]{FIWARE} subsection, this project also provide a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
+This graphical interface has been developed using the well-known tkinter library in python, it requires also a few python modules that are specified in \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/pip_requirements.txt}{\cmd{/src/api-gui/pip\_requirements.txt}}, in fact the user should be able to install them by performing in a terminal the following command \cmd{pip install -r pip\_requirements.txt}.
+
+When the requirements are satisfied and all docker containers are running properly\footnote{In a first place the GUI will retrieve the versions for Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap, so if they are not running the GUI wont open.} the GUI can be run performing \cmd{./api-gui.py}; this should open the window shown by figure \ref{fig:GUI_app_keys}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_app_keys.png}
+  \caption{First view of \cmd{api-gui.py}}
+  \label{fig:GUI_app_keys}
+\end{figure}
+
+To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without complete all the required info won't be enabled the \enquote{Devices Management} tab.
+Actually this tab performs a series of checks regard in TTN keys.
+Following the same TTN requirements,
+
+\begin{itemize}
+  \item TTN Application ID must be greater than 2 characters.
+  \item TTN Application Access Key must starts with \cmd{ttn-account} string.
+  \item TTN Application EUI must be 16 characters long and be hexadecimal. 
+\end{itemize}
+When any of this requirements are not fulfilled the related label will turn red as figure \ref{fig:GUI_app_keys_fail} shows
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_app_keys_fail.png}
+  \caption{TTN Application EUI must be hexadecimal} 
+  \label{fig:GUI_app_keys_fail}
+\end{figure}
+
 \newpage

### 2020-06-29 docs/LaTex/Paper.tex
Completes docs

@@
 This document gives a detailed description of this project, which is focused on researching possibles low-cost alternatives for wireless dendrometry systems.
-Currently there exist a lot of expensive and professional systems in the market, that's because this project is intended to reduce costs and increase the versatility, scalability and accessibility.
+Currently there exist a lot of expensive and professional systems in the market, that's why this project is intended to reduce costs and increase the versatility, scalability and accessibility.
 
@@
 
-Since the idea is to produce a low-cost device in order to distribute a high number of them, it must be a simple design; that's because it consist only of three parts,
+Since the idea is to produce a low-cost device in order to distribute a high number of them, it must be a simple design; that's why it consist only of three parts,
 
@@
 LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite[]{LoRaGeneral}.
-This is because some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
+This is why some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
 So this low availability (of resources) along with their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
@@
 \subsubsection{Raspberry Pi, a powerful microcomputer}
-As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, Raspberry Pi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's because this devices are usually called \textit{single-board computers}.
+As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, Raspberry Pi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's why this devices are usually called \textit{single-board computers}.
 
@@
 \subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
-These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) have not the ability to comunicate through LoRa physical layer, that's because they need a physical interface in order to manage those LoRa packages.
+These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) have not the ability to comunicate through LoRa physical layer, that's why they need a physical interface in order to manage those LoRa packages.
 Both shields are based on the SX1276/SX1278 transceiver.
@@
 \subsubsubsection{Linear Potentiometer}
-As said, ideally this potentiometer should be a linear potentiometer with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it's recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that's because the figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
+As said, ideally this potentiometer should be a linear potentiometer with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it's recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that's why the figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
 
@@
 
-It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead a variable resistor, this is because taking advantage of the Arduino power source, particularly the 5 \si{\volt} and GND pins.
+It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead a variable resistor, this is why taking advantage of the Arduino power source, particularly the 5 \si{\volt} and GND pins.
 
@@
 Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the context data will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persistent.
-Also a \textbf{Redis database} is used for geocoding purposes, however this project will not deal with it ---this is because of this database is not even being included in the figure \ref{fig:GenView}.
+Also a \textbf{Redis database} is used for geocoding purposes, however this project will not deal with it ---this is why of this database is not even being included in the figure \ref{fig:GenView}.
   \item \textbf{Grafana}, this container is intended to provide a web interface to visualize and manage the retrieved context data.
@@
 Anyway and after of this little introduction, the most important thing is to deploy all FIWARE generic enablers and its dependencies in a secure and stable way.
-This is because this project uses \cmd{docker} and \cmd{docker-compose} in order to ensure these modules deployment and execution.
+This is why this project uses \cmd{docker} and \cmd{docker-compose} in order to ensure these modules deployment and execution.
 Again this project provides the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/fiware/docker-compose.yml}{\cmd{docker-compose.yml}} file to ensure a correct deployment.
@@
 
+This information of course must match exactly with information provided by TTN, actually the \enquote{TTN Application Acess Key} and \enquote{TTN Application EUI} are provided by TTN, however the \enquote{TTN Application ID} must be chosen by the user when a new application is created using the TTN web interface, but this ID must be unique and it must be at least 3 characters long, that is why the developed GUI follows the same criteria.
+
+When the keys are updated properly the labels over the text fields are also updated with the chosen values and the \enquote{Devices Management} tab is enabled.
+This tab provides also a more friendly interface to manage devices.
+In a similar way this info will must match with the info provided by TTN interface, so that as figure \ref{fig:device_management_tab}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_decice_management.png}
+  \caption{Device Management tab}
+  \label{fig:device_management_tab}
+\end{figure}
+
+This view allows the user to create and delete devices ---along with its entities and subscriptions--- nevertheless, the \enquote{Delete device and entities} button remains disabled until a device is chosen in the right list.
+In a similar way than \enquote{Applications Keys}, to add a new device to FIWARE deployed instance by docker, all keys must fullfil a set of minimum requirements; as can be seen in picture \ref{fig:adding_dev}
+
+\begin{itemize}
+  \item Device entity and ID must be also unique and it has to be at least 3 characters long. Moreover, this ID must match exactly with the ID chosen in TTN interface when the device is created,
+  \item Application Session Key must be 32 characters long and also an hexadecimal string.
+  \item Device EUI has to be 16 characters long and again an hexadecimal string.
+\end{itemize}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_add_device.png}
+  \caption{Trying to add a new device}
+  \label{fig:adding_dev}
+\end{figure}
+
 \newpage

### 2020-06-29 docs/LaTex/Paper.tex
Finishes technical description

@@
 
-% To avoid 'underfull box' warning caused because bibliography line-break
+% To avoid 'underfull box' warning caused because of bibliography line-break
 \usepackage{etoolbox}
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../pictures/GUI_decice_management.png}
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_device_management.png}
   \caption{Device Management tab}
@@
 
-\begin{figure}[ht]
+These keys can be obtained from TTN as can be seen in screenshot \ref{fig:dev_overview} ---this screenshot shows info about \cmd{node\_0}, though.
+As can be seen in picture \ref{fig:adding_dev}, there was a problem with \enquote{Device EUI}, this was caused because of the final \cmd{Z} letter in the string which doesn't belong to hexadecimal characters set.
+
+\begin{figure}[htp]
   \centering
@@
 \end{figure}
+
+\begin{figure}[htp]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_about_tab.png}
+  \caption{Running versions in About tab.}
+  \label{fig:about_tab}
+\end{figure}
+
+Finally, as indicated, the \enquote{About} tab shows the versions info as can be seen in \ref{fig:about_tab}.
+The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why could be helpful to retrieve these.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/GUI_device_management_2nodes.png}
+  \caption{Graphical interface showing 2 created nodes}
+  \label{fig:GUI_2nodes}
+\end{figure}
+
+As figure \ref{fig:GUI_2nodes} shows, once a node is created is added to the list, subscription and entity ID are useful to remove manually just a subscription or an entity by performing a manual \cmd{DELETE} request, in fact have been also included several \cmd{DELETE} requests in the attached Postman collection, however, when a node is deleted using this graphical interface button is also being removed its entity and its subscription.
 

### 2020-06-29 docs/LaTex/Paper.tex
Finishes a first docs draft

@@
 \section{Conclusions}
+Finally there are a few considerations that could be done,
+
+\begin{itemize}
+  \item A rotary potentiometer does work well enough as a proof of concept, but would have been more appropriate to have the expected linear potentiometer to try its sensibility, which is a good point in order to measure such little variations as produced in the stem diameter in a tree.
+  \item The fact that the Raspberry Pi HAT uses a single channel chip (SX1276) has two important implications:
+    \begin{itemize}
+      \item In a first place, as indicated in its corresponding section, these kind of gateways are LoRaWAN compatible but not compliant.
+So it would be desirable to replace this HAT by another one, or simple to acquire a proper gateway for a production environment.
+      \item Nodes cannot comunicate simultaneously with the gateway because it is listening only one channel, so this is causing a time gap in data acquisition as can be seen in the picture \ref{fig:times_gap}
+      \begin{figure}[ht]
+        \centering
+        \includegraphics[width=.9\textwidth]{../pictures/Grafana_times_gap.png}
+        \caption{Little gap between times caused by the single channel chip.}
+        \label{fig:times_gap}
+      \end{figure}
+    \end{itemize}
+    \item Maybe the developed GUI could have handle also the context data plots, however this project couldn't have covered in anyway this development with the available time, so this is not considered as part of the project because Grafana is complete enough to this task for now and perhaps for ever due it is extremely powerful handling databases.
+    \item For a production environment it seems clear that arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems, this would lead a very compact device which could be also improved adding more sensor due versatility of CayenneLPP encoding standard which allows to send multiple measurements of different sensors.
+    \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; so it likely keeps developing, depending on the needs of the researchers.
+\end{itemize}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/2prototypes_1gateway.jpg}  
+  \caption{2 prototypes transmitting to the gateway}
+  \label{fig:2nodes_gateway}
+\end{figure}
 

### 2020-06-29 docs/LaTex/Paper.tex
Minor fixes in the first docs draft

@@
 So it would be desirable to replace this HAT by another one, or simple to acquire a proper gateway for a production environment.
-      \item Nodes cannot comunicate simultaneously with the gateway because it is listening only one channel, so this is causing a time gap in data acquisition as can be seen in the picture \ref{fig:times_gap}
+      \item Secondly, Nodes cannot comunicate simultaneously with the gateway because it is listening only one channel, so this is causing a time gap in data acquisition as can be seen in the picture \ref{fig:times_gap}
       \begin{figure}[ht]

### 2020-06-29 docs/LaTex/Paper.tex
Minor fixes

@@
 
-\enquote{Host} must be \cmd{crate:5432} because is the tcp port that grafana uses to queries, this can be seen also in figure \ref{fig:docker_containers} another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
+\enquote{Host} must be \cmd{crate:5432} because is the port which grafana uses to queries under postgres protocol, this can be seen also in figure \ref{fig:docker_containers} another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
 

### 2020-06-29 docs/LaTex/Paper.tex
Fixes in docs

@@
 Basically the sketch is based on the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino}{ttn-abp example}, as the reader can see in the first comment block; that example sends a LoRaWAN valid package with the payload \enquote{Hello World!}.
-Besides it uses the ABP (Activation-by-personalization) method ---in opposition to OTAA (Over-the-air-activation) method---, depending on the used method it is necessary to include different security keys provided by The Things Network stack (TTN stack) in the code.
-\cite{TTN_Security}
+Besides it uses the ABP (Activation-by-personalization) method ---in opposition to OTAA (Over-the-air-activation) method---, depending on the used method it is necessary to include different security keys provided by The Things Network stack (TTN stack) in the code \cite{TTN_Security}.
+\textbf{It is important to note that frame counters in TTN console must be reset every time the device is reset, powered off or flashed again} \cite{TTN_frame_counters}.
 
@@
     \end{itemize}
+    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in the case that downlinks packets were transmitted, this is why due some security implications frame counters must be reset for ABP devices every time they are shutdown for any circumstance; in particular when downlinks are transmitted, could be necessary to reset the frame counter in the devices, so in this situation would be interesting to consider swap to OTAA method.
+In fact, TTN offers the possibility to reset the frame counter in TTN console but this could be removed in a future.
     \item Maybe the developed GUI could have handle also the context data plots, however this project couldn't have covered in anyway this development with the available time, so this is not considered as part of the project because Grafana is complete enough to this task for now and perhaps for ever due it is extremely powerful handling databases.
@@
   \centering
-  \includegraphics[width=.9\textwidth]{../pictures/2prototypes_1gateway.jpg}  
-  \caption{2 prototypes transmitting to the gateway}
+  \includegraphics[width=.8\textwidth]{../pictures/2prototypes_1gateway.jpg}  
+  \caption{Two prototypes transmitting through the gateway}
   \label{fig:2nodes_gateway}

### 2020-06-30 docs/LaTex/Paper.tex
Completes introduction section in docs

@@
 
+\subsubsection{The Things Network}
+TTN provides a backend system to route the traffic between devices and applications.
+It is basically a network server placed between gateways and applications servers.
+It is necessary because LoRaWAN is a non-IP protocol, but a MAC layer protocol and does need some for of routing and processing messages before they can reach the application side.
+TTN takes care of these steps.
+
+Moreover, this project is based on FIWARE (described in the next subsection) and the LoRaWAN agent in FIWARE is developed to interact with TTN infrastructure,
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/TTN_overview.png}
+  \caption{TTN basic diagram.}
+  \label{fig:TTN_overview}
+\end{figure}
+
+End-node devices broadcast LoRaWAN packets over LoRa radio protocol to the gateways, this gateways will forward these packets to the TTN backend, where a software module called Router will manage the gateway status and schedule transmissions, these Routers are connected to another software modules called Brokers, that are responsible for map a device to an application ---into TTN environment--- allowing downlinks and uplinks transmissions.
+Once these uplinks transmission are inside TTN environment, the FIWARE LoRaWAN agent will retrieve them, however this is actually the application server shown in figure \ref{fig:TTN_overview}.
+\enquote{The goal of The Things Network is to be very flexible in terms of deployment options.} \cite{TTN_net_arch}.
+As can be read in the docs, TTN infrastructure provides different \enquote{core functions} divided in four main groups:
+
+\begin{itemize}
+  \item Gateway-related functions. In order to schedule and manage the gateways.
+  \item Device-related functions. To manage the state and received context data from specific devices.
+  \item Application-related functions, to group different devices used for similar purposes.
+  \item Service discovery functionality, which helps components to determine where traffic should be routed to. 
+\end{itemize}
+
+After this, the traffic is routed to FIWARE side, which is deployed locally and the user has total control over it.
+
 \subsubsection{FIWARE}
@@
 \textit{Orion Context Broker} is designed to manage this context data through concepts such as \textit{subscriptions} or \textit{entities}, to name but a few; however, all these \textit{Generics Enablers} communicate with each other using the \textit{FIWARE NGSI RESTful API}\cite{NGSI}.\footnote{\textit{RESTful} term comes from the software architectural style \textit{REST}, which stands for \textit{Representational State Transfer}}
+
+\subsubsection{Simple Graphical Interface}
+Despite FIWARE is a powerful platform, is not friendly enough regarding to final user.
+This is why a the end of this document is described a simple graphical interface developed to manage end-nodes in FIWARE side.
+As will be indicated, this graphical interface has been developed using python and its tkinter module intended to build graphical interfaces.
 

### 2020-07-01 testing/docs/LaTex/Paper.tex
Adds docs to testing folder to make a second version

--- 2020-06-30 docs/LaTex/Paper.tex
@@
 \subsubsection{Simple Graphical Interface}
-Despite FIWARE is a powerful platform, is not friendly enough regarding to final user.
+Despite FIWARE is a powerful platform, is not friendly enough regarding final user.
 This is why a the end of this document is described a simple graphical interface developed to manage end-nodes in FIWARE side.
@@
 
-Regarding to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{pleas, refer to that subsection for further details}.
+With regards to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{pleas, refer to that subsection for further details}.
 The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
@@
 This is a platform for the LoRaWAN protocol where is possible to manage gateways, devices and applications.
-It provides a centralized way to manage these kind of things regard in security too, generating several keys to tight the end-devices to the chosen server cloud solution (FIWARE in this case).
+It provides a centralized way to manage these kind of things regarding security too, generating several keys to tight the end-devices to the chosen server cloud solution (FIWARE in this case).
 There are three fundamental process the reader must follow to include nodes and gateways in the platform, those are luckily pretty well documented.
@@
 
-Once these steps are completed, the TTN web interface will provide an overview with all data regarding to the device like figure \ref{fig:dev_overview} shows.
+Once these steps are completed, the TTN web interface will provide an overview with all data regarding the device like figure \ref{fig:dev_overview} shows.
 
@@
 To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without complete all the required info won't be enabled the \enquote{Devices Management} tab.
-Actually this tab performs a series of checks regard in TTN keys.
+Actually this tab performs a series of checks regarding TTN keys.
 Following the same TTN requirements,

### 2020-07-01 docs/LaTex/Paper.tex
Completes docs

@@
 % \renewcommand{\maketitlehooka}{\headingfont}
-
-% To change TOC appearance; see 2.1 section in tocloft.pdf for info about 'titles' option
-% it does not mention fancyhdr package but it do the work to flushright page numbers for 
-% table of contents pages. see: https://tex.stackexchange.com/a/137301/83189
-% \usepackage[titles]{tocloft}
-\usepackage{tocloft}
-\renewcommand{\cftsecfont}{\headingfont\bfseries}
-\renewcommand{\cftsecpagefont}{\headingfont\bfseries}
-\renewcommand{\cftsubsecfont}{\headingfont}
-\renewcommand{\cftsubsecpagefont}{\headingfont}
-\renewcommand{\cftsubsubsecfont}{\headingfont}
-\renewcommand{\cftsubsubsecpagefont}{\headingfont}
-\tocloftpagestyle{fancy} % However, I find better this way bc 'titles' option does not 
-% produce a compact TOC, it does expand bc probably it uses the same spacing rules in 
-% the TOC than in (subsub)(sub)sections titles at the rest of the text.
-% -------------------
-% Remove the TOC title (it is redundant if you have included a TOC section inside TOC)
-\makeatletter
-\renewcommand{\@cftmaketoctitle}{}
-\makeatother
-% But in the case you still want to print the TOC title, this other one modifies TOC title font
-\renewcommand{\cfttoctitlefont}{\Large\headingfont\bfseries}
 
@@
 % To include References and TOC in the TOC
-\usepackage[]{tocbibind}
+\usepackage[nottoc]{tocbibind}
 
@@
 \setlength{\headheight}{14pt}% ...at least 51.60004pt
+\fancypagestyle{plain}{%
+  \fancyhf{}% clear all header and footer fields
+  \fancyfoot[R]{\thepage}%
+  \renewcommand{\headrulewidth}{0pt}%
+  \renewcommand{\footrulewidth}{0pt}%
+}
+
+% To change TOC appearance; see 2.1 section in tocloft.pdf for info about 'titles' option
+% it does not mention fancyhdr package but it do the work to flushright page numbers for 
+% table of contents pages. see: https://tex.stackexchange.com/a/137301/83189
+% \usepackage[titles]{tocloft}
+\usepackage{tocloft}
+\renewcommand{\cftsecfont}{\headingfont\bfseries}
+\renewcommand{\cftsecpagefont}{\headingfont\bfseries}
+\renewcommand{\cftsubsecfont}{\headingfont}
+\renewcommand{\cftsubsecpagefont}{\headingfont}
+\renewcommand{\cftsubsubsecfont}{\headingfont}
+\renewcommand{\cftsubsubsecpagefont}{\headingfont}
+\tocloftpagestyle{empty} % However, I find better this way bc 'titles' option does not 
+% produce a compact TOC, it does expand bc probably it uses the same spacing rules in 
+% the TOC than in (subsub)(sub)sections titles at the rest of the text.
+% -------------------
+% Remove the TOC title (it is redundant if you have included a TOC section inside TOC)
+\makeatletter
+\renewcommand{\@cftmaketoctitle}{}
+\makeatother
+% But in the case you still want to print the TOC title, this other one modifies TOC title font
+% \renewcommand{\cfttoctitlefont}{\Large\headingfont\bfseries}
 
@@
 \title{\headingfont%
-  University of Extremadura\\
-  Faculty of Science\\[1cm]
+  UNIVERSIDAD DE EXTREMADURA\\
+  FACULTAD DE CIENCIAS\\[1cm]
   \textbf{%
-  	Physics degree\\
-    Degree Final Project\\[3cm]
+  	Grado en Física\\
+    TRABAJO DE FIN DE GRADO\\[3cm]
     \flushleft{\large{Developement of a FIWARE-based application for tree species monitoring (dendrometry)}}\\[3cm]
@@
 \begin{document}
+\pagenumbering{Roman} 
 \noindent % to prevent the indentation before first minipage and be able to expand minipages
@@
 \hskip 30pt minus 5pt % It is a dynamic space (glue) that starts on 30pt and is 5pt shrinkable
-\begin{minipage}[t]{.6\textwidth}
+\begin{minipage}[t]{.65\textwidth}
   \centering
@@
 \newpage
-% \pagenumbering{arabic}% Arabic page numbers (and reset to 1)
-\setcounter{page}{1}
-\tableofcontents
+
+\newpage\thispagestyle{empty}
+\vspace*{\fill}
+Fernando Javier Álvarez Franco, profesor del Departamento de Ingeniería Eléctrica, Electrónica y Automática de la Universidad de Extremadura.
+
+Informa:\\ Que D. Javier Fernández Aparicio ha realizado bajo su dirección el Trabajo de Fin de Grado. Considera que la memoria reúne los requisitos necesarios para su evaluación
+
+\begin{center}
+  Badajoz, 8 de julio de 2020
+\end{center}
+\vspace{2cm}
+\begin{center}
+  Fdo. Fernando J. Álvarez Franco.
+\end{center}
+\vspace*{\fill}
+%
+
+\newpage\thispagestyle{empty}
+\vspace*{.5cm}
 
 \newpage
+\pagestyle{empty}\tableofcontents\clearpage\pagestyle{fancy} % set the pagestyle as empty for all toc pages
+
+\newpage\thispagestyle{plain}\setcounter{page}{1}
+\pagenumbering{arabic}% Arabic page numbers (and reset to 1)
+\phantomsection\section*{Resumen}
+Este documento proporciona una descripción detallada del proyecto, el cual está centrado en la investigación de posibles alternativas de bajo coste para sistemas de dendrometría inalámbricos.
+Actualmente existen una amplia gama de sistemas profesionales en el mercado, sin embargo debido a su elevado coste, este proyecto pretende abordar la reducción en el mismo, incrementando la versatilidad, accesibilidad y escalabilidad de este tipo de sistemas.
+
+Para conseguir estos objetivos el proyecto se apoyará en tecnologías de hardware libre como Arduino \cite{Arduino} o Raspberry Pi \cite{Raspberrypi}, y en sistemas de software libre o código abierto, como FIWARE \cite{FIWARE}, The Things Network \cite{TTN} y Python \cite{Python}.
 
@@
 
-In order to reach these objetives the project will be supported over free software such as FIWARE\cite{Fiware} or free hardware such as Arduino\cite{Arduino} and Raspberry Pi\cite{Raspberrypi}.
-
-\section{Introduction}
-This project arises itself from a direct interaction with professionals inside forestal sector.
+In order to reach these objetives the project will be supported with free software technologies such as FIWARE\cite{FIWARE}, The Things Network\cite{TTN} and Python\cite{Python} or free hardware systems such as Arduino\cite{Arduino} and Raspberry Pi\cite{Raspberrypi}.
+
+\newpage
+\section{Introduction}\thispagestyle{plain}
+This project arises itself from a direct interaction with professionals inside forestry sector.
 The original idea was to give technical coverage for particular necessities which professionals in this sector had to face off with.
 At this point is easy to notice this solution will need to be a distributed solution, due high samples dispersion.
-As indicated, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts \doubt{between $157$-$170$ indviduals per hectare}\cite{ForestStandVol} (depending on the used model).
-So according to this and sample size determination theories, to get a great resolution could be necessary a big size for samples and the necessity of a big wireless network of distributed devices, since each device will correspond with an individual.
-
-This is more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
-However, also according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept, but according to \cite[p.~2(920)]{IoTObjetives}
+As indicated, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts between $157$-$170$ individuals per hectare\cite{ForestStandVol} (depending on the used model).
+
+So according to indicated above and these sample size determination theories, to get a great resolution could be necessary a big size for samples and the need for a big wireless network of distributed devices, since each device will correspond with an individual.
+
+\subsection{Internet of Things}
+This previously exposed is, more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
+However, also according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept yet, but according to \cite[p.~2(920)]{IoTObjetives}
 
@@
 
-Over the years there have existed analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way.
-This could turn out bothering because the big size for this statistical population, as previously stated.
+Over the years there have existed analog and handled by hand dendrometers, thus data acquisition had to follow a manual process in the same way.
+This could turn out bothering because of the big size for this statistical population, as previously stated.
 So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.
@@
 
-This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point could be interesting to extend this project to include also a sensor to heights measurement.\footnote{This project is already considered extensive enough.}
+This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point could be interesting to extend this project to include also a sensor to heights measurement; however, this project is already considered extensive enough.
 
@@
 This project is focused on the former kind, so for this project is being developed a \enquote{contact dendrometer} based on a linear potentiometer.
+
+\subsection{Objectives}
+As previously stated the Abstract section, the most important objective for this project is to research about the possibility of make this dendrometry system cheaper and more scalable and accessible.
+
+Forests are remote and large surface areas which will require long-range communication technologies to be controlled.
+That's why LoRa is the perfect wireless technology to achieve this and these kind of automatic systems seem to be the proper solution to this hard and slow process.
+Hence IoT has much to contribute to this task.
+Forestry professionals could take advantage of this technologies to make this in an easier, cheaper and automated way; being this the main purpose of this project.
+
+\newpage\thispagestyle{plain}
+\section{System Architecture}
+\doubt{COMPLETE}
 
@@
 This is why some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
-So this low availability (of resources) along with their tendency to be distributed/scattered causes the necessity for a low-power consumption and a long-range telecommunication.
+So this low availability (of resources) along with their tendency to be distributed/scattered causes the need for a low-power consumption and a long-range telecommunication.
 
@@
 
-In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances:
+In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances; a comparative diagram shows different wireless technologies in figure \ref{fig:LoraComparison}
 
@@
 \subsubsection{FIWARE}
-It is defined as \enquote{The Open Source Platform for Our Smart Digital Future}\cite{Fiware}; in a deeper sense is \enquote{an open source initiative defending an universal set of standards for context data management}\cite{Fiware}, which basically means that \textit{FIWARE} is an open source platform where develop IoT and smart solutions.
+It is defined as \enquote{The Open Source Platform for Our Smart Digital Future}\cite{FIWARE}; in a deeper sense is \enquote{an open source initiative defending an universal set of standards for context data management}\cite{FIWARE}, which basically means that \textit{FIWARE} is an open source platform where develop IoT and smart solutions.
 The platform provides a number of software modules called \textit{Generic Enablers}; among all those \textit{Generic Enablers} there is one particularly important called \textit{Orion Context Broker}.
@@
 \newpage
-\section{Technical description}
+\section{Design}\thispagestyle{plain}
 Like in the Introduction, the most proper way to present this technical description is to difference between software and hardware parts.
@@
 
-\subsubsubsection{Arduino sketch}
+\subsubsubsection{Arduino}
 In a first place, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino}{the sketch} can be found in \cmd{src/dendro/dendro.ino} path.
@@
 
-It is important to note the necessity of a relatively stable internet connection; those gateways will forward the traffic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.
+It is important to note the need for a relatively stable internet connection; those gateways will forward the traffic to the cloud server, so they are receiving LoRa packets through its LoRa interface and then forwarding them to the specified server.
 
@@
   \item \textbf{Grafana}, this container is intended to provide a web interface to visualize and manage the retrieved context data.
-It stores its configuration and customization values in an embebed sqlite3 database; so there is not necessity for a separate container.
+It stores its configuration and customization values in an embebed sqlite3 database; so there is not need for a separate container.
 \end{itemize}
@@
 
-\newpage
+\newpage\thispagestyle{plain}
+\section{Proof of concept}
+\doubt{COMPLETE}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.8\textwidth]{../pictures/2prototypes_1gateway.jpg}  
+  \caption{Two prototypes transmitting through the gateway}
+  \label{fig:2nodes_gateway}
+\end{figure}
+
+\newpage\thispagestyle{plain}
 \section{Conclusions}
@@
 
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=.8\textwidth]{../pictures/2prototypes_1gateway.jpg}  
-  \caption{Two prototypes transmitting through the gateway}
-  \label{fig:2nodes_gateway}
-\end{figure}
-
 \newpage

### 2020-07-01 docs/LaTex/Paper.tex
Sorting files and folders

@@
 \section{System Architecture}
-\doubt{COMPLETE}
-
+A simple block diagram is shown in figure \ref{fig:block_diagram}, this diagram intends to emphasize the difference between two big layers, a physical layer that should be placed ---IoT dendrometers at least--- close to the sample, and another one called abstract layer which can be deployed locally ---TTN stack is a cloud server, although.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../schemes/simple_scheme_tbg.png}
+  \caption{Block diagram overview}
+  \label{fig:block_diagram}
+\end{figure}
+
+The following subsections will detail each of one.
+As can be seen in \ref{fig:block_diagram}, the physical layer is made up of two main modules, the first one is an Arduino microcontroller with its LoRa interface and a linear potentiometer acting as a sensor ---this whole set is the IoT dendrometer--- and the second one; a Raspberry Pi 3b+ with its LoRa interface to act as a gateway.
+
+Meanwhile, the abstract layer is made up of three software modules.
+TTN stack is a cloud server to retrieve LoRa packets from all existent gateways, FIWARE is a local application server to gather all this context data; and finally Grafana to plot the context data.
 \subsection{Physical Layer}
@@
 \begin{itemize}
-  \item \href{https://es.rs-online.com/web/p/products/0317780/}{\doubt{Linear potentiometer:}} which is the sensor itself due is directly in contact with the stem.
-In order \doubt{to improve data acquisition it will be necessary to use a High Input Impedance Amplifier.}
+  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself due is directly in contact with the stem.
+In order to improve data acquisition it will be necessary to use a High Input Impedance Amplifier.
   \item Arduino microcontroller: this is the core part for the device, it will be responsible for acquire linear potentiometer data and send it to a gateway through LoRa interface. 
@@
 
-In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances; a comparative diagram shows different wireless technologies in figure \ref{fig:LoraComparison}
-
 \begin{figure}[htp]
@@
 \end{figure}
+
+In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances; a comparative diagram shows different wireless technologies in figure \ref{fig:LoraComparison}
 
@@
 
-% \newpage
-% \section{Background}
+There are a lot of HTTP clients available, however the reader would be still forced to make manually these HTTP requests, these could result a little complex for a non advanced user, that's why this graphical interface has been developed.
+
+It's not perfect, it has a few deficiencies in fact, but is by the moment an appropriate way to manage devices, entities and subscriptions in locally deployed FIWARE instance.
+\href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{The code} of this graphical interface is available in the project repository at GitHub.
+
 \newpage
@@
   \item A linear potentiometer, ideally an \href{https://docs.rs-online.com/37bf/0900766b814f0bd0.pdf}{RS Pro Conductive Polymer
-  Potentiometer for Automotive Applications}, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, \doubt{this project will implement a regular rotary potentiometer.}
-  \item \doubt{A signal conditioning stage, which is needed in order to improve the quality of the voltage signal produced by the potentiometer.}
+  Potentiometer for Automotive Applications}, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, this project will implement a regular rotary potentiometer.
+  \item A signal conditioning stage, which is needed in order to improve the quality of the voltage signal produced by the potentiometer.
   \item Arduino + Dragino LoRa Shield; of course it is necessary to schedule data transmission and acquisition, as well as transmit them using the LoRa physical layer and LoRaWAN protocol.

### 2020-07-01 docs/LaTex/Paper.tex
Completes docs and minor fixes

@@
 % To avoid 'underfull box' warning caused because of bibliography line-break
-\usepackage{etoolbox}
-\apptocmd{\thebibliography}{\raggedright}{}{}
+% \usepackage{etoolbox}
+% \apptocmd{\thebibliography}{\raggedright}{}{}
 
@@
 \usepackage{fancyhdr}
-% Turn on the style
-\pagestyle{fancy}
+% Plain style that removes headers but keeps page numbers
+\fancypagestyle{plain}{%
+  \fancyhf{} % Clears all header and footer fields
+  \fancyfoot[R]{\thepage}%
+  \renewcommand{\headrulewidth}{0pt}%
+  \renewcommand{\footrulewidth}{0pt}%
+}
+% This style is my custom style
+\fancypagestyle{myfancy}{%
 	\fancyhf{} % Start with clearing everything in the header and footer
-\renewcommand{\headrulewidth}{1pt} % Removes the horizontal rule in header by default
+	\renewcommand{\headrulewidth}{1pt} % Makes 1pt width the horizontal rule in header by default
 	\fancyfoot[R]{\thepage} % Set the right side of the footer to be the page number
@@
 	\setlength{\headheight}{14pt}% ...at least 51.60004pt
-\fancypagestyle{plain}{%
-  \fancyhf{}% clear all header and footer fields
-  \fancyfoot[R]{\thepage}%
-  \renewcommand{\headrulewidth}{0pt}%
-  \renewcommand{\footrulewidth}{0pt}%
-}
+}
+\pagestyle{myfancy} % Enables by default the custom style.
 
@@
 \endgroup
-\newpage
-
-\newpage\thispagestyle{empty}
+
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{empty}
 \vspace*{\fill}
@@
 
-\newpage\thispagestyle{empty}
-\vspace*{.5cm}
-
-\newpage
-\pagestyle{empty}\tableofcontents\clearpage\pagestyle{fancy} % set the pagestyle as empty for all toc pages
-
-\newpage\thispagestyle{plain}\setcounter{page}{1}
-\pagenumbering{arabic}% Arabic page numbers (and reset to 1)
+\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{empty}
+
+\tableofcontents\cleardoublepage\pagestyle{myfancy} % set the pagestyle as empty for all toc pages
+
+\newpage\thispagestyle{plain}\pagenumbering{arabic}% Arabic page numbers (and reset to 1)
 \phantomsection\section*{Resumen}
@@
 
-\newpage
+\clearpage{\pagestyle{empty}\cleardoublepage}
 \section{Introduction}\thispagestyle{plain}
@@
 
-\newpage\thispagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}
 \section{System Architecture}
@@
 
-In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances; a comparative diagram shows different wireless technologies in figure \ref{fig:LoraComparison}
+In comparison with other common telecommunication technologies, LoRa uses lower bandwidth, but it reaches longer distances; a comparative diagram shows different wireless technologies in figure \ref{fig:LoraComparison}.
 
@@
 
-\newpage
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}
 \section{Design}\thispagestyle{plain}
@@
 
-\newpage\thispagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}
 \section{Proof of concept}
@@
 
-\newpage\thispagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{plain}
 \section{Conclusions}
@@
 
-\newpage
-\phantomsection
-\addcontentsline{toc}{section}{References}
+\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{myfancy}
+\phantomsection\addcontentsline{toc}{section}{References}
 \printbibliography

### 2020-07-01 docs/LaTex/Paper.tex
Minor fixes

@@
 
-\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{empty}
-
+\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{empty} % Forces the next section to start in an odd page.
 \tableofcontents\cleardoublepage\pagestyle{myfancy} % set the pagestyle as empty for all toc pages
 
-\newpage\thispagestyle{plain}\pagenumbering{arabic}% Arabic page numbers (and reset to 1)
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain} % Forces the next section to start in an odd page.
+\pagenumbering{arabic}% Arabic page numbers (and reset to 1)
 \phantomsection\section*{Resumen}
@@
 
-\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain} % Forces the next section to start in an odd page.
 \section{System Architecture}
@@
 
-\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain} % Forces the next section to start in an odd page.
 \section{Design}\thispagestyle{plain}
@@
 
-\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}  % Forces the next section to start in an odd page.
 \section{Proof of concept}
@@
 
-\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{plain}
+\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{plain}  % Forces the next section to start in an odd page.
 \section{Conclusions}
@@
 
-\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{myfancy}
+\clearpage{\pagestyle{empty}\cleardoublepage}\pagestyle{myfancy} % Forces the next section to start in an odd page.
 \phantomsection\addcontentsline{toc}{section}{References}

### 2020-07-02 docs/LaTex/Paper.tex
Completes docs

@@
 
-\subsubsection{Gateway}
+\subsubsection{Gateway}\label{sssec:Gateway}
 
@@
 
-\subsubsection{Grafana}
+\subsubsection{Grafana}\label{sssec:Grafana}
 Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; due it's actually capable to acquire the data from multiple sources.
@@
 
-\subsubsection{FIWARE API Graphical User Interface}
+\subsubsection{FIWARE API Graphical User Interface}\label{sssec:API_GUI}
 Finally, to make more friendly the FIWARE interaction exposed in \hyperref[sssec:FIWARE]{FIWARE} subsection, this project also provide a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
@@
 \clearpage{\pagestyle{empty}\cleardoublepage}\thispagestyle{plain}  % Forces the next section to start in an odd page.
-\section{Proof of concept}
-\doubt{COMPLETE}
-
-\begin{figure}[ht]
-  \centering
-  \includegraphics[width=.8\textwidth]{../pictures/2prototypes_1gateway.jpg}  
+\section{Proof of concept, deployment}
+Due circumstances at early 2020, this section has been reduced, instead of test the system in a real production environment, this section will focus on summarize needed steps/stages to deploy the system properly.
+Although these steps are already detailed separately, this section tries to put them together to give a consistent and practical point of view.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/2prototypes_1gateway.jpg}  
   \caption{Two prototypes transmitting through the gateway}
@@
 \end{figure}
+
+\subsection{Set up the gateway}
+This process has been widely detailed in \ref{sssec:Gateway} subsection.
+The most important point on this stage is to ensure the systemd service is working properly and to retrieve the Gateway ID, which will be needed in the next stage.
+
+\subsection{TTN Stack}
+The first thing that must be done is to follow the three steps described in subsection \ref{sssec:TTN_Stack} ---Register a Gateway using its \enquote{Gateway ID}, Add an application and register a device for that application; this is the first step because the reader will need the security keys to two following stages:
+
+\begin{itemize}
+  \item Embebe these keys in Arduino sketch. This sketch will require the \enquote{Application Session Key}, \enquote{Network Session Key} and \enquote{Device Address}.
+  \item Use them to create devices, and subscriptions in FIWARE.
+To create the device ---and also its corresponding entity and subscription when using the designed graphical interface shown in \ref{sssec:API_GUI} subsection--- the IoTagent will require the \enquote{Application ID}, \enquote{Application Access Key}, \enquote{Device ID}, \enquote{Application EUI} and again \enquote{Application Session Key}.
+\end{itemize}
+
+Once these keys are available in TTN platform, the process can continue.
+
+\subsection{Customize the Arduino sketch for a specific device}
+As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}, this latter is the most important key, because several nodes chan use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check those are different for every node created in TTN platform.
+
+Of course, to be able to compile the sketch some dependencies must be satisfied, such as CayenneLPP ArduinoJson and lmic-arduino libraries.
+Moreover, the \textbf{linear potentiometer has been replaced for a rotary potentiometer}, due circumstances again.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.9\textwidth]{../pictures/prototype.jpg}
+  \caption{Prototype broadcasting LoRa packets.}
+  \label{fig:prototype}
+\end{figure}
+
+It's important to note that has been chosen the analog pin 2 for analog signal acquisition.
+Besides the reader can check the status of the device in serial monitor, however, the reader must also note that serial monitor is setup at 115200 bits per second.
+
+\subsection{Deploy Docker containers}
+Once the hardware has been properly setup, the next step is deploy locally all Docker containers, so as indicated in \ref{sssec:FIWARE} subsection, using \cmd{docker-compose} is possible to do this easily.
+
+This will deploy essentially the IoTagent-LoRaWAN, Orion CB, QuantumLeap and Grafana. The status of these containers can be checked by performing \cmd{docker ps -a}.
+
+\subsection{Devices, entities and subscriptions}
+When Docker containers are deployed, it should be possible to send HTTP requests to create a device (with its entity) and a subscription to notify QuantumLeap at any change for a particular entity.
+This process can be don by two different ways as indicated previously,
+
+\begin{itemize}
+  \item Using developed graphical user interface (recommended, see \ref{sssec:API_GUI} subsection)
+  \item Using \cmd{curl} or Postman to perform directly the HTTP requests
+\end{itemize}
+
+\subsection{Grafana}
+Once any device (with an entity) and particularly its subscription are properly created, CrateDB should start journaling the context data, therefore next step is to add CrateDB as a data source in Grafana.
+
+So if the data source is properly added, it should be possible to create a new panel into a new dashboard, setting up the proper query for that panel, Grafana should start to plot the context data.
+(see \ref{sssec:Grafana} subsection for further details)
+
+Finally the system it would be completely deployed and journaling the context data, as can be seen in figure \ref{fig:final_panel}.
 

### 2020-07-02 docs/LaTex/Paper.tex
Minor fixes

@@
 \begin{itemize}
-  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, \doubt{(I'm not sure about this cite (It's a forum))} so this means this hat is designed to be a node-class radio, not a gateway.}
+  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, so this means this hat is designed to be a node-class radio, not a gateway.}
   \item \textit{LoRa Shield for Arduino}\cite{DraginoArdShield}, which is plugged through analog and digital pins.
@@
 So, as can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
-%So, as can be seen, in order to prepare the payload, the first is to read the analog pin and store the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type in \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}--- after this, the \cmd{potVal} variable is scaled to a $[0,1.0)$ range, this is due the function \cmd{LMIC\_f2sflt16()} (line 195), which is essentially a kind of encoder to encode floats to 16-bit fields.
-The working of this function is explained in \cite{MCCI_lmic_16bitencoder}.\footnote{A javascript decoder is also provided by the \href{https://github.com/mcci-catena/arduino-lmic\#uflt16-javascript-decoder}{\cmd{README.md}} but \doubt{it will be described in next sections.}}
 
@@
 This range will fit in the range that can be encoded in CayenneLPP.
-%This encoded 16-bit integer variable is stored in \cmd{uint16\_t payloadPot}, after this the \cmd{payloadPot} 16-bit integer is split into two bytes called \cmd{potLow} and \cmd{potHigh} ---using native \cmd{lowByte} and \cmd{highByte} functions.
-Finally these bytes are stored in the first two elements of initial \cmd{payload[3]} string declared at line 65, as the reader can see in example \ref{ej:potmeter_vars}, and the payload is prepared to send at the next possible time using the function \cmd{LMIC\_setTxData2()}.
 

### 2020-07-03 docs/LaTex/Paper.tex
Minor fixes
Redaction review

@@
 \addcontentsline{toc}{section}{Abstract}%
-This document gives a detailed description of this project, which is focused on researching possibles low-cost alternatives for wireless dendrometry systems.
-Currently there exist a lot of expensive and professional systems in the market, that's why this project is intended to reduce costs and increase the versatility, scalability and accessibility.
+This document gives a detailed description of this project, which is focused on researching possible low-cost alternatives for wireless dendrometry systems.
+Currently there are a lot of expensive and professional systems in the market, that's why this project is intended to reduce costs and increase the versatility, scalability and accessibility.
 
@@
 This project arises itself from a direct interaction with professionals inside forestry sector.
-The original idea was to give technical coverage for particular necessities which professionals in this sector had to face off with.
-At this point is easy to notice this solution will need to be a distributed solution, due high samples dispersion.
-As indicated, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts between $157$-$170$ individuals per hectare\cite{ForestStandVol} (depending on the used model).
-
-So according to indicated above and these sample size determination theories, to get a great resolution could be necessary a big size for samples and the need for a big wireless network of distributed devices, since each device will correspond with an individual.
+The original idea was to give technical coverage for particular needs that professionals in this sector had to face off with.
+At this point it is easy to notice this solution will need to be a distributed one, due to high samples dispersion.
+As we can see, there are even remote techniques to predict this sample density/dispersion using remote methods which predicts between $157$-$170$ individuals per hectare\cite{ForestStandVol} (depending on the model used).
+
+So according to these sample size determination theories, a big size for samples and the need for a big wireless network of distributed devices is essential for a great resolution, since each device will correspond with an individual.
 
 \subsection{Internet of Things}
-This previously exposed is, more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier concept called M2M (Machine-to-Machine) communications.
-However, also according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept yet, but according to \cite[p.~2(920)]{IoTObjetives}
+This big wireless network of distributed devices, more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier one called M2M (Machine-to-Machine) communications.
+Although according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept yet, \cite[p.~2(920)]{IoTObjetives} defined it as:
 
@@
 
-This, at least, covers a little part what this project is intended to do: \enquote{Interconnect ordinary physical objects with the identifiable adresses} to provide intelligent services.
+This, at least, covers a small part of what this project is intended to do: \enquote{Interconnect ordinary physical objects with the identifiable addresses} to provide intelligent services.
 These physical objects are in this case ordinary dendrometers.
 
-Over the years there have existed analog and handled by hand dendrometers, thus data acquisition had to follow a manual process in the same way.
-This could turn out bothering because of the big size for this statistical population, as previously stated.
-So it was traditionally necessary to go there and as part of the field work, take individual by individual the whole sample data.
+Over the years there have been analog and handled by hand dendrometers, thus data acquisition had to follow a manual process in the same way.
+This could become a time-intensive task because of the big size for this statistical population, as previously stated.
+Traditionally, one would need to go there and as part of the field work, take the whole sample data individual by individual.
 
@@
 
-This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point could be interesting to extend this project to include also a sensor to heights measurement; however, this project is already considered extensive enough.
-
-A lot of comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems, consequently as stated, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, because those professional systems are still expensive.
-So this is intended to get a cheaper system and make it accessible to everyone who wants to monitor one or more trees growth.
+This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point extending the project could be interesting to include also a heights measurement sensor; however, this could be addressed in future projects.
+
+Many comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems.
+Consequently as stated, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system; since those professional systems are still expensive.
+This research is intended to get a cheaper system and make it accessible to everyone who wants to monitor the growth of one or more trees.
 
 There are quite a few types of dendrometers but according to \cite{DendroResearch} \enquote{It is possible to define two broad categories of dendrometer: those that contact the stem and those that do not}.
-This project is focused on the former kind, so for this project is being developed a \enquote{contact dendrometer} based on a linear potentiometer.
+This project is focused on the former kind, so we are developing a \enquote{contact dendrometer} based on a linear potentiometer.
 
 \subsection{Objectives}
-As previously stated the Abstract section, the most important objective for this project is to research about the possibility of make this dendrometry system cheaper and more scalable and accessible.
+As previously stated in the Abstract, the most important objective for this project is to research about the possibility to make this dendrometry system cheaper, more scalable and accessible.
 
 Forests are remote and large surface areas which will require long-range communication technologies to be controlled.
-That's why LoRa is the perfect wireless technology to achieve this and these kind of automatic systems seem to be the proper solution to this hard and slow process.
+That's why LoRa is the perfect wireless technology to achieve this, and these kind of automatic systems seem to be the proper solution to this hard and slow process.
 Hence IoT has much to contribute to this task.
-Forestry professionals could take advantage of this technologies to make this in an easier, cheaper and automated way; being this the main purpose of this project.
+Forestry professionals could take advantage of these technologies to make this in an easier, cheaper and automated way.
+Hence this is the main purpose of this project.
 
@@
 \section{System Architecture}
-A simple block diagram is shown in figure \ref{fig:block_diagram}, this diagram intends to emphasize the difference between two big layers, a physical layer that should be placed ---IoT dendrometers at least--- close to the sample, and another one called abstract layer which can be deployed locally ---TTN stack is a cloud server, although.
+A simple block diagram is shown in figure \ref{fig:block_diagram}, this diagram intends to emphasize the difference between two big layers, a physical layer that should be placed ---IoT dendrometers at least--- close to the sample, and another one called abstract layer which can be deployed locally ---TTN stack is a cloud server, though.
 
@@
 The following subsections will detail each of one.
-As can be seen in \ref{fig:block_diagram}, the physical layer is made up of two main modules, the first one is an Arduino microcontroller with its LoRa interface and a linear potentiometer acting as a sensor ---this whole set is the IoT dendrometer--- and the second one; a Raspberry Pi 3b+ with its LoRa interface to act as a gateway.
+As we can see in figure \ref{fig:block_diagram}, the physical layer is made up of two main modules: the first one is an Arduino microcontroller with its LoRa interface and a linear potentiometer acting as a sensor ---this whole set is the IoT dendrometer--- and a second one consisting of a Raspberry Pi 3b+ with its LoRa interface to act as a gateway.
 
@@
 \subsubsection{Arduino, multipurpose microcontroller}
-It's not difficult to justify the use of such an interesting platform as Arduino.
+Justifying the use of such an interesting platform as Arduino is not difficult.
 The adaptability is one of its strengths, therefore; it is able to acquire and process certain data coming from a set of sensors and manage it to send this via any plugged wireless network interface.
 
-Due circumstances exposed in introduction section is needed an accessible and multipurpose platform to be the basis for the device design itself, this is to say the core part of the dendrometer is an Arduino microcontroller.
+Due to circumstances above, an accessible and multipurpose platform to be the basis for the device design itself is needed; i.e. the core part of the dendrometer is an Arduino microcontroller.
 
@@
 \begin{itemize}
-  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself due is directly in contact with the stem.
+  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself due it is directly in contact with the stem.
 In order to improve data acquisition it will be necessary to use a High Input Impedance Amplifier.
-  \item Arduino microcontroller: this is the core part for the device, it will be responsible for acquire linear potentiometer data and send it to a gateway through LoRa interface. 
+  \item Arduino microcontroller: this is the core part for the device, it will be responsible to acquire linear potentiometer data and send it to a gateway through LoRa interface. 
   \item LoRa interface: similar to other existing wireless interfaces, it is necessary to forward the sensor data to a concentrator (gateway).
-Usually and due its complexity these kind of interfaces are integrated circuits which are mounted on a PCB in order to obtain a pluggable card/shield.
+Usually and due to its complexity these kind of interfaces are integrated circuits which are mounted on a PCB in order to obtain a pluggable card/shield.
 \end{itemize}
@@
 
-By other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
-The former one (the physical layer), is a proprietary technology developed by Semtech.
-So this does mean this layer is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer,  It is intended to sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite[p.~9]{LoRaGeneral}.
+In the other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
+The physical layer, is a proprietary technology developed by Semtech.
+So it is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer, it is intended for sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite[p.~9]{LoRaGeneral}.
 This particularly means that LoRaWAN protocol is perfect for the purpose of this project.
@@
 \subsubsection{Raspberry Pi, a powerful microcomputer}
-As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware point of view, Raspberry Pi implements a better one than Arduino, this also has an impact on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's why this devices are usually called \textit{single-board computers}.
+As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware perspective, Raspberry Pi implements a better one than Arduino, this also results on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's why these devices are usually called \textit{single-board computers}.
 
@@
 The role for Raspberry Pi in this project is to act like a gateway, receiving all the data from different nodes.
-So in order to perform this task correctly it is necessary to provide a LoRa interface, which is also embebed (as in Arduino) in a separate pluggable shield/card/hat.
+So in order to perform this task correctly it is necessary to provide a LoRa interface, which is also embedded (as in Arduino) in a separate pluggable shield/card/hat.
 
 \subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
-These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) have not the ability to comunicate through LoRa physical layer, that's why they need a physical interface in order to manage those LoRa packages.
+These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) do not have the ability to comunicate through LoRa physical layer, that's why they need a physical interface in order to manage those LoRa packages.
 Both shields are based on the SX1276/SX1278 transceiver.
-However the Raspberry Pi hat also has a L80 GPS interface (Base on MTK MT3339), meanwhile Arduino has not.
+However the Raspberry Pi hat also has a L80 GPS interface (Base on MTK MT3339), while Arduino does not.
 
@@
 It is basically a network server placed between gateways and applications servers.
-It is necessary because LoRaWAN is a non-IP protocol, but a MAC layer protocol and does need some for of routing and processing messages before they can reach the application side.
+It is necessary because LoRaWAN is a non-IP protocol, but a MAC layer protocol and does need some form of routing and processing messages before they can reach the application side.
 TTN takes care of these steps.
@@
 
-End-node devices broadcast LoRaWAN packets over LoRa radio protocol to the gateways, this gateways will forward these packets to the TTN backend, where a software module called Router will manage the gateway status and schedule transmissions, these Routers are connected to another software modules called Brokers, that are responsible for map a device to an application ---into TTN environment--- allowing downlinks and uplinks transmissions.
-Once these uplinks transmission are inside TTN environment, the FIWARE LoRaWAN agent will retrieve them, however this is actually the application server shown in figure \ref{fig:TTN_overview}.
+End-node devices broadcast LoRaWAN packets over LoRa radio protocol to the gateways, these gateways will forward these packets to the TTN backend, where a software module called Router will manage the gateway status and schedule transmissions.
+These routers are connected to other software modules called Brokers, that are responsible for mapping a device to an application ---into TTN environment--- allowing downlink and uplink transmissions.
+Once these uplink transmissions are inside TTN environment, the FIWARE LoRaWAN agent will retrieve them, however this is actually the application server shown in figure \ref{fig:TTN_overview}.
 \enquote{The goal of The Things Network is to be very flexible in terms of deployment options.} \cite{TTN_net_arch}.
-As can be read in the docs, TTN infrastructure provides different \enquote{core functions} divided in four main groups:
+As explained in the docs, TTN infrastructure provides different \enquote{core functions} divided in four main groups:
 
 \begin{itemize}
-  \item Gateway-related functions. In order to schedule and manage the gateways.
-  \item Device-related functions. To manage the state and received context data from specific devices.
+  \item Gateway-related functions, in order to schedule and manage the gateways.
+  \item Device-related functions, to manage the state and received context data from specific devices.
   \item Application-related functions, to group different devices used for similar purposes.
@@
 \subsubsection{FIWARE}
-It is defined as \enquote{The Open Source Platform for Our Smart Digital Future}\cite{FIWARE}; in a deeper sense is \enquote{an open source initiative defending an universal set of standards for context data management}\cite{FIWARE}, which basically means that \textit{FIWARE} is an open source platform where develop IoT and smart solutions.
+FIWARE is defined as \enquote{The Open Source Platform for Our Smart Digital Future}\cite{FIWARE}; in a deeper sense it is \enquote{an open source initiative defending an universal set of standards for context data management}\cite{FIWARE}, which basically means that \textit{FIWARE} is an open source platform where IoT and smart solutions can be developed.
 The platform provides a number of software modules called \textit{Generic Enablers}; among all those \textit{Generic Enablers} there is one particularly important called \textit{Orion Context Broker}.
@@
   ]
-  // Prepare upstream data transmission at the next possible time.
+  // Prepare upstream data transmission.
 

### 2020-07-03 docs/LaTex/Paper.tex
Minor redaction fixes

@@
 
-Informa:\\ Que D. Javier Fernández Aparicio ha realizado bajo su dirección el Trabajo de Fin de Grado. Considera que la memoria reúne los requisitos necesarios para su evaluación
+Informa:\\ Que D. Javier Fernández Aparicio ha realizado bajo su dirección el Trabajo de Fin de Grado. Considera que la memoria reúne los requisitos necesarios para su evaluación.
 
@@
 \begin{itemize}
-  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself due it is directly in contact with the stem.
+  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself due to it is directly in contact with the stem.
 In order to improve data acquisition it will be necessary to use a High Input Impedance Amplifier.
@@
 
-By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
+By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as it can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
 As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.
@@
 With regards to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{pleas, refer to that subsection for further details}.
-The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
+The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as it can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
 
@@
 
-So, as can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
+So, as it can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
 
@@
 After a little researching, it is possible to conclude that there is not other operative system which improves the performance of Raspberry Pi OS in a Raspberry Pi.
-Due this former argument this project is going to use Raspberry Pi OS.
+Due to this former argument this project is going to use Raspberry Pi OS.
 
@@
 Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role.
-These usually are generic devices due LoRaWAN protocol versatility.
+These usually are generic devices due to LoRaWAN protocol versatility.
 LoRaWAN is a cloud-based medium access control (MAC) layer protocol, but actually acts as a network layer protocol for managing communication between gateways and nodes, similar to a routing protocol.
@@
 
-Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be the following as stated by \cite{Dragino_DualChannelController_Rpi}
+Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due to it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be the following as stated by \cite{Dragino_DualChannelController_Rpi}
 
@@
 
-As can be seen, this service was modified in order to reach properly the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even set up as suggested with \cmd{pi} user.}
+As it can be seen, this service was modified in order to reach properly the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even set up as suggested with \cmd{pi} user.}
 
@@
 Another important consideration inside TTN platform is to check which decoder is chosen for the received payloads, this can be checked by clicking on \cmd{Payload Formats} tab inside Application Overview.
-As can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so importan to choose the \cmd{CayenneLPP} decoder, because as indicated, \myuline{the payload is being encoded in the end-nodes using the CayenneLPP standard}.
+As it can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so importan to choose the \cmd{CayenneLPP} decoder, because as indicated, \myuline{the payload is being encoded in the end-nodes using the CayenneLPP standard}.
 
@@
 This can be done following the FIWARE API documentation.
-These configuration can be donde using HTTP requests due FIWARE essentially provides a RESTful API \cite{FIWARE_RESTful}, these HTTP requests can be performed using multiple clients like \cmd{curl} or \cmd{Postman} among others.
+These configuration can be donde using HTTP requests due to FIWARE essentially provides a RESTful API \cite{FIWARE_RESTful}, these HTTP requests can be performed using multiple clients like \cmd{curl} or \cmd{Postman} among others.
 This project in fact also provides a \href{https://github.com/WyRe/lora-arduino-dendrometer/tree/master/src/postman}{\cmd{Postman} collection and environment} that can be imported to manage them using a confortable graphical interface.
@@
 \subsubsection{Grafana}\label{sssec:Grafana}
-Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; due it's actually capable to acquire the data from multiple sources.
+Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; due to it is actually capable to acquire the data from multiple sources.
 In this project will be used in order to monitor the context data along the time.
@@
 This view allows the user to create and delete devices ---along with its entities and subscriptions--- nevertheless, the \enquote{Delete device and entities} button remains disabled until a device is chosen in the right list.
-In a similar way than \enquote{Applications Keys}, to add a new device to FIWARE deployed instance by docker, all keys must fullfil a set of minimum requirements; as can be seen in picture \ref{fig:adding_dev}
+In a similar way than \enquote{Applications Keys}, to add a new device to FIWARE deployed instance by docker, all keys must fullfil a set of minimum requirements; as it can be seen in picture \ref{fig:adding_dev}
 
@@
 
-These keys can be obtained from TTN as can be seen in screenshot \ref{fig:dev_overview} ---this screenshot shows info about \cmd{node\_0}, though.
-As can be seen in picture \ref{fig:adding_dev}, there was a problem with \enquote{Device EUI}, this was caused because of the final \cmd{Z} letter in the string which doesn't belong to hexadecimal characters set.
+These keys can be obtained from TTN as it can be seen in screenshot \ref{fig:dev_overview} ---this screenshot shows info about \cmd{node\_0}, though.
+We can see in picture \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because of the final \cmd{Z} letter in the string which doesn't belong to hexadecimal characters set.
 
@@
 
-Finally, as indicated, the \enquote{About} tab shows the versions info as can be seen in \ref{fig:about_tab}.
+Finally, as indicated, the \enquote{About} tab shows the versions info as it can be seen in \ref{fig:about_tab}.
 The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why could be helpful to retrieve these.
@@
 \section{Proof of concept, deployment}
-Due circumstances at early 2020, this section has been reduced, instead of test the system in a real production environment, this section will focus on summarize needed steps/stages to deploy the system properly.
+Due to circumstances at early 2020, this section has been reduced, instead of test the system in a real production environment, this section will focus on summarize needed steps/stages to deploy the system properly.
 Although these steps are already detailed separately, this section tries to put them together to give a consistent and practical point of view.
@@
 Of course, to be able to compile the sketch some dependencies must be satisfied, such as CayenneLPP ArduinoJson and lmic-arduino libraries.
-Moreover, the \textbf{linear potentiometer has been replaced for a rotary potentiometer}, due circumstances again.
+Moreover, the \textbf{linear potentiometer has been replaced for a rotary potentiometer}, again due to circumstances in the moment when the project is developed.
 
@@
 
-Finally the system it would be completely deployed and journaling the context data, as can be seen in figure \ref{fig:final_panel}.
+Finally the system it would be completely deployed and journaling the context data, as it can be seen in figure \ref{fig:final_panel}.
 
@@
 So it would be desirable to replace this HAT by another one, or simple to acquire a proper gateway for a production environment.
-      \item Secondly, Nodes cannot comunicate simultaneously with the gateway because it is listening only one channel, so this is causing a time gap in data acquisition as can be seen in the picture \ref{fig:times_gap}
+      \item Secondly, Nodes cannot comunicate simultaneously with the gateway because it is listening only one channel, so this is causing a time gap in data acquisition as it can be seen in the picture \ref{fig:times_gap}
       \begin{figure}[ht]
@@
     \end{itemize}
-    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in the case that downlinks packets were transmitted, this is why due some security implications frame counters must be reset for ABP devices every time they are shutdown for any circumstance; in particular when downlinks are transmitted, could be necessary to reset the frame counter in the devices, so in this situation would be interesting to consider swap to OTAA method.
+    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in the case that downlinks packets were transmitted, this is why due to some security implications frame counters must be reset for ABP devices every time they are shutdown for any circumstance; in particular when downlinks are transmitted, could be necessary to reset the frame counter in the devices, so in this situation would be interesting to consider swap to OTAA method.
 In fact, TTN offers the possibility to reset the frame counter in TTN console but this could be removed in a future.
-    \item Maybe the developed GUI could have handle also the context data plots, however this project couldn't have covered in anyway this development with the available time, so this is not considered as part of the project because Grafana is complete enough to this task for now and perhaps for ever due it is extremely powerful handling databases.
-    \item For a production environment it seems clear that arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems, this would lead a very compact device which could be also improved adding more sensor due versatility of CayenneLPP encoding standard which allows to send multiple measurements of different sensors.
+    \item Maybe the developed GUI could have handle also the context data plots, however this project couldn't have covered in anyway this development with the available time, so this is not considered as part of the project because Grafana is complete enough to this task for now and perhaps for ever due to it is extremely powerful handling databases.
+    \item For a production environment it seems clear that arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems, this would lead a very compact device which could be also improved adding more sensor due to versatility of CayenneLPP encoding standard which allows to send multiple measurements of different sensors.
     \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; so it likely keeps developing, depending on the needs of the researchers.

### 2020-07-03 docs/LaTex/Paper.tex
Fixes cover page

@@
   UNIVERSIDAD DE EXTREMADURA\\
-  FACULTAD DE CIENCIAS\\[1cm]
+  FACULTAD DE CIENCIAS\\[1.5cm]
   \textbf{%
   	Grado en Física\\
-    TRABAJO DE FIN DE GRADO\\[3cm]
-    \flushleft{\large{Developement of a FIWARE-based application for tree species monitoring (dendrometry)}}\\[3cm]
-    %\flushleft{\large{About implementation of LoRa-based wireless monitoring systems for tree species dendrometry.}}\\[3cm]
-  }
+    TRABAJO DE FIN DE GRADO\\[3.5cm]}
+    {\setstretch{1.25}\flushleft{\large{\textbf{Developement of a FIWARE-based application for tree species monitoring (dendrometry)}}\\[3.5cm]}}
 }

### 2020-07-03 docs/LaTex/Paper.tex
Minor fixes

@@
 
-By other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
+In the other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
 The \cmd{arduino-lmic} library is, as its description in the README.md says, a port from the IBM LoRaWAN MAC in C, but \enquote{slightly modified to run in the Arduino environment allowing using the SX1272, SX1276 transceivers and compatible modules} \cite{MCCI_lmic}.
@@
 
-As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L30}{lines 30 and 31} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
+As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
 This is to enable specific functions in the library for Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must looks like in te example \ref{ej:lmic_project_config.h}
@@
 
-By other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as it can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
+In the other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as it can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
 As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.
@@
 
-So generating a pair of keys and configuring properly the ssh access;\footnote{The reader will probably need to figure out the Raspberry IP in their local network, by other hand, inside a local network will not be necessary specify any port for ssh, though it usually is the 22 by default.} it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
+So generating a pair of keys and configuring properly the ssh access;\footnote{The reader will probably need to figure out the Raspberry IP in their local network, otherwise, inside a local network will not be necessary specify any port for ssh, though it usually is the 22 by default.} it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
 
@@
 
-Once these keys are available in TTN platform, the process can continue.
+We must remember also it is so important to choose the CayenneLPP as the payload decoder. Once these keys are available in TTN platform, the process can continue.
 

### 2020-07-03 docs/LaTex/Paper.tex
Minor fixes

@@
 
-In the other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
+On the other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
 The physical layer, is a proprietary technology developed by Semtech.
@@
 
-In the other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
+On the other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
 The \cmd{arduino-lmic} library is, as its description in the README.md says, a port from the IBM LoRaWAN MAC in C, but \enquote{slightly modified to run in the Arduino environment allowing using the SX1272, SX1276 transceivers and compatible modules} \cite{MCCI_lmic}.
@@
 
-In the other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as it can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
+On the other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as it can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
 As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.

### 2020-07-03 docs/LaTex/Paper.tex
Minor fixes

@@
 \begin{document}
-\pagenumbering{Roman} 
+% \pagenumbering{Roman}  % To use upper case roman numerals in page numbering.
+\pagenumbering{roman}  % To use lower case roman numerals in page numbering.
 \noindent % to prevent the indentation before first minipage and be able to expand minipages
@@
 On the other hand, LoRa can commonly refer to two distinct layers; a physical layer (LoRa itself) and a MAC layer protocol (LoRaWAN).
-The physical layer, is a proprietary technology developed by Semtech.
-So it is not fully open; LoRaWAN, however, is a protocol built to use LoRa physical layer, it is intended for sensor networks, wherein those sensors exchange packages with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite[p.~9]{LoRaGeneral}.
+The physical layer, is a proprietary technology developed by Semtech, so it is not fully open.
+LoRaWAN, however, is a protocol built to use LoRa physical layer, it is intended for sensor networks, wherein those sensors exchange packets with some server with a \enquote{low data rate and relatively long time intervals (one transmission per hour or even days).}\cite[p.~9]{LoRaGeneral}.
 This particularly means that LoRaWAN protocol is perfect for the purpose of this project.
@@
 \subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
-These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) do not have the ability to comunicate through LoRa physical layer, that's why they need a physical interface in order to manage those LoRa packages.
+These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) do not have the ability to comunicate through LoRa physical layer, that's why they need a physical interface in order to manage those LoRa packets.
 Both shields are based on the SX1276/SX1278 transceiver.

### 2020-07-05 docs/LaTex/Paper.tex
Redaction fixes

@@
 
-This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point extending the project could be interesting to include also a heights measurement sensor; however, this could be addressed in future projects.
+This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point extending this project could be interesting to include also a heights measurement sensor; however, this could be an interesting research premise for future projects.
 
 Many comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems.
-Consequently as stated, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system; since those professional systems are still expensive.
+Consequently, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, since those professional systems are still expensive.
 This research is intended to get a cheaper system and make it accessible to everyone who wants to monitor the growth of one or more trees.
@@
 
-Due to circumstances above, an accessible and multipurpose platform to be the basis for the device design itself is needed; i.e. the core part of the dendrometer is an Arduino microcontroller.
+With these constraints, an accessible and multipurpose platform to be the basis for the device design itself is needed; i.e. the core part of the dendrometer is an Arduino microcontroller.
 
@@
 \subsubsection{Raspberry Pi, a powerful microcomputer}
-As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are a lot of Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware perspective, Raspberry Pi implements a better one than Arduino, this also results on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's why these devices are usually called \textit{single-board computers}.
+As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are various Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware perspective, Raspberry Pi implements a better one than Arduino, this also results on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's why these devices are usually called \textit{single-board computers}.
 
@@
 \begin{itemize}
-  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, so this means this hat is designed to be a node-class radio, not a gateway.}
+  \item \textit{LoRa GPS HAT for Raspberry Pi,}\cite{DraginoRpiHat} which makes use of the extended 40-pin GPIO header to be plugged\footnote{It is important to note that this LoRa HAT is not actually designed to play a gateway role, in fact, this is considered a \enquote{Hack where a node-class radio tries to impersonate a gateway}\cite{RpiHatHack}, so this means this hat is designed to be a node-class radio, not a gateway.}
   \item \textit{LoRa Shield for Arduino}\cite{DraginoArdShield}, which is plugged through analog and digital pins.

### 2020-07-05 docs/LaTex/Paper.tex
Redaction fixes

@@
 \subsubsection{Simple Graphical Interface}
-Despite FIWARE is a powerful platform, is not friendly enough regarding final user.
-This is why a the end of this document is described a simple graphical interface developed to manage end-nodes in FIWARE side.
-As will be indicated, this graphical interface has been developed using python and its tkinter module intended to build graphical interfaces.
-
-There are a lot of HTTP clients available, however the reader would be still forced to make manually these HTTP requests, these could result a little complex for a non advanced user, that's why this graphical interface has been developed.
-
-It's not perfect, it has a few deficiencies in fact, but is by the moment an appropriate way to manage devices, entities and subscriptions in locally deployed FIWARE instance.
+Despite FIWARE is a powerful platform, we cannot consider it to be user-friendly.
+This is why at the end of this document ---\ref{sssec:API_GUI} subsection--- a simple graphical interface developed to manage end-nodes in FIWARE side is described.
+As it will be indicated, this graphical interface has been developed using python and its \cmd{tkinter} module intended to build graphical interfaces.
+
+There are a lot of HTTP clients available, however the reader would still be forced to make these HTTP requests manually, and these could add complexity for a non advanced user, hence the need to develop a graphical interface.
+
+It's not perfect, it could be improved in a future, but is by the moment an appropriate way to manage devices, entities and subscriptions in locally deployed FIWARE instance.
 \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{The code} of this graphical interface is available in the project repository at GitHub.
@@
 \section{Design}\thispagestyle{plain}
-Like in the Introduction, the most proper way to present this technical description is to difference between software and hardware parts.
-However, is still interesting to present a simple and general diagram for whole designed solution.
-So as the reader can see in the figure \ref{fig:GenView} is shown the general diagram.
+Like in the Introduction, the most proper way to present this technical description is to differentiate between software and hardware parts.
+However, it is still interesting to present a simple and general diagram for the designed solution.
+This can be seen at figure \ref{fig:GenView}.
 
 Firstly the TTN stack send through MQTT protocol the payload to FIWARE IoTagent-LoRaWAN generic enabler.
-Secondly, FIWARE generic enablers process that payload and store the contained information (also called \textit{context data}) along with a timestamp in the persisting data base (CrateDB), after this, the context data can be visualized in a web view provided by Grafena.
+Secondly, FIWARE generic enablers process that payload and store the contained information (also called \textit{context data}) along with a timestamp in the persisting data base (CrateDB).
+After this, the context data can be visualized in a web view provided by Grafena.
 
 At this moment this solution requires to send manually the HTTP requests in order to communicate with FIWARE generic enablers manually.
-The ideal situation would be to develop a desktop or web application capable of provide any kind of graphical user interface to easily set up the two most important requirements to get the solution working as intended ---Create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
+The ideal situation would be to develop a desktop or web application capable to provide any kind of graphical user interface to easily set up the two most important requirements to get the solution working as intended ---create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
 
@@
   \includegraphics[width=\textwidth]{../schemes/main_scheme_tbg.png}
-  \caption{General view of this solution.}
+  \caption{Solution overview.}
   \label{fig:GenView}
@@
 \subsubsection{Nodes}
-Nodes are the parts of the system in direct contact with trees themselves (one node per each tree).
-Most important features in those devices are: low-cost, low-powered, wireless communication and small size.
-The figure \ref{fig:NodeDiag} shows a diagram of its most relevant parts:
+Nodes are the parts of the system in direct contact with trees themselves (one node per tree).
+Some essential features in those devices are: low-cost, low-powered, wireless communication and small size.
+Figure \ref{fig:NodeDiag} shows a diagram of its most relevant parts:
 
@@
   \item A linear potentiometer, ideally an \href{https://docs.rs-online.com/37bf/0900766b814f0bd0.pdf}{RS Pro Conductive Polymer
-  Potentiometer for Automotive Applications}, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, this project will implement a regular rotary potentiometer.
+  Potentiometer for Automotive Applications}%, however, by circumstances is not possible to use it and in order to perform a kind of proof of concept, this project will implement a regular rotary potentiometer.
   \item A signal conditioning stage, which is needed in order to improve the quality of the voltage signal produced by the potentiometer.
-  \item Arduino + Dragino LoRa Shield; of course it is necessary to schedule data transmission and acquisition, as well as transmit them using the LoRa physical layer and LoRaWAN protocol.
-For this purpose an Arduino-like microcontroller is perfect.
+  \item Arduino + Dragino LoRa Shield; of course it is necessary to schedule data transmission and acquisition, as well as transmitting them using the LoRa physical layer and LoRaWAN protocol.
+An Arduino-like microcontroller is perfect for this purpose.
 \end{itemize}
@@
 \subsubsubsection{Linear Potentiometer}
-As said, ideally this potentiometer should be a linear potentiometer with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it's recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that's why the figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
+As said, ideally this potentiometer should be linear with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it's recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that's why the figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
 
 \subsubsubsection{Signal conditioning}
-This stage is focused on improve the signal acquisition, according to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option.
+This stage is focused on improving the signal acquisition.
+According to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option.
 \doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage.}
 
-It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead a variable resistor, this is why taking advantage of the Arduino power source, particularly the 5 \si{\volt} and GND pins.
+It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead of a variable resistor, this is we use the Arduino power source, particularly the 5 \si{\volt} and GND pins.
 
@@
 In a first place, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino}{the sketch} can be found in \cmd{src/dendro/dendro.ino} path.
-This is the sketch that must be uploaded to arduino microcontroller.
+This is the sketch that must be uploaded to Arduino microcontroller.
 MEGA 2560 in this case.
 
-On the other hand, in order to this sketch takes advantage of the plugged LoRa Shield, is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
+On the other hand, for this sketch to use the plugged LoRa Shield, it is required to base it on the LoRaMAC-in-C library \cite{MCCI_lmic} (\cmd{arduino-lmic})\footnote{Dragino has its own \href{https://github.com/dragino/arduino-lmic}{fork of this library} but has been considered less upgraded than the MCCI version, which is a fork itself from the first port of LMIC library by \href{https://github.com/matthijskooijman}{Matthijs Kooijman}.} implemented by \href{https://mcci.com/}{MCCI}, whose technical documentation can be found in \cite{MCCI_lmic_docs}.
 The \cmd{arduino-lmic} library is, as its description in the README.md says, a port from the IBM LoRaWAN MAC in C, but \enquote{slightly modified to run in the Arduino environment allowing using the SX1272, SX1276 transceivers and compatible modules} \cite{MCCI_lmic}.
@@
 
-As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after install the library with the library manager ---in the Arduino IDE or any other chosen IDE---, the reader will need to go the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
-This is to enable specific functions in the library for Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must looks like in te example \ref{ej:lmic_project_config.h}
+As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after installing the library with the library manager ---in the Arduino IDE or any other chosen IDE, the reader will need go to the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
+This is to enable specific functions in the library for Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must look like in the example \ref{ej:lmic_project_config.h}
 
@@
 After this, the sketch may use specific functions to perform radio transmissions in Europe region.
-This in fact, will has an impact in the \cmd{void setup()\{\}} function ---in the example sketch ttn-abp--- so lines between \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L232}{line 232} and line 254 will be executed when the library is configured for Europe region.
-
-Once the library has been properly configured, the second most important thing is set the pin mapping for the chip, according to the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} in the \cmd{README.md}, \enquote{a variety of configurations are possible}.
+This in fact, will have an impact in the \cmd{void setup()\{\}} function ---in the example sketch ttn-abp--- so lines between \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L232}{line 232} and line 254 will be executed when the library is configured for Europe region.
+
+Once the library has been properly configured, the second most important thing is setting the pin mapping for the chip.
+According to the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} in the \cmd{README.md}, \enquote{a variety of configurations are possible}.
 Then, the most appropriate way to make the pin mapping is following the Dragino indications, so following the figure \ref{fig:Ard_LoRa_Shield_pin_mapping} ---which can also be found in \href{http://wiki.dragino.com/index.php?title=Lora_Shield#Pin_Mapping_and_Unused_Pins}{Dragino documentation}.
@@
 
-With regards to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so are necessary the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address); these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{pleas, refer to that subsection for further details}.
+With regards to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address) are necessary; these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{please, refer to that subsection for further details}.
 The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as it can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
@@
   firstnumber=181,
-  caption={Prpareing the payload.},
+  caption={Preparing the payload.},
   label=ej:prep_payload
@@
 
-So, as it can be seen, in order to prepare the payload firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}, finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
+So, as it can be seen, in order to prepare the payload, firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this, the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}.
+Finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
 
 These \cmd{lpp.*()} functions are part of the CayenneLPP library \cite{CayenneLPP_lib} ---which is also dependent of ArduinoJson library \cite{ArduinoJson_lib}; in particular, the \cmd{lpp.addAnalogInput()} function encodes the raw value as a number encoded in 2 bytes (16 bits signed) with a unit of 0.01.
-In 16 bits signed is possible to represent numbers from -32768 until 32767, so with a unit of 0.01, that amounts to a range from -327.68 to +327.67; that is because the sketch includes the line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L185}{185}, where the raw value \cmd{potVal} is scaled to a [0, 100.0] range.
+In 16 bits signed it is possible to represent numbers from -32768 to 32767, so with a unit of 0.01, that amounts to a range from -327.68 to +327.67; that is because the sketch includes the line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L185}{185}, where the raw value \cmd{potVal} is scaled to a [0, 100.0] range.
 This range will fit in the range that can be encoded in CayenneLPP.
@@
 
-It is also important to note that, as will be stated in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and only can listen in one channel, so despite inside \cmd{void setup()} function the channels are set up as shows the example \ref{ej:chan_setup}, the gateway only will be able to listen the channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
+It is also important to note that, as we will see in \hyperlink{Raspi_HAT}{2.1.2.2 Dragino GPS and LoRa HAT} subsection, the SX1276 is a single channel chip and can only listen in one channel, so despite the channels inside \cmd{void setup()} function being set up as shown in example \ref{ej:chan_setup}, the gateway only will be able to listen to channel 0 (868.1 \si{\mega\hertz}), which is the frequency set in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/global_conf.json}{global\_conf.json} at \cite{Dragino_DualChannelController_Rpi}
 
@@
 
-Because of explained above, the gateway will just receive one package per each $\sim$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening the channel 0, the microcontroller is transmitting at this channel once in 9---
+Because of what's been explained above, the gateway will just receive one package per each $\sim$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening to channel 0, the microcontroller is transmitting at this channel once in 9 times---
 
@@
 This model of Raspberry Pi is able to run almost every GNU/Linux distribution ported to ARM architecture, Raspberry Pi OS, formerly known as Raspbian, \cite{RaspberryPiOs} is in fact a Debian port to ARM architecture.
-After a little researching, it is possible to conclude that there is not other operative system which improves the performance of Raspberry Pi OS in a Raspberry Pi.
+After a little research, it is possible to conclude that there is no other operative system that improves the performance of Raspberry Pi OS in a Raspberry Pi.
 Due to this former argument this project is going to use Raspberry Pi OS.
 
-One of the most interesting features of Raspberry Pi is precisely that OS is installed and run from a microSD card, so the hardware is loading the operative system from this microSD card to the RAM directly, which improves notably the system load times, and increments its portability.
+One of the most interesting features of Raspberry Pi is precisely that the OS is installed and run from a microSD card, so the hardware is loading the operative system from this microSD card to the RAM directly, which improves notably the system load times, and increments its portability.
 
 Raspberry Pi foundation provides the \href{https://www.raspberrypi.org/downloads/noobs/}{\textit{Raspberry Pi Imager}} to perform the installation in a microSD card.\footnote{There are many ways to perform the Raspberry Pi OS installation in a microSD card, this document leaves it to readers to use their preferred method.
-However, this document also considers the \textit{Raspberry Pi Imager} method as the best one.} This tool allows the reader to choose between three different versions of Raspberry Pi OS.
-Recommended, Lite and Full version.
-Lite version is the same than Recommended version but without graphical user interface (GUI or Desktop Environment), meanwhile Full version is the same than Recommended version but with a few extra applications.
-
-This project is using the Recommended version for Raspberry Pi OS, nevertheless, the project does not require absolutely the desktop environment, so to maximize the available free space in the SD card, is better to install the Lite version instead of the Recommended version.
+However, this document also considers the \textit{Raspberry Pi Imager} method as the best one.} This tool allows the reader to choose between three different versions of Raspberry Pi OS: Recommended, Lite and Full version.
+Lite version is the same than Recommended version but without graphical user interface (GUI or Desktop Environment), while Full version includes a few extra applications.
+
+This project is using the Recommended version for Raspberry Pi OS, nevertheless, the project does not require the desktop environment at all, so to maximize the available free space in the SD card, it is better to install the Lite version instead of the Recommended version.
 
@@
 
-One of the first and most important configurations is the internet access; this can be done via two different ways, using the ethernet port (which does not require any extra configuration, just tu plug the cable) or using the WiFi interface, which can be configured using also \cmd{raspi-config} \cite{RaspiWIFIcli}.
+One of the first and most important configurations is the internet access; this can be done via two different ways, using the ethernet port (which does not require any extra configuration, just to plug the cable) or using the WiFi interface, which can be configured using also \cmd{raspi-config} \cite{RaspiWIFIcli}.
 Due to its versatility, there are multiple setup that could fulfil the requirements of this projects:
@@
 \begin{itemize}
-  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} which is using the GPIO connector, so this makes impossible to use it along the LoRa HAT.
+  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} using the GPIO connector, so this makes it impossible to use it along the LoRa HAT.
 Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}.
@@
   \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one exposed, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
-  \item \textbf{GSM module} and battery or mains powered; this document consider this as the ideal setup.
+  \item \textbf{GSM module} and battery or mains powered; this document considers this as the ideal setup.
 This setup doesn't require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone.
@@
       \item \href{https://www.itead.cc/wiki/RPI_SIM800_GSM/GPRS_ADD-ON_V2.0}{Itead Raspberry Pi GSM Board (SIM800)}.
-This is not the most interesting way, because though GPIO pins comes through this does mean the HAT supports \href{http://www.pi-in-the-sky.com/index.php?id=stacking-guide}{stacking}, and even if that does, doesn't mean every other HAT would.
+This is not the most interesting way, because although GPIO pins comes through, this does not mean the HAT supports \href{http://www.pi-in-the-sky.com/index.php?id=stacking-guide}{stacking}, and even if that does, it doesn't mean every other HAT would.
       \item \href{https://tutorials-raspberrypi.com/raspberry-pi-gsm-module-mobile-internet-lte-3g-umts/}{USB GSM module}.
@@
 These kind of GSM modules are in general compliant with GNU/Linux systems and using a regular SIM card, they would be eventually able to provide an internet access point.
-Of course, this also will to increment the size of the whole device.
+Of course, this will also increment the size of the whole device.
     \end{itemize}
@@
 
-So generating a pair of keys and configuring properly the ssh access;\footnote{The reader will probably need to figure out the Raspberry IP in their local network, otherwise, inside a local network will not be necessary specify any port for ssh, though it usually is the 22 by default.} it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
+So, generating a pair of keys and configuring properly the ssh access;\footnote{The reader will probably need to figure out the Raspberry IP in their local network, otherwise, inside a local network it will not be necessary specify any port for ssh, though it is usually the 22 by default.} it will probably be necessary to paste the \cmd{*.pub} key content into \cmd{~/.ssh/authorized\_keys} file (located in the local folder in the Raspberry Pi); and give it the right permissions (\cmd{700}), after this the reader should be able to connect through ssh service.
 
@@
 
-Once is available the remote control for Raspberry Pi through ssh, is possible to manage it completely from the command line, even upgrade the system and compile the required controller to get properly working the LoRa transceiver.
-
-At this point and depending on the type of access point used to provide internet service to the Raspberry, the reader will be able even to remote control the Raspberry over internet.
-For this purpose could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh} ---in the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not.
+Once the remote control for Raspberry Pi through ssh is available, is possible to manage it completely from the command line, even upgrading the system and compiling the required controller to get the LoRa transceiver working properly.
+
+At this point and depending on the type of access point used to provide internet service to the Raspberry, the reader will be able to even remote control the Raspberry over internet.
+For this purpose it could be useful to read the access point documentation in order to perform port forwarding or alternatively \href{http://www.thirdway.ch/En/projects/raspberry_pi_3g/index.php}{a reverse tunnel over ssh} ---in the case the ISP does not allow ssh when using a GSM module, also the reader may find that IPv6 works but IPv4 does not.
 Anyway, as said the internet access point could be variable and it is not determined by this project.\footnote{The most important thing is to work comfortably with the Raspberry Pi with no HDMI, keyboard and mouse plugged, and this can be done locally in a LAN to configure it in a first place.}
@@
 \subsubsubsection{Dragino GPS and LoRa HAT}
-Gateways doesn't need actually a diagram or detailed description because there are multiple devices which could play this role.
+Gateways don't actually need a diagram or detailed description because there are multiple devices which could play this role.
 These usually are generic devices due to LoRaWAN protocol versatility.
@@
 Instead, data transmitted by a node is typically received by multiple gateways.
-Each gateway will forward the received packets from the end-node to de cloud-based network server.
-Besides, this project is using a Raspberry Pi 3B+ with a Dragino Hat which mounts a SX1276 LoRa \textbf{transceiver}\cite{SX1276}; this is so important because means, according to Semtech, this transceiver is not intended to play a gateway role, but a end-node role.
+Each gateway will forward the received packets from the end-node to the cloud-based network server.
+Besides, this project is using a Raspberry Pi 3B+ with a Dragino Hat which mounts a SX1276 LoRa \textbf{transceiver}\cite{SX1276}; this is so important because it means that, according to \textit{Semtech}, this transceiver is not intended to play a gateway role, but a end-node role.
 
@@
 The main reason the SX1276 transceiver is not suitable to work as a gateway is that \textbf{it is actually a single channel transceiver}.
-Despite all of this, there still exist the possibility of use it as a gateway, because it is technically possible.
+Despite all of this, there still exists the possibility of using it as a gateway, because it is technically possible.
 
@@
 
-In order to get the HAT operative, the reader will need to enable the SPI (Serial Peripheral Interface). This can be done also with \cmd{raspi-config} like shows the figure \ref{fig:EnablingSPI}
+In order to get the HAT operative, the reader will need to enable the SPI (Serial Peripheral Interface). This can be done also with \cmd{raspi-config} like shown in the figure \ref{fig:EnablingSPI}
 
@@
 
-After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi} and compile it performing the command shown in example \ref{ej:compiling_dual_channel_pkg} ---Like will be explained below, it is also important to note where \cmd{git} will clone the repository and where the reader will compile it, because the location of the \cmd{dual\_chan\_pkt\_fwd} binary is so critical for that systemd service (also included in the GitHub repository) \cmd{dual\_chan\_pkt\_fwd.service} works properly.
+After this, the reader will be in position to clone the controller from the github repository \cite{Dragino_DualChannelController_Rpi} and compiling it, performing the command shown in example \ref{ej:compiling_dual_channel_pkg} ---As it will be explained below, it is also important to note where \cmd{git} will clone the repository and where the reader will compile it, because the location of the \cmd{dual\_chan\_pkt\_fwd} binary is critical so that the systemd service (also included in the GitHub repository) \cmd{dual\_chan\_pkt\_fwd.service} works properly.
 
@@
 
-Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, for this projects, due to it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, the pins configuration in \cmd{global\_conf.json} must be the following as stated by \cite{Dragino_DualChannelController_Rpi}
+Once the controller is compiled, the reader must check the contents of \cmd{global\_conf.json}, because it is using the LoRa GPS HAT Single Channel LoRa \cite{DraginoRpiHat}, pins setup in \cmd{global\_conf.json} must be the following as stated by \cite{Dragino_DualChannelController_Rpi}
 
@@
 
-I order to obtain the gateway ID the reader will need to perform the following command to run for the first time the \cmd{dual\_chan\_pkt\_fwd} binary, i.e.
-once the compilation process has produced the \cmd{dual\_chan\_pkt\_fwd} binary is necessary to run it directly to check the terminal output and find for the gateway ID.
+In order to obtain the gateway ID the reader will need to perform the following command to run for the first time the \cmd{dual\_chan\_pkt\_fwd} binary, i.e.
+once the compilation process has produced the \cmd{dual\_chan\_pkt\_fwd} binary it is necessary to run it directly to check the terminal output and find for the gateway ID.
 
@@
 
-After to have obtained the gateway ID (\cmd{b8:27:eb:ff:ff:1b:14:9b} for this Dragino HAT), the reader can proceed then to install the system service by perform \cmd{sudo make install}.
+After having obtained the gateway ID (\cmd{b8:27:eb:ff:ff:1b:14:9b} for this Dragino HAT), the reader can proceed then to install the system service by performing \cmd{sudo make install}.
 As the content of \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/Makefile#L22}{\cmd{Makefile}} shows at the line 22, \cmd{sudo make install} will copy the system service \cmd{dual\_chan\_pkt\_fwd.service} to the path \cmd{/lib/systemd/system/} and it will enable it in order to start the service ---the service will start the controller as the reader can see at line 9 in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{\cmd{dual\_chan\_pkt\_fwd.service}}--- at every system startup.\footnote{These kind of services are the way in which \cmd{systemd} manages the daemons running in background.}
 
-The reader must be aware of the path where is located the \cmd{dual\_chan\_pkt\_fwd} binary, this is so important because the service \cmd{dual\_chan\_pkt\_fwd.service} is using the absolute path to run the \cmd{dual\_chan\_pkt\_fwd} binary; So this could result in an error if the binary is not properly located or service is not properly modified.
-
-Once the service is installed and all these point have been checked, the reader will be able to manage the service to
+The reader must be aware of the path where the \cmd{dual\_chan\_pkt\_fwd} binary is located, this is so important because the service \cmd{dual\_chan\_pkt\_fwd.service} is using the absolute path to run the \cmd{dual\_chan\_pkt\_fwd} binary; So this could result in an error if the binary is not properly located or service is not properly modified.
+
+Once the service is installed and all these points have been checked, the reader will be able to manage the service to
 
@@
 
-These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status; in the picture \ref{fig:dual_chan_pkt_fwd-status} it can be seen the service active (running)
+These last two instructions show information about service status and \cmd{dual\_chan\_pkt\_fwd} status; in the picture \ref{fig:dual_chan_pkt_fwd-status} the service can be seen active (running)
 
@@
 \begin{itemize}
-  \item \textbf{IoTagent-LoRaWAN}, which is in charge of forward the traffic from TTN stack to Orion Context Broker.
-  \item \textbf{Orion Context Broker}, this is the \myuline{core generic enabler}, in fact as indicated, this is the only one required component to the solution can be considered as \enquote{powered by FIWARE}.
-Besides this generic enabler requires a \textbf{Mongo database} to store entities and subscriptions among other things.
+  \item \textbf{IoTagent-LoRaWAN}, which is in charge of forwarding the traffic from TTN stack to Orion Context Broker.
+  \item \textbf{Orion Context Broker}, this is the \myuline{core generic enabler}, in fact as indicated, this is the only one required component so the solution can be considered as \enquote{powered by FIWARE}.
+Besides, this generic enabler requires a \textbf{Mongo database} to store entities and subscriptions among other things.
   \item \textbf{QuantumLeap}, in order to make the context data persistent, this generic enabler is needed.
@@
 
-These bold highlighted are the deployed containers that will be detailed later, many of them are not even part of FIWARE platform, though.
+The terms above highlighted in bold are the deployed containers that will be detailed later, many of them are not even part of FIWARE platform, though.
 First this document will expose how to set up the gateway, create an application and add devices inside The Things Network stack.
@@
 \subsubsection{The Things Network Stack}\label{sssec:TTN_Stack}
-This is a platform for the LoRaWAN protocol where is possible to manage gateways, devices and applications.
-It provides a centralized way to manage these kind of things regarding security too, generating several keys to tight the end-devices to the chosen server cloud solution (FIWARE in this case).
-There are three fundamental process the reader must follow to include nodes and gateways in the platform, those are luckily pretty well documented.
+This is a platform for the LoRaWAN protocol where it is possible to manage gateways, devices and applications.
+It provides a centralized way to manage these kind of things regarding security too, generating several keys to tie the end-devices to the chosen server cloud solution (FIWARE in this case).
+There are three fundamental processes the reader must follow to include nodes and gateways in the platform, those are luckily pretty well documented.
 
@@
   \item \textbf{Register a gateway}; this is the process where \cmd{Gateway ID} shown in example \ref{ej:Running_DualChann_Controller} is needed; this process is, as indicated, well documented in \cite{Gateway_reg}.
-  \item \textbf{Add an application}; this application will provide a set of security keys as well as a zone where manage different devices (these are basically the end-nodes i.e.
+  \item \textbf{Add an application}; this application will provide a set of security keys as well as a zone where to manage different devices (these are basically the end-nodes i.e.
 the arduino microcontrollers).
-Is also documented in \cite{Add_application}.
+It is also documented in \cite{Add_application}.
   \item \textbf{Register a device} within added application; these devices will be ultimately the microcontrollers. Again the reader can find the relative documents in \cite{Device_reg}. 
@@
 Another important consideration inside TTN platform is to check which decoder is chosen for the received payloads, this can be checked by clicking on \cmd{Payload Formats} tab inside Application Overview.
-As it can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so importan to choose the \cmd{CayenneLPP} decoder, because as indicated, \myuline{the payload is being encoded in the end-nodes using the CayenneLPP standard}.
+As it can be seen in figure \ref{fig:CayenneLPP_decoder}, it is so important to choose the \cmd{CayenneLPP} decoder, because as indicated, \myuline{the payload is being encoded in the end-nodes using the CayenneLPP standard}.
 
@@
 
-Once this has been done, the microcontrollers has also been programmed and the gateway configured, the reader should be able to read the sensor data in the \cmd{Data} tab like figure \ref{fig:sens_data_TTN}
+Once this has been done, the microcontrollers have also been programmed and the gateway configured, the reader should be able to read the sensor data in the \cmd{Data} tab like figure \ref{fig:sens_data_TTN}
 

### 2020-07-06 docs/LaTex/Paper.tex
Redaction improvements

@@
 % \renewcommand{\maketitlehooka}{\headingfont}
+
+% Titles spacing (I'm not using it because I've almos finished the document when I included it.)
+% \titlespacing*{\section}
+% {0pt}{4ex plus 1.5ex minus 3ex}{3ex minus 2.5ex}
+% \titlespacing*{\subsection}
+% {0pt}{4ex plus 1.5ex minus 3ex}{3ex minus 2.5ex}
+% \titlespacing*{\subsubsection}
+% {0pt}{4ex plus 1.5ex minus 3ex}{3ex minus 2.5ex}
 
@@
 \subsubsection{FIWARE}\label{sssec:FIWARE}
-As shown in figure \ref{fig:GenView}, this solution is being implemented using particularly three FIWARE generic enablers, the most important one, of course, \textbf{Orion Context Broker}, also a LoRaWAN agent called \textbf{IoTagent-LoRaWAN} and finally to make data persistent, \textbf{QuantumLeap}.
-
-The workflow is like picture \ref{fig:GenView} shows, first the IoTagent receives the context data from TTN, this IoTagent sends this context data to the Orion Context Broker and this latter notifies QuantumLeap the variations on the context data, thus QuantumLeap can proceed to store them into Crate database.
+As it is shown in figure \ref{fig:GenView}, this solution is being implemented using particularly three FIWARE generic enablers, the most important one, of course is \textbf{Orion Context Broker}, also a LoRaWAN agent called \textbf{IoTagent-LoRaWAN} and finally to make data persistent, \textbf{QuantumLeap}.
+
+The workflow is like picture \ref{fig:GenView} shows, first the IoTagent receives the context data from TTN, this IoTagent sends this context data to the Orion Context Broker and this latter notifies QuantumLeap any variation on the context data, thus QuantumLeap can proceed to store them into Crate database.
 
@@
 This can be done following the FIWARE API documentation.
-These configuration can be donde using HTTP requests due to FIWARE essentially provides a RESTful API \cite{FIWARE_RESTful}, these HTTP requests can be performed using multiple clients like \cmd{curl} or \cmd{Postman} among others.
+This setup can be donde using HTTP requests because FIWARE essentially provides a RESTful API \cite{FIWARE_RESTful}, these HTTP requests can be performed using multiple clients like \cmd{curl} or \cmd{Postman} among others.
 This project in fact also provides a \href{https://github.com/WyRe/lora-arduino-dendrometer/tree/master/src/postman}{\cmd{Postman} collection and environment} that can be imported to manage them using a confortable graphical interface.
-
-Anyway and after of this little introduction, the most important thing is to deploy all FIWARE generic enablers and its dependencies in a secure and stable way.
-This is why this project uses \cmd{docker} and \cmd{docker-compose} in order to ensure these modules deployment and execution.
-Again this project provides the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/fiware/docker-compose.yml}{\cmd{docker-compose.yml}} file to ensure a correct deployment.
+Also a graphical user interface has been developed to interact with FIWARE, it is described in \ref{sssec:API_GUI} subsection.
+
+Anyway and after this little introduction, the most important thing is to deploy all FIWARE generic enablers and its dependencies in a secure and stable way.
+This is why this project uses \cmd{docker} and \cmd{docker-compose} in order to ensure the deployment and execution of these modules.
+Again this project provides the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/fiware/docker-compose.yml}{\cmd{docker-compose.yml}} file to ease this deployment.
 
@@
 
-So cloning the project and navigating to \cmd{src/fiware/} folder in the terminal the reader will must perform \cmd{docker-compose up -d} to deploy all containers and keep detached from the logs output.
+So cloning the project and navigating to \cmd{src/fiware/} folder in the terminal, the reader will must perform \cmd{docker-compose up -d} to deploy all containers and keep detached from the logs output.
 Once this has been done, performing \cmd{docker ps -a} will show the status of deployed containers, as figure \ref{fig:docker_containers} shows.
-In a similar way, to check the logs the reader could perform \cmd{docker-compose logs -f <NAME>} where the argument \cmd{<NAME>} is optional and when is not provided \cmd{docker-compose} will show logs for all containers; this is also shown in picture \ref{fig:docker_containers}.
+In a similar way, to check these logs the reader could perform \cmd{docker-compose logs -f <NAME>} where the argument \cmd{<NAME>} is optional and when is not provided \cmd{docker-compose} will show these logs for all containers; this is also shown in picture \ref{fig:docker_containers}.
 
@@
 \begin{itemize}
-  \item In a first place the reader will need to provision a new device for the \textbf{IoTagent-LoRaWAN}, this can be done performing a \cmd{POST} HTTP request whose body will include the necessary data to retrieve the TTN stack context data.
+  \item In a first place the reader will need to provision a new device for the \textbf{IoTagent-LoRaWAN}, this can be done performing a \cmd{POST} HTTP request whose body must include the necessary data to retrieve the TTN stack context data.
 
@@
 
-  As example \ref{ej:provisioning_devices} shows, is mandatory to provide a \cmd{"device\_id"}, which would be desirable to match with de TTN device ID, an \cmd{"entity\_name"} in order to make Orion Context Broker create that entity and also an \cmd{"username"} and \cmd{"application\_id"} that must match with the application ID in TTN.
-Besides It will be needed the \cmd{"password"}, which can be obtained from the Application Overview in TTN website and the \cmd{"dev\_eui"}, \cmd{"app\_eui"} and \cmd{"application\_key"}, all of these ---relative to the created device--- are available into TTN dashboard.
-This \cmd{POST} request is included in the \cmd{Postman} collection provided by this project repository; of course \myuline{the reader will need to replace the keys in the environment for their own}.
-
-  When a new device is created, \textbf{IoTagent-LoRaWAN} causes that \textbf{Orion Context Broker} creates a new entity. In fact this can be checked by performing a \cmd{GET} HTTP request,
+  As example \ref{ej:provisioning_devices} shows, is mandatory to provide a \cmd{"device\_id"}, which must match the TTN device ID, an \cmd{"entity\_name"}\footnote{\cmd{LORA-N-0} is just a FIWARE internal identifier for the entity, so it can be chosen by the reader.
+In fact, When nodes, entities and subscriptions are created using the developed graphical user interface described in \ref{sssec:API_GUI} subsection, the \cmd{"entity\_name"} is the same than \cmd{"device\_id"}.} that will be used by Orion Context Broker to create the entity and also an \cmd{"username"} and \cmd{"application\_id"} that must match the application ID in TTN.
+Besides It will be needed the \cmd{"password"}, which can be obtained from the Application Overview in TTN website and the \cmd{"dev\_eui"}, \cmd{"app\_eui"} and \cmd{"application\_key"}, all of these ---related with the created device--- are available into TTN dashboard.
+This \cmd{POST} request is included in the \cmd{Postman} collection stored in this project repository; of course \myuline{the reader will need to replace the keys in the environment with their own}.
+
+  When a new device is created, \textbf{IoTagent-LoRaWAN} causes that \textbf{Orion Context Broker} creates a new entity. In fact this can be checked performing a \cmd{GET} HTTP request,
 
@@
 
-  \item Secondly, after have created the device and therefore the entity, the next thing to do is to create a subscription in order to Orion Context Broker notifies somewhere ---QuantumLeap in this case.
-In a similar way, to create this subscription the reader can perform a \cmd{POST} request against Orion Context Broker, so as example \ref{ej:OrionCB_subscription} shows,
+  \item Secondly, after having created the device and therefore the entity, the next thing to do is to create a subscription in order to Orion Context Broker notifies somewhere ---QuantumLeap in this case.
+In a similar way, to create this subscription the reader must perform a \cmd{POST} request against Orion Context Broker, as it is shown in example \ref{ej:OrionCB_subscription},
 
@@
   As the reader can see, for this subscription has been set up the \cmd{"url": "http://quantumleap:8668/v2/notify"} i.e.
-QuantumLeap will be notified at any change in the entity with \cmd{"id": "LORA-N-0"}; also it's a good practice to check that all went as expected, so sending a \cmd{GET} request the response should be like example \ref{ej:OrionCB_subscriptions_get}
+QuantumLeap will be notified at any change in the entity with \cmd{"id": "LORA-N-0"}.
+It is a good practice to check that all went as expected, so sending a \cmd{GET} request the response should be like example \ref{ej:OrionCB_subscriptions_get}
 
@@
 
-Once these two steps have been performed successfully, the context data should be routed to the Crate database where will be persistent.
+Once these two steps have been successfully performed, the context data should be routed to the Crate database where it will persist.
 
 \subsubsection{Grafana}\label{sssec:Grafana}
-Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; due to it is actually capable to acquire the data from multiple sources.
+Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; because of this, it is capable of acquiring the data from multiple sources.
 In this project will be used in order to monitor the context data along the time.
@@
 The default credentials for the first access are \cmd{username: admin} and \cmd{password: admin}.
-Once the reader has accessed for a first time is possible to modify those credentials or even create new users with different privileges.
+Once the reader has accessed for a first time it is possible to modify those credentials or even create new users with different privileges.
 
@@
 
-AS the central panel in the home view indicates, there are two fundamental task that must be performed to get a detailed view of our context data; these are to add a data source and set up a new dashboard.
+As the central panel in the home view indicates, there are two fundamental task that must be performed to get a detailed view of our context data: add a data source and set up a new dashboard.
 
@@
 
-\enquote{Host} must be \cmd{crate:5432} because is the port which grafana uses to queries under postgres protocol, this can be seen also in figure \ref{fig:docker_containers} another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
+\enquote{Host} must be \cmd{crate:5432} because this is the port which grafana uses to queries under postgres protocol, this can be seen also in figure \ref{fig:docker_containers} another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
 
@@
 
-To get the panel drawing the CrateDB stored context data, the reader must specify a query to retrieve the data from the database.
-This can be done in the \enquote{Query} tab below the graph shown in picture \ref{fig:panel_setup}, here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should looks as follows in example \ref{ej:CrateDB_query}
+To get the panel plotting the context data stored in CrateDB, the reader must specify a query to retrieve the data from the database.
+This can be done in the \enquote{Query} tab below the graph shown in picture \ref{fig:panel_setup}, here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should look as follows in example \ref{ej:CrateDB_query}
 
@@
 The table name \cmd{mtatosioe.etloradevice} can be found inside CrateDB admin panel, which can be accessed in \cmd{http://localhost:4200/\#!/tables} from the web browser.
-So after clicking on \enquote{Apply} button on top right corner the reader should have something similar to figure \ref{fig:final_panel}
+So after clicking on \enquote{Apply} button on top right corner, the reader should have something similar to figure \ref{fig:final_panel}
 
@@
 
-Grafana has a wide set of options intended to monitor different kind of systems, it can represent data in different type of graphics and even allows export to csv format, this can be useful for a more deep analysis using other technologies like, for instance, R, a well-known language widely used for statistical analysis.
-
-Of course, this context data could be acquired directly from CrateDB by means of some custom application developed specifically to this task, however this is beyond of the purpose of this project.
-Maybe in a future a feature like this could be implemented.
+Grafana has a wide options set intended to monitor different kind of systems, it can plot data in variety of graphics and it allows export to csv format even, this can be useful for a more deep analysis using other technologies like, for instance, R, a well-known language widely used for statistical analysis.
+
+Of course, this context data could have been acquired directly from CrateDB through some custom application developed specifically to this task, however this is beyond of the purpose of this project.
+Perhaps in a future, a feature like this could be implemented.
 
 \subsubsection{FIWARE API Graphical User Interface}\label{sssec:API_GUI}
-Finally, to make more friendly the FIWARE interaction exposed in \hyperref[sssec:FIWARE]{FIWARE} subsection, this project also provide a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
-This graphical interface has been developed using the well-known tkinter library in python, it requires also a few python modules that are specified in \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/pip_requirements.txt}{\cmd{/src/api-gui/pip\_requirements.txt}}, in fact the user should be able to install them by performing in a terminal the following command \cmd{pip install -r pip\_requirements.txt}.
-
-When the requirements are satisfied and all docker containers are running properly\footnote{In a first place the GUI will retrieve the versions for Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap, so if they are not running the GUI wont open.} the GUI can be run performing \cmd{./api-gui.py}; this should open the window shown by figure \ref{fig:GUI_app_keys}
+Finally, to make user-friendly the FIWARE interaction seen in \hyperref[sssec:FIWARE]{FIWARE} subsection, this project also provide a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
+This graphical interface has been developed using the \cmd{tkinter} module in python, it requires also a few python modules that are specified in \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/pip_requirements.txt}{\cmd{/src/api-gui/pip\_requirements.txt}}, in fact the user should be able to install them by performing in a terminal the following command: \cmd{pip install -r pip\_requirements.txt}.
+
+When the requirements are satisfied and all docker containers are properly running\footnote{In a first place the GUI will retrieve the versions for Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap, so if they are not running the GUI will not open.} the GUI can be run performing \cmd{./api-gui.py}; this should open the window shown by figure \ref{fig:GUI_app_keys}
 
@@
 \end{itemize}
+
 When any of this requirements are not fulfilled the related label will turn red as figure \ref{fig:GUI_app_keys_fail} shows
@@
 
-This information of course must match exactly with information provided by TTN, actually the \enquote{TTN Application Acess Key} and \enquote{TTN Application EUI} are provided by TTN, however the \enquote{TTN Application ID} must be chosen by the user when a new application is created using the TTN web interface, but this ID must be unique and it must be at least 3 characters long, that is why the developed GUI follows the same criteria.
+This information of course must match the one provided by TTN, actually the \enquote{TTN Application Acess Key} and \enquote{TTN Application EUI} are provided by TTN, however the \enquote{TTN Application ID} must be chosen by the user when a new application is created using the TTN web interface, but this ID must be unique and it must be at least 3 characters long, this is why the developed GUI follows the same criteria.
 
 When the keys are updated properly the labels over the text fields are also updated with the chosen values and the \enquote{Devices Management} tab is enabled.
-This tab provides also a more friendly interface to manage devices.
-In a similar way this info will must match with the info provided by TTN interface, so that as figure \ref{fig:device_management_tab}
+This tab provides a friendly-user interface to manage devices.
+In a similar way, this info will must be the same than info provided by TTN interface, so that as figure \ref{fig:device_management_tab}
 
@@
 \begin{itemize}
-  \item Device entity and ID must be also unique and it has to be at least 3 characters long. Moreover, this ID must match exactly with the ID chosen in TTN interface when the device is created,
+  \item Device entity and ID must be also unique and it has to be at least 3 characters long. Moreover, this ID must match the chosen ID in TTN interface when the device is created,
   \item Application Session Key must be 32 characters long and also an hexadecimal string.
@@
 These keys can be obtained from TTN as it can be seen in screenshot \ref{fig:dev_overview} ---this screenshot shows info about \cmd{node\_0}, though.
-We can see in picture \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because of the final \cmd{Z} letter in the string which doesn't belong to hexadecimal characters set.
+We can see in picture \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because the final \cmd{Z} letter in the string doesn't belong to hexadecimal characters set.
 
@@
 Finally, as indicated, the \enquote{About} tab shows the versions info as it can be seen in \ref{fig:about_tab}.
-The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why could be helpful to retrieve these.
+The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why could be helpful to retrieve them.
 
@@
 
-As figure \ref{fig:GUI_2nodes} shows, once a node is created is added to the list, subscription and entity ID are useful to remove manually just a subscription or an entity by performing a manual \cmd{DELETE} request, in fact have been also included several \cmd{DELETE} requests in the attached Postman collection, however, when a node is deleted using this graphical interface button is also being removed its entity and its subscription.
+As figure \ref{fig:GUI_2nodes} shows, once a node is created is added to the list, subscription and entity ID are useful to remove manually only a subscription or an entity by performing a manual \cmd{DELETE} request, in fact have been also included several \cmd{DELETE} requests in the attached Postman collection, however, when a node is deleted using this graphical interface button is also being removed its entity and its subscription.
 
@@
 \section{Proof of concept, deployment}
-Due to circumstances at early 2020, this section has been reduced, instead of test the system in a real production environment, this section will focus on summarize needed steps/stages to deploy the system properly.
-Although these steps are already detailed separately, this section tries to put them together to give a consistent and practical point of view.
+Due to circumstances at early 2020, this section has been reduced, instead of test the system in a real production environment, this section will focus on summarize needed steps/stages to properly deploy the system.
+Although these steps are already separately detailed, this section tries to put them together to give a consistent and practical point of view.
 
@@
 \subsection{TTN Stack}
-The first thing that must be done is to follow the three steps described in subsection \ref{sssec:TTN_Stack} ---Register a Gateway using its \enquote{Gateway ID}, Add an application and register a device for that application; this is the first step because the reader will need the security keys to two following stages:
+Then, the reader should follow the three steps described in subsection \ref{sssec:TTN_Stack}: Register a Gateway using its \enquote{Gateway ID}, Add an application and register a device for that application.
+This is the first step because the reader will need the security keys to two following stages:
 
 \begin{itemize}
-  \item Embebe these keys in Arduino sketch. This sketch will require the \enquote{Application Session Key}, \enquote{Network Session Key} and \enquote{Device Address}.
+  \item Embed these keys in Arduino sketch. This sketch will require the \enquote{Application Session Key}, \enquote{Network Session Key} and \enquote{Device Address}.
   \item Use them to create devices, and subscriptions in FIWARE.
@@
 
-We must remember also it is so important to choose the CayenneLPP as the payload decoder. Once these keys are available in TTN platform, the process can continue.
+We must remember also it is so important to choose the CayenneLPP as the payload decoder. So whether these keys are available in TTN platform, the process can continue.
 
 \subsection{Customize the Arduino sketch for a specific device}
-As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}, this latter is the most important key, because several nodes chan use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check those are different for every node created in TTN platform.
-
-Of course, to be able to compile the sketch some dependencies must be satisfied, such as CayenneLPP ArduinoJson and lmic-arduino libraries.
+As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}, this latter is the most important one, because several nodes chan use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check that all newly created node has a different one in TTN platform.
+
+Of course, to be able to compile the sketch some dependencies must be satisfied, such as CayenneLPP, ArduinoJson and lmic-arduino libraries.
 Moreover, the \textbf{linear potentiometer has been replaced for a rotary potentiometer}, again due to circumstances in the moment when the project is developed.
@@
 
-It's important to note that has been chosen the analog pin 2 for analog signal acquisition.
-Besides the reader can check the status of the device in serial monitor, however, the reader must also note that serial monitor is setup at 115200 bits per second.
+It is important to note that analog pin 2 has been chosen the for analog signal acquisition.
+Besides the reader can check the status of the device in serial monitor, however, the reader may also note that serial monitor is setup at 115200 bits per second.
 
@@
 
-This will deploy essentially the IoTagent-LoRaWAN, Orion CB, QuantumLeap and Grafana. The status of these containers can be checked by performing \cmd{docker ps -a}.
+This will deploy essentially the IoTagent-LoRaWAN, Orion CB, QuantumLeap and Grafana. The status of these containers can be checked performing \cmd{docker ps -a}.
 
 \subsection{Devices, entities and subscriptions}
-When Docker containers are deployed, it should be possible to send HTTP requests to create a device (with its entity) and a subscription to notify QuantumLeap at any change for a particular entity.
-This process can be don by two different ways as indicated previously,
+When Docker containers are deployed, it should be possible to send HTTP requests to create a device (along its entity) and a subscription to notify QuantumLeap any change in any entity.
+This process can be done through two different ways as indicated previously,
 
 \begin{itemize}
-  \item Using developed graphical user interface (recommended, see \ref{sssec:API_GUI} subsection)
-  \item Using \cmd{curl} or Postman to perform directly the HTTP requests
+  \item Using developed graphical user interface (recommended, see \ref{sssec:API_GUI} subsection).
+  \item Using \cmd{curl} or Postman to perform the HTTP requests directly.
 \end{itemize}
@@
 \subsection{Grafana}
-Once any device (with an entity) and particularly its subscription are properly created, CrateDB should start journaling the context data, therefore next step is to add CrateDB as a data source in Grafana.
-
-So if the data source is properly added, it should be possible to create a new panel into a new dashboard, setting up the proper query for that panel, Grafana should start to plot the context data.
+Once any device (along its entity) and particularly its subscription are properly created, CrateDB should start journaling the context data, therefore next step is to add CrateDB as a data source in Grafana.
+
+So if the data source is properly added, it should be possible to create a new panel into a new dashboard, setting up the proper query for that panel.
+Then, Grafana should start to plot the context data.
 (see \ref{sssec:Grafana} subsection for further details)
@@
       \item In a first place, as indicated in its corresponding section, these kind of gateways are LoRaWAN compatible but not compliant.
-So it would be desirable to replace this HAT by another one, or simple to acquire a proper gateway for a production environment.
-      \item Secondly, Nodes cannot comunicate simultaneously with the gateway because it is listening only one channel, so this is causing a time gap in data acquisition as it can be seen in the picture \ref{fig:times_gap}
+So it would be desirable to replace this HAT by another one, or simply to acquire a proper gateway for a production environment.
+      \item Secondly, Nodes cannot comunicate simultaneously with the gateway because it is listening to only one channel, so this is causing a time gap in data acquisition as it can be seen in the picture \ref{fig:times_gap}
       \begin{figure}[ht]
@@
     \end{itemize}
-    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in the case that downlinks packets were transmitted, this is why due to some security implications frame counters must be reset for ABP devices every time they are shutdown for any circumstance; in particular when downlinks are transmitted, could be necessary to reset the frame counter in the devices, so in this situation would be interesting to consider swap to OTAA method.
+    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in the case that downlink packets were transmitted.
+This is why due to some security implications, frame counters must be reset for ABP devices every time they are shutdown for any circumstance.
+In particular when downlink packets are transmitted, could be necessary to reset the frame counter in the devices, so in this situation it would be interesting to consider swap to OTAA method.
 In fact, TTN offers the possibility to reset the frame counter in TTN console but this could be removed in a future.
-    \item Maybe the developed GUI could have handle also the context data plots, however this project couldn't have covered in anyway this development with the available time, so this is not considered as part of the project because Grafana is complete enough to this task for now and perhaps for ever due to it is extremely powerful handling databases.
-    \item For a production environment it seems clear that arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems, this would lead a very compact device which could be also improved adding more sensor due to versatility of CayenneLPP encoding standard which allows to send multiple measurements of different sensors.
+    \item Maybe the developed GUI could have handle also the context data plots, however this task is left as an objective to be addressed in future researches, so this is not considered as part of the project because Grafana is complete capable to handle this for now and perhaps forever because it is extremely powerful handling databases.
+    \item For a production environment it seems clear that Arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems, this would lead a very compact device which could be also improved adding more sensor due to versatility of CayenneLPP encoding standard which allows to send multiple measurements of different sensors.
     \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; so it likely keeps developing, depending on the needs of the researchers.

### 2020-07-06 docs/LaTex/Paper.tex
Fixes docs redaction

@@
 \subsubsection{FIWARE}\label{sssec:FIWARE}
-As it is shown in figure \ref{fig:GenView}, this solution is being implemented using particularly three FIWARE generic enablers, the most important one, of course is \textbf{Orion Context Broker}, also a LoRaWAN agent called \textbf{IoTagent-LoRaWAN} and finally to make data persistent, \textbf{QuantumLeap}.
+As it is shown in figure \ref{fig:GenView}, this solution is being implemented using particularly three FIWARE generic enablers, the most important one, of course, is the \textbf{Orion Context Broker}, also a LoRaWAN agent called \textbf{IoTagent-LoRaWAN} and finally to make data persistent, \textbf{QuantumLeap}.
 
@@
 This setup can be donde using HTTP requests because FIWARE essentially provides a RESTful API \cite{FIWARE_RESTful}, these HTTP requests can be performed using multiple clients like \cmd{curl} or \cmd{Postman} among others.
-This project in fact also provides a \href{https://github.com/WyRe/lora-arduino-dendrometer/tree/master/src/postman}{\cmd{Postman} collection and environment} that can be imported to manage them using a confortable graphical interface.
+This project in fact also provides a \href{https://github.com/WyRe/lora-arduino-dendrometer/tree/master/src/postman}{\cmd{Postman} collection and environment} that can be imported to manage them using a comfortable graphical interface.
 Also a graphical user interface has been developed to interact with FIWARE, it is described in \ref{sssec:API_GUI} subsection.
@@
 
-So cloning the project and navigating to \cmd{src/fiware/} folder in the terminal, the reader will must perform \cmd{docker-compose up -d} to deploy all containers and keep detached from the logs output.
+So cloning the project and navigating to \cmd{src/fiware/} folder in the terminal, the reader will have to perform \cmd{docker-compose up -d} to deploy all containers and stay detached from the logs output.
 Once this has been done, performing \cmd{docker ps -a} will show the status of deployed containers, as figure \ref{fig:docker_containers} shows.
-In a similar way, to check these logs the reader could perform \cmd{docker-compose logs -f <NAME>} where the argument \cmd{<NAME>} is optional and when is not provided \cmd{docker-compose} will show these logs for all containers; this is also shown in picture \ref{fig:docker_containers}.
+In a similar way, to check these logs the reader could perform \cmd{docker-compose logs -f <NAME>} where the argument \cmd{<NAME>} is optional and when it is not provided, \cmd{docker-compose} will show these logs for all containers; this is also shown in picture \ref{fig:docker_containers}.
 
@@
   As example \ref{ej:provisioning_devices} shows, is mandatory to provide a \cmd{"device\_id"}, which must match the TTN device ID, an \cmd{"entity\_name"}\footnote{\cmd{LORA-N-0} is just a FIWARE internal identifier for the entity, so it can be chosen by the reader.
-In fact, When nodes, entities and subscriptions are created using the developed graphical user interface described in \ref{sssec:API_GUI} subsection, the \cmd{"entity\_name"} is the same than \cmd{"device\_id"}.} that will be used by Orion Context Broker to create the entity and also an \cmd{"username"} and \cmd{"application\_id"} that must match the application ID in TTN.
-Besides It will be needed the \cmd{"password"}, which can be obtained from the Application Overview in TTN website and the \cmd{"dev\_eui"}, \cmd{"app\_eui"} and \cmd{"application\_key"}, all of these ---related with the created device--- are available into TTN dashboard.
+In fact, when nodes, entities and subscriptions are created using the developed graphical user interface described in \ref{sssec:API_GUI} subsection, the \cmd{"entity\_name"} is the same than \cmd{"device\_id"}.} that will be used by Orion Context Broker to create the entity and also an \cmd{"username"} and \cmd{"application\_id"} that must match the application ID in TTN.
+Besides, it will need the \cmd{"password"}, which can be obtained from the Application Overview in TTN website and the \cmd{"dev\_eui"}, \cmd{"app\_eui"} and \cmd{"application\_key"}, all of these ---related with the created device--- are available inside the TTN dashboard.
 This \cmd{POST} request is included in the \cmd{Postman} collection stored in this project repository; of course \myuline{the reader will need to replace the keys in the environment with their own}.
@@
 
-  \item Secondly, after having created the device and therefore the entity, the next thing to do is to create a subscription in order to Orion Context Broker notifies somewhere ---QuantumLeap in this case.
+  \item Secondly, after having created the device and therefore the entity, the next thing to do is to create a subscription so that the Orion Context Broker notifies somewhere ---QuantumLeap in this case.
 In a similar way, to create this subscription the reader must perform a \cmd{POST} request against Orion Context Broker, as it is shown in example \ref{ej:OrionCB_subscription},
@@
 
-  As the reader can see, for this subscription has been set up the \cmd{"url": "http://quantumleap:8668/v2/notify"} i.e.
+  As we can see, for this subscription the \cmd{"url": "http://quantumleap:8668/v2/notify"} has been set up, i.e.
 QuantumLeap will be notified at any change in the entity with \cmd{"id": "LORA-N-0"}.
@@
 \subsubsection{Grafana}\label{sssec:Grafana}
-Grafana is a powerful dashboard used at many scenarios in the software industry, it allows to monitor almost everything the reader can imagine; because of this, it is capable of acquiring the data from multiple sources.
-In this project will be used in order to monitor the context data along the time.
+Grafana is a powerful dashboard used at many scenarios in the software industry, it allows monitoring almost everything the reader can imagine; because of this, it is capable of acquiring the data from multiple sources.
+In this project it will be used in order to monitor the context data along the time.
 
@@
 
-As the central panel in the home view indicates, there are two fundamental task that must be performed to get a detailed view of our context data: add a data source and set up a new dashboard.
-
-To add a new data source the reader can use the left sidebar or click where says \enquote{Add your first dada source} in the home view.
-CrateDB is PostgreSQL compatible, so it must be chosen a PostgreSQL database and fill the form with the data shown in figure \ref{fig:Grafana_add_datasource}
+As the central panel in the home view indicates, there are two fundamental tasks that must be performed to get a detailed view of our context data: adding a data source and setting up a new dashboard.
+
+To add a new data source the reader can use the left sidebar or click on \enquote{Add your first dada source} in the home view.
+CrateDB is PostgreSQL compatible, so a PostgreSQL database must be chosen, and the form filled with the data shown in figure \ref{fig:Grafana_add_datasource}
 
@@
 
-\enquote{Host} must be \cmd{crate:5432} because this is the port which grafana uses to queries under postgres protocol, this can be seen also in figure \ref{fig:docker_containers} another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
+\enquote{Host} must be \cmd{crate:5432} because this is the port which grafana uses to query under postgres protocol, this can be seen also in figure \ref{fig:docker_containers}, another important credential is the \enquote{User} that must be \cmd{crate} with no password because CrateDB does not set anyone by default, \enquote{Database} can be set arbitrarily, and finally the \enquote{SSL} has to be disabled.
 
@@
 To get the panel plotting the context data stored in CrateDB, the reader must specify a query to retrieve the data from the database.
-This can be done in the \enquote{Query} tab below the graph shown in picture \ref{fig:panel_setup}, here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should look as follows in example \ref{ej:CrateDB_query}
+This can be done in the \enquote{Query} tab below the graph shown in picture \ref{fig:panel_setup}: here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should look as follows in example \ref{ej:CrateDB_query}
 
@@
 The table name \cmd{mtatosioe.etloradevice} can be found inside CrateDB admin panel, which can be accessed in \cmd{http://localhost:4200/\#!/tables} from the web browser.
-So after clicking on \enquote{Apply} button on top right corner, the reader should have something similar to figure \ref{fig:final_panel}
+So after clicking on \enquote{Apply} at the top right corner, the reader should have something similar to figure \ref{fig:final_panel}
 
@@
 
-Grafana has a wide options set intended to monitor different kind of systems, it can plot data in variety of graphics and it allows export to csv format even, this can be useful for a more deep analysis using other technologies like, for instance, R, a well-known language widely used for statistical analysis.
-
-Of course, this context data could have been acquired directly from CrateDB through some custom application developed specifically to this task, however this is beyond of the purpose of this project.
+Grafana has a wide options set intended to monitor different kind of systems, it can plot data in a variety of graphics and it even allows exporting to csv format, this can be useful for a more deep analysis using other technologies like, for instance, R, a well-known language widely used for statistical analysis.
+
+Of course, this context data could have been acquired directly from CrateDB through some custom application developed specifically for this task, however this is beyond the purpose of this project.
 Perhaps in a future, a feature like this could be implemented.
@@
 \subsubsection{FIWARE API Graphical User Interface}\label{sssec:API_GUI}
-Finally, to make user-friendly the FIWARE interaction seen in \hyperref[sssec:FIWARE]{FIWARE} subsection, this project also provide a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
-This graphical interface has been developed using the \cmd{tkinter} module in python, it requires also a few python modules that are specified in \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/pip_requirements.txt}{\cmd{/src/api-gui/pip\_requirements.txt}}, in fact the user should be able to install them by performing in a terminal the following command: \cmd{pip install -r pip\_requirements.txt}.
+Finally, to make the FIWARE interaction seen in \hyperref[sssec:FIWARE]{FIWARE} subsection user-friendly, this project also provides a simple and light \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{graphical user interface}.
+This graphical interface has been developed using the \cmd{tkinter} module in python, it requires also a few python modules that are specified in \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/pip_requirements.txt}{\cmd{/src/api-gui/pip\_requirements.txt}}.
+In fact the user should be able to install them by performing the following command in a terminal: \cmd{pip install -r pip\_requirements.txt}.
 
@@
 
-To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without complete all the required info won't be enabled the \enquote{Devices Management} tab.
+To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without completing all the required info the \enquote{Devices Management} tab won't be enabled.
 Actually this tab performs a series of checks regarding TTN keys.
@@
   \item TTN Application Access Key must starts with \cmd{ttn-account} string.
-  \item TTN Application EUI must be 16 characters long and be hexadecimal. 
+  \item TTN Application EUI must be 16 characters long and hexadecimal. 
 \end{itemize}
 
-When any of this requirements are not fulfilled the related label will turn red as figure \ref{fig:GUI_app_keys_fail} shows
+When any of these requirements are not fulfilled the related label will turn red as figure \ref{fig:GUI_app_keys_fail} shows
 
@@
 
-This information of course must match the one provided by TTN, actually the \enquote{TTN Application Acess Key} and \enquote{TTN Application EUI} are provided by TTN, however the \enquote{TTN Application ID} must be chosen by the user when a new application is created using the TTN web interface, but this ID must be unique and it must be at least 3 characters long, this is why the developed GUI follows the same criteria.
+This information of course must match the one provided by TTN.
+Actually the \enquote{TTN Application Acess Key} and \enquote{TTN Application EUI} are provided by TTN, however the \enquote{TTN Application ID} must be chosen by the user when a new application is created using the TTN web interface, but this ID must be unique and it must be at least 3 characters long.
+This is why the developed GUI follows the same criteria.
 
 When the keys are updated properly the labels over the text fields are also updated with the chosen values and the \enquote{Devices Management} tab is enabled.
-This tab provides a friendly-user interface to manage devices.
-In a similar way, this info will must be the same than info provided by TTN interface, so that as figure \ref{fig:device_management_tab}
+This tab provides a user-friendly interface to manage devices.
+In a similar way, this info will must be the same than the info provided by TTN interface, so as in figure \ref{fig:device_management_tab}
 
@@
 This view allows the user to create and delete devices ---along with its entities and subscriptions--- nevertheless, the \enquote{Delete device and entities} button remains disabled until a device is chosen in the right list.
-In a similar way than \enquote{Applications Keys}, to add a new device to FIWARE deployed instance by docker, all keys must fullfil a set of minimum requirements; as it can be seen in picture \ref{fig:adding_dev}
+In a similar way than \enquote{Applications Keys}, to add a new device to the FIWARE instance deployed by docker, all keys must fullfil a set of minimum requirements; as it can be seen in picture \ref{fig:adding_dev}
 
 \begin{itemize}
-  \item Device entity and ID must be also unique and it has to be at least 3 characters long. Moreover, this ID must match the chosen ID in TTN interface when the device is created,
+  \item Device entity and ID must be also unique and it has to be at least 3 characters long. In fact, this ID must match the chosen ID in TTN interface when the device is created,
   \item Application Session Key must be 32 characters long and also an hexadecimal string.
@@
 Finally, as indicated, the \enquote{About} tab shows the versions info as it can be seen in \ref{fig:about_tab}.
-The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why could be helpful to retrieve them.
+The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why retrieving them could be helpful.
 
@@
 
-As figure \ref{fig:GUI_2nodes} shows, once a node is created is added to the list, subscription and entity ID are useful to remove manually only a subscription or an entity by performing a manual \cmd{DELETE} request, in fact have been also included several \cmd{DELETE} requests in the attached Postman collection, however, when a node is deleted using this graphical interface button is also being removed its entity and its subscription.
+As figure \ref{fig:GUI_2nodes} shows, once a node is created it is added to the list, subscription and entity IDs are useful to remove manually only a subscription or an entity ---by performing a manual \cmd{DELETE} request, in fact several \cmd{DELETE} requests have been also included in the attached Postman collection, however, when a node is deleted using this graphical interface button its entity and its subscription are also being removed.
 
@@
 \section{Proof of concept, deployment}
-Due to circumstances at early 2020, this section has been reduced, instead of test the system in a real production environment, this section will focus on summarize needed steps/stages to properly deploy the system.
+Due to inconvenient circumstances at early 2020, this section is not as long as initially planted, instead of testing the system in a real production environment, this section will focus on summarizing needed steps/stages to properly deploy the system.
 Although these steps are already separately detailed, this section tries to put them together to give a consistent and practical point of view.
@@
 \subsection{TTN Stack}
-Then, the reader should follow the three steps described in subsection \ref{sssec:TTN_Stack}: Register a Gateway using its \enquote{Gateway ID}, Add an application and register a device for that application.
-This is the first step because the reader will need the security keys to two following stages:
+Then, the reader should follow the three steps described in subsection \ref{sssec:TTN_Stack}: register a Gateway using its \enquote{Gateway ID}, add an application and register a device for that application.
+This is the first step because the reader will need the security keys for the two following stages:
 
@@
 
-We must remember also it is so important to choose the CayenneLPP as the payload decoder. So whether these keys are available in TTN platform, the process can continue.
+We must remember also that it is very important to choose the CayenneLPP as the payload decoder. So if these keys are available in TTN platform, the process can continue.
 
 \subsection{Customize the Arduino sketch for a specific device}
-As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}, this latter is the most important one, because several nodes chan use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check that all newly created node has a different one in TTN platform.
+As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}, this latter is the most important one, because several nodes can use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check that every newly created node has a different one in TTN platform.
 
@@
 It is important to note that analog pin 2 has been chosen the for analog signal acquisition.
-Besides the reader can check the status of the device in serial monitor, however, the reader may also note that serial monitor is setup at 115200 bits per second.
+Besides, the reader can check the status of the device in the serial monitor, however, the reader may also note that serial monitor is setup at 115200 bits per second.
 
 \subsection{Deploy Docker containers}
-Once the hardware has been properly setup, the next step is deploy locally all Docker containers, so as indicated in \ref{sssec:FIWARE} subsection, using \cmd{docker-compose} is possible to do this easily.
+Once the hardware has been properly setup, the next step is to deploy locally all Docker containers, as indicated in \ref{sssec:FIWARE} subsection.
+Using \cmd{docker-compose} this becomes an easy task.
 
@@
 So if the data source is properly added, it should be possible to create a new panel into a new dashboard, setting up the proper query for that panel.
-Then, Grafana should start to plot the context data.
-(see \ref{sssec:Grafana} subsection for further details)
-
-Finally the system it would be completely deployed and journaling the context data, as it can be seen in figure \ref{fig:final_panel}.
+Then, Grafana should start to plot the context data (see \ref{sssec:Grafana} subsection for further details)
+
+Finally the system would be completely deployed and journaling the context data, as it can be seen in figure \ref{fig:final_panel}.
 
@@
 \begin{itemize}
-  \item A rotary potentiometer does work well enough as a proof of concept, but would have been more appropriate to have the expected linear potentiometer to try its sensibility, which is a good point in order to measure such little variations as produced in the stem diameter in a tree.
+  \item A rotary potentiometer does work well enough as a proof of concept, but it would have been more appropriate to have the expected linear potentiometer to try its sensitivity, which is a good point in order to measure such little variations as those produced in the stem diameter in a tree.
   \item The fact that the Raspberry Pi HAT uses a single channel chip (SX1276) has two important implications:
@@
 So it would be desirable to replace this HAT by another one, or simply to acquire a proper gateway for a production environment.
-      \item Secondly, Nodes cannot comunicate simultaneously with the gateway because it is listening to only one channel, so this is causing a time gap in data acquisition as it can be seen in the picture \ref{fig:times_gap}
+      \item Secondly, nodes cannot comunicate simultaneously with the gateway because it is listening to only one channel, so this is causing a time gap in data acquisition as it can be seen in picture \ref{fig:times_gap}
       \begin{figure}[ht]
@@
         \includegraphics[width=.9\textwidth]{../pictures/Grafana_times_gap.png}
-        \caption{Little gap between times caused by the single channel chip.}
+        \caption{Small gap between times caused by the single channel chip.}
         \label{fig:times_gap}
@@
     \end{itemize}
-    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in the case that downlink packets were transmitted.
+    \item End-nodes are being activated using ABP method, this method could not be as effective as OTAA method in case that downlink packets were transmitted.
 This is why due to some security implications, frame counters must be reset for ABP devices every time they are shutdown for any circumstance.
-In particular when downlink packets are transmitted, could be necessary to reset the frame counter in the devices, so in this situation it would be interesting to consider swap to OTAA method.
-In fact, TTN offers the possibility to reset the frame counter in TTN console but this could be removed in a future.
-    \item Maybe the developed GUI could have handle also the context data plots, however this task is left as an objective to be addressed in future researches, so this is not considered as part of the project because Grafana is complete capable to handle this for now and perhaps forever because it is extremely powerful handling databases.
-    \item For a production environment it seems clear that Arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems, this would lead a very compact device which could be also improved adding more sensor due to versatility of CayenneLPP encoding standard which allows to send multiple measurements of different sensors.
-    \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; so it likely keeps developing, depending on the needs of the researchers.
+In particular when downlink packets are transmitted, reset the frame counter in the devices could be necessary, so in this situation swapping to OTAA method might be worth considering.
+In fact, TTN offers the possibility to reset the frame counter in TTN console but this could be removed in the future.
+    \item Maybe the developed GUI could have handled also the context data plots, however this task is left as an objective to be addressed in future researches, so this is not considered as part of the project because Grafana is completely capable to handle this for now and perhaps forever because it is extremely powerful handling databases.
+    \item For a production environment it seems clear that Arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems.
+This would lead to a very compact device which could be also improved adding more sensors due to CayenneLPP encoding standard's versatility which allows sending multiple measurements of different sensors.
+    \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; it has potential to be developed further, depending on the needs of the researches.
 \end{itemize}

### 2020-07-06 docs/LaTex/Paper.tex
Fixes a typo

@@
 \section{Proof of concept, deployment}
-Due to inconvenient circumstances at early 2020, this section is not as long as initially planted, instead of testing the system in a real production environment, this section will focus on summarizing needed steps/stages to properly deploy the system.
+Due to inconvenient circumstances at early 2020, this section is not as long as initially planned, instead of testing the system in a real production environment, this section will focus on summarizing needed steps/stages to properly deploy the system.
 Although these steps are already separately detailed, this section tries to put them together to give a consistent and practical point of view.

### 2020-07-08 docs/LaTex/Paper.tex
Redaction fixes

@@
 This document gives a detailed description of this project, which is focused on researching possible low-cost alternatives for wireless dendrometry systems.
-Currently there are a lot of expensive and professional systems in the market, that's why this project is intended to reduce costs and increase the versatility, scalability and accessibility.
-
-In order to reach these objetives the project will be supported with free software technologies such as FIWARE\cite{FIWARE}, The Things Network\cite{TTN} and Python\cite{Python} or free hardware systems such as Arduino\cite{Arduino} and Raspberry Pi\cite{Raspberrypi}.
+Currently there are a lot of expensive and professional systems in the market, that is why this project is intended to reduce costs and increase the versatility, scalability and accessibility.
+
+In order to reach these objetives the project will be supported with free hardware technologies such as Arduino\cite{Arduino} and Raspberry Pi\cite{Raspberrypi} and free software or open source systems such as FIWARE\cite{FIWARE}, The Things Network\cite{TTN} and Python\cite{Python}.
 
@@
 \section{Introduction}\thispagestyle{plain}
-This project arises itself from a direct interaction with professionals inside forestry sector.
-The original idea was to give technical coverage for particular needs that professionals in this sector had to face off with.
+This project arises from a direct interaction with professionals inside forestry sector.
+The original idea was to give technical coverage for particular needs that professionals in this sector had to tackle with.
 At this point it is easy to notice this solution will need to be a distributed one, due to high samples dispersion.
@@
 This big wireless network of distributed devices, more or less, the definition of the IoT (Internet of Things) concept; according to the abstract in \cite{IoTOverview} IoT concept comes from an earlier one called M2M (Machine-to-Machine) communications.
-Although according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for IoT concept yet, \cite[p.~2(920)]{IoTObjetives} defined it as:
+Although according to \cite[p.~1(71)]{IoTOverview} there is not an official definition for the IoT concept yet, \cite[p.~2(920)]{IoTObjetives} defined it as:
 
@@
 
-Over the years there have been analog and handled by hand dendrometers, thus data acquisition had to follow a manual process in the same way.
+Over the years there have been analog and manual dendrometers, thus data acquisition had to follow a manual process in the same way.
 This could become a time-intensive task because of the big size for this statistical population, as previously stated.
@@
 
-This one is a bit wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point extending this project could be interesting to include also a heights measurement sensor; however, this could be an interesting research premise for future projects.
+This one is a rather wide definition because nowadays most dendrometry researches are focused on stem diameter; however, at this point extending this project could be interesting to include also a heights measurement sensor; however, this could be an interesting starting point for future projects.
 
 Many comercial dendrometry systems are available in the market, nevertheless more than single and manual dendrometers those are complex and professional distributed systems.
-Consequently, one of the most important objectives in this project is to research about the possibility to get lower the costs of the whole system, since those professional systems are still expensive.
+Consequently, one of the most important objectives in this project is to research about the possibility to lower the costs of the whole system, since those professional systems are still expensive.
 This research is intended to get a cheaper system and make it accessible to everyone who wants to monitor the growth of one or more trees.
@@
 Forests are remote and large surface areas which will require long-range communication technologies to be controlled.
-That's why LoRa is the perfect wireless technology to achieve this, and these kind of automatic systems seem to be the proper solution to this hard and slow process.
+That is why LoRa is the perfect wireless technology to achieve this, and this kind of automatic systems seems to be the proper solution to this hard and slow process.
 Hence IoT has much to contribute to this task.
 Forestry professionals could take advantage of these technologies to make this in an easier, cheaper and automated way.
-Hence this is the main purpose of this project.
+This is the main purpose of this project.
 
@@
 
-The following subsections will detail each of one.
+The following subsections will detail each one.
 As we can see in figure \ref{fig:block_diagram}, the physical layer is made up of two main modules: the first one is an Arduino microcontroller with its LoRa interface and a linear potentiometer acting as a sensor ---this whole set is the IoT dendrometer--- and a second one consisting of a Raspberry Pi 3b+ with its LoRa interface to act as a gateway.
 
-Meanwhile, the abstract layer is made up of three software modules.
-TTN stack is a cloud server to retrieve LoRa packets from all existent gateways, FIWARE is a local application server to gather all this context data; and finally Grafana to plot the context data.
+On the other hand, the abstract layer is made up of three software modules.
+TTN stack is a cloud server to retrieve LoRa packets from all existing gateways, FIWARE is a local application server to gather all these context data; and finally Grafana to plot the context data.
 \subsection{Physical Layer}
@@
 Justifying the use of such an interesting platform as Arduino is not difficult.
-The adaptability is one of its strengths, therefore; it is able to acquire and process certain data coming from a set of sensors and manage it to send this via any plugged wireless network interface.
+The adaptability is one of its strengths, therefore; it is able to acquire and process certain data coming from a set of sensors and manage it to send them via any plugged wireless network interface.
 
@@
 
-Since the idea is to produce a low-cost device in order to distribute a high number of them, it must be a simple design; that's why it consist only of three parts,
+Since the idea is to produce a low-cost device in order to distribute a high number of them, it must be a simple design; that is why it consist only of three parts,
 
 \begin{itemize}
-  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself due to it is directly in contact with the stem.
+  \item \href{https://es.rs-online.com/web/p/products/0317780/}{Linear potentiometer:} which is the sensor itself since it is directly in contact with the stem.
 In order to improve data acquisition it will be necessary to use a High Input Impedance Amplifier.
-  \item Arduino microcontroller: this is the core part for the device, it will be responsible to acquire linear potentiometer data and send it to a gateway through LoRa interface. 
+  \item Arduino microcontroller: this is the core part for the device, it will be responsible for acquiring the linear potentiometer data and sending it to a gateway through a LoRa interface.
   \item LoRa interface: similar to other existing wireless interfaces, it is necessary to forward the sensor data to a concentrator (gateway).
-Usually and due to its complexity these kind of interfaces are integrated circuits which are mounted on a PCB in order to obtain a pluggable card/shield.
+Usually and due to its complexity this kind of interfaces are integrated circuits which are mounted on a PCB in order to obtain a pluggable card/shield.
 \end{itemize}
@@
 LoRa is a \enquote{long-range, low-power, low-bitrate, wireless telecommunications system}\cite[]{LoRaGeneral}.
-This is why some devices inside IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been pointed.
+This is why some devices inside the IoT paradigm tend to be economical and low-resources devices, in order to get them distributed/scattered, as it has been already pointed.
 So this low availability (of resources) along with their tendency to be distributed/scattered causes the need for a low-power consumption and a long-range telecommunication.
 
-In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), so as claimed by \cite[]{LoRaGeneral}
+In a more general sense, there is a wider concept to include all these kind of technologies which fullfil the IoT communication requirements, this is the \enquote{Low-Power Wide Area Networks} (LPWAN), as claimed by \cite[]{LoRaGeneral}
 
@@
 
-It is important to note that when talking about \enquote{low-power consumption}, in many cases it is actually meaning battery-powered devices.
+It is important to note that when talking about \enquote{low-power consumption}, in many cases it actually means battery-powered devices.
 
@@
 \subsubsection{Raspberry Pi, a powerful microcomputer}
-As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are various Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware perspective, Raspberry Pi implements a better one than Arduino, this also results on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that's why these devices are usually called \textit{single-board computers}.
-
-A few of interesting hardware specs are for instance, a \textit{Broadcom BCM2837B0, Cortex-A53 (ARMv8) 64-bit SoC @ 1.4GHz} as its CPU, \textit{1GB LPDDR2 SDRAM}, \textit{Gigabit Ethernet over USB 2.0 (maximum throughput 300 Mbps)} or even an \textit{Extended 40-pin GPIO header} among others.\cite{RPiSpecs}
-
-The role for Raspberry Pi in this project is to act like a gateway, receiving all the data from different nodes.
+As in the case of Arduino, Raspberry Pi provides a powerful platform, however in the case of a Raspberry Pi this platform is slightly more complex than for an Arduino.\footnote{There are various Arduino models, and maybe some of them could be able to support, for example, multiple kind of Real Time Operative Systems, but it is not the case of this project, wherein the Arduino role is just to act as a core for the nodes.} From the hardware perspective, Raspberry Pi implements a better one than Arduino, this also results on a higher cost; however, this hardware allows a Raspberry Pi to support a whole operating system; that is why these devices are usually called \textit{single-board computers}.
+
+A few interesting hardware specs are for instance, a \textit{Broadcom BCM2837B0, Cortex-A53 (ARMv8) 64-bit SoC @ 1.4GHz} as its CPU, \textit{1GB LPDDR2 SDRAM}, \textit{Gigabit Ethernet over USB 2.0 (maximum throughput 300 Mbps)} or even an \textit{Extended 40-pin GPIO header} among others.\cite{RPiSpecs}
+
+The role of the Raspberry Pi in this project is to act like a gateway, receiving all data from different nodes.
 So in order to perform this task correctly it is necessary to provide a LoRa interface, which is also embedded (as in Arduino) in a separate pluggable shield/card/hat.
@@
 \subsubsection{Dragino shields/hats}\label{sssec:DraginoShields}
-These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) do not have the ability to comunicate through LoRa physical layer, that's why they need a physical interface in order to manage those LoRa packets.
+These shields/expansion cards are necessary because those devices (Arduino and Raspberry Pi) do not have the ability to comunicate through LoRa physical layer, that is why they need a physical interface in order to manage those LoRa packets.
 Both shields are based on the SX1276/SX1278 transceiver.
@@
 
-Moreover, this project is based on FIWARE (described in the next subsection) and the LoRaWAN agent in FIWARE is developed to interact with TTN infrastructure,
+Moreover, this project is based on FIWARE (described in the next subsection) and the LoRaWAN agent in FIWARE is developed to interact with the TTN infrastructure,
 
@@
 \enquote{The goal of The Things Network is to be very flexible in terms of deployment options.} \cite{TTN_net_arch}.
-As explained in the docs, TTN infrastructure provides different \enquote{core functions} divided in four main groups:
+As explained in the documentation, TTN infrastructure provides different \enquote{core functions} divided in four main groups:
 
@@
 
-After this, the traffic is routed to FIWARE side, which is deployed locally and the user has total control over it.
+After this, the traffic is routed to the FIWARE side, which is deployed locally and the user has total control over it.
 
@@
 The platform provides a number of software modules called \textit{Generic Enablers}; among all those \textit{Generic Enablers} there is one particularly important called \textit{Orion Context Broker}.
-In fact, for a solution to be considered as \textit{Powered by FIWARE} it must to use the \textit{Orion Context Broker} at least.
+In fact, for a solution to be considered as \textit{Powered by FIWARE} it must at least use the \textit{Orion Context Broker}.
 
 This \textit{Context data} is just a way to name any kind of data coming from any kind of sensor.
-\textit{Orion Context Broker} is designed to manage this context data through concepts such as \textit{subscriptions} or \textit{entities}, to name but a few; however, all these \textit{Generics Enablers} communicate with each other using the \textit{FIWARE NGSI RESTful API}\cite{NGSI}.\footnote{\textit{RESTful} term comes from the software architectural style \textit{REST}, which stands for \textit{Representational State Transfer}}
+\textit{Orion Context Broker} is designed to manage this context data through concepts such as \textit{subscriptions} or \textit{entities}, to name a few; however, all these \textit{Generics Enablers} communicate with each other using the \textit{FIWARE NGSI RESTful API}\cite{NGSI}.\footnote{\textit{RESTful} term comes from the software architectural style \textit{REST}, which stands for \textit{Representational State Transfer}}
 
@@
 
-It's not perfect, it could be improved in a future, but is by the moment an appropriate way to manage devices, entities and subscriptions in locally deployed FIWARE instance.
+This graphical interface is a simple solution that could be clearly improved in the future, but is by the moment an appropriate way to manage devices, entities and subscriptions in locally deployed FIWARE instance.
 \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/api-gui/api-gui.py}{The code} of this graphical interface is available in the project repository at GitHub.
@@
 \section{Design}\thispagestyle{plain}
-Like in the Introduction, the most proper way to present this technical description is to differentiate between software and hardware parts.
+As in the Introduction, the most proper way to present this technical description is to differentiate between software and hardware parts.
 However, it is still interesting to present a simple and general diagram for the designed solution.
-This can be seen at figure \ref{fig:GenView}.
-
-Firstly the TTN stack send through MQTT protocol the payload to FIWARE IoTagent-LoRaWAN generic enabler.
+This diagram can be seen in figure \ref{fig:GenView}.
+
+Firstly the TTN stack sends through MQTT protocol the payload to the FIWARE IoTagent-LoRaWAN generic enabler.
 Secondly, FIWARE generic enablers process that payload and store the contained information (also called \textit{context data}) along with a timestamp in the persisting data base (CrateDB).
-After this, the context data can be visualized in a web view provided by Grafena.
-
-At this moment this solution requires to send manually the HTTP requests in order to communicate with FIWARE generic enablers manually.
-The ideal situation would be to develop a desktop or web application capable to provide any kind of graphical user interface to easily set up the two most important requirements to get the solution working as intended ---create a new device using the IoTagent-LoRaWAN, create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
+After this, the context data can be visualized in a web view provided by Grafana.
+
+Out of the box, this solution requires to manually send the HTTP requests in order to communicate with FIWARE generic enablers.
+However, the ideal situation would be to develop a desktop or web application capable of providing any kind of graphical user interface to easily set up the two most important requirements to get the solution working as intended: create a new device in the IoTagent-LoRaWAN and create a subscription in Orion Context Broker which sends notifications to QuantumLeap.
+This graphical interface will be described later.
 
@@
 Nodes are the parts of the system in direct contact with trees themselves (one node per tree).
-Some essential features in those devices are: low-cost, low-powered, wireless communication and small size.
-Figure \ref{fig:NodeDiag} shows a diagram of its most relevant parts:
+Some essential features of those devices are: low-cost, low-powered, wireless communication and small size.
+Figure \ref{fig:NodeDiag} shows a diagram of their most relevant parts:
 
@@
   \item A signal conditioning stage, which is needed in order to improve the quality of the voltage signal produced by the potentiometer.
-  \item Arduino + Dragino LoRa Shield; of course it is necessary to schedule data transmission and acquisition, as well as transmitting them using the LoRa physical layer and LoRaWAN protocol.
+  \item Arduino + Dragino LoRa Shield; of course it is necessary to schedule data transmission and acquisition, as well as to transmit them using the LoRa physical layer and LoRaWAN protocol.
 An Arduino-like microcontroller is perfect for this purpose.
@@
 \subsubsubsection{Linear Potentiometer}
-As said, ideally this potentiometer should be linear with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it's recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that's why the figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
+As said, ideally this potentiometer should be linear with $5\ \si{\kilo\ohm}$ of maximum resistance, however, in order to improve its performance, it is recommended to use it as a voltage divider; also it would be convenient to buffer the resulting output with a high impedance amplifier, that is why figure \ref{fig:NodeDiag} includes a signal conditioning stage, which will be described below.
 
@@
 According to this potentiometer specifications, the best way to achieve this objetive is using a high input amplifier, in fact, a single-supply, rail-to-rail operational amplifier is the best option.
-\doubt{Figure \ref{fig:SignalCond} shows a possible schematics for this stage.}
-
-It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead of a variable resistor, this is we use the Arduino power source, particularly the 5 \si{\volt} and GND pins.
+Figure \ref{fig:SignalCond} shows a possible schematics for this stage.
+
+It is also recommended by the manufacturer to use the potentiometer as a voltage divider instead of a variable resistor, this is the reason why we use the Arduino power source, particularly the 5 \si{\volt} and GND pins.
 
@@
 In a first place, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino}{the sketch} can be found in \cmd{src/dendro/dendro.ino} path.
-This is the sketch that must be uploaded to Arduino microcontroller.
-MEGA 2560 in this case.
+This is the sketch that must be uploaded to the Arduino microcontroller.
+A MEGA 2560 in this case.
 
@@
 
-As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after installing the library with the library manager ---in the Arduino IDE or any other chosen IDE, the reader will need go to the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and the line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
-This is to enable specific functions in the library for Europe region (where the radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must look like in the example \ref{ej:lmic_project_config.h}
+As commented \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L29}{lines 29 and 30} point out, the most important thing before start working on the sketch is to configure the \cmd{arduino-lmic} library, so after installing the library with the library manager ---in the Arduino IDE or any other chosen IDE, the reader will need go to the library path and locate the \href{https://github.com/mcci-catena/arduino-lmic/blob/master/project_config/lmic_project_config.h}{\cmd{<lib\_path>/project\_config/lmic\_project\_config.h}} file to ensure the line 2 (\cmd{\#define CFG\_eu868 1}) is \textbf{uncommented} and line 3 (\cmd{\#define CFG\_us915 1}) is \textbf{commented}.
+This is necessary to enable specific functions in the library for the Europe region (where radio regulations are slightly different than in other regions); so this \cmd{lmic\_project\_config.h} must look like as in example \ref{ej:lmic_project_config.h}
 
@@
 
-After this, the sketch may use specific functions to perform radio transmissions in Europe region.
-This in fact, will have an impact in the \cmd{void setup()\{\}} function ---in the example sketch ttn-abp--- so lines between \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L232}{line 232} and line 254 will be executed when the library is configured for Europe region.
+After this, the sketch may use specific functions to perform radio transmissions in the Europe region.
+This in fact, will have an impact in the \cmd{void setup()\{\}} function ---in the example sketch ttn-abp--- so lines between \href{https://github.com/mcci-catena/arduino-lmic/blob/master/examples/ttn-abp/ttn-abp.ino#L232}{line 232} and line 254 will be executed when the library is configured for the Europe region.
 
@@
 According to the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} in the \cmd{README.md}, \enquote{a variety of configurations are possible}.
-Then, the most appropriate way to make the pin mapping is following the Dragino indications, so following the figure \ref{fig:Ard_LoRa_Shield_pin_mapping} ---which can also be found in \href{http://wiki.dragino.com/index.php?title=Lora_Shield#Pin_Mapping_and_Unused_Pins}{Dragino documentation}.
+Then, the most appropriate way to make the pin mapping is following the Dragino indications, so following figure \ref{fig:Ard_LoRa_Shield_pin_mapping} ---which can also be found in the \href{http://wiki.dragino.com/index.php?title=Lora_Shield#Pin_Mapping_and_Unused_Pins}{Dragino documentation}.
 
@@
 Besides, a brand new board will only have loaded the \cmd{R9} resistor, this means the LoRa CS pin will be the \cmd{D10}.
-This leads to set the pin mapping shown in example \ref{ej:pin_mapping}
+This leads to the pin mapping shown in example \ref{ej:pin_mapping}
 
@@
 On the other hand \enquote{Any pins that are not needed should be specified as \cmd{LMIC\_UNUSED\_PIN}} as it can be also read in the \href{https://github.com/mcci-catena/arduino-lmic#pin-mapping}{Pin Mapping section} of the library \cmd{README.md}.
-As the reader can see, the example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.
-
-With regards to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address) are necessary; these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, so \textbf{please, refer to that subsection for further details}.
+As the reader can see, example \ref{ej:pin_mapping} starts at line 76, that is because this is the line where the pin mapping is placed in the \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L76}{written sketch}.
+
+With regards to LoRaWAN and TTN security, as previously indicated, the chosen method is ABP, so the \cmd{NWKSKEY} (Network Session Key), \cmd{APPSKEY} (Application Session Key) and \cmd{DEVADDR} (End-Device Address) are necessary; these constants placed at lines \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L48}{48}, \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L52}{52} and \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L57}{57} must be obtained from TTN console as will be explained in \hyperref[sssec:TTN_Stack]{The Things Network Stack} subsection, \textbf{Please, refer to that subsection for further details}.
 The sketch also sets analog pin \cmd{A2} to acquire the sensor signal as it can be seen in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L70}{70}.% as well as the payload to send through LoRa interface in line \href{https://github.com/WyRe/lora-arduino-dendrometer/blob/master/src/arduino/dendro/dendro.ino#L65}{65}
@@
 
-So, as it can be seen, in order to prepare the payload, firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it makes \enquote{On the Arduino Due} \cite{Ard_double_type}; after this, the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}.
+So, as it can be seen, in order to prepare the payload, firstly the sketch reads the analog pin and stores the read value in \cmd{potVal} variable ---which is a \cmd{double} type and this makes no difference with \cmd{float} type \enquote{On the Uno and other ATMEGA based boards} but it does make \enquote{On the Arduino Due} \cite{Ard_double_type}; after this, the function \cmd{lpp.reset()} is used to reset the CayenneLPP buffer, then the measure \cmd{potVal} is added to the CayenneLPP buffer in the first channel using the function \cmd{lpp.addAnalogInput()}.
 Finally the buffer content is prepared to be transmitted at the next possible time using the function \cmd{LMIC\_setTxData2()}.
@@
 
-Because of what's been explained above, the gateway will just receive one package per each $\sim$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening to channel 0, the microcontroller is transmitting at this channel once in 9 times---
+Because of what has been explained above, the gateway will just receive one package every $\sim$10 minutes ---at line 74 in the sketch the transmission interval is set at 60 seconds, so as there are 9 channels but the gateway is just listening to channel 0, the microcontroller is transmitting at this channel once in 9 times---
 
@@
 
-Of course this parameter can be adjusted, but for the final purpose of this project 60 seconds are even too many, in order to monitor changes in stem diameter of a tree, a more appropriate value would be 3600 seconds (1 hour) or even greater.
+Of course this parameter can be adjusted, but for the final purpose of this project 60 seconds are even too little, in order to monitor changes in the stem diameter of a tree, and a more appropriate value would be 3600 seconds (1 hour) or even greater.
 
@@
 
-Even so, independently of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-config} application \cite{RaspiConf}.
+Even so, regardless of the installed version, the reader will be able to disable the graphical environment using the console based \cmd{raspi-config} application \cite{RaspiConf}.
 
@@
 Probably the WiFi radio is not enough to make it the difference between this setup and the next one suggested.
-  \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one exposed, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
+  \item Simple \textbf{ethernet} and mains power, this setup is similar to the previous one, but omitting the WiFi limitations/configurations, the access point and the Raspberry Pi should be practically both at the same location.
   \item \textbf{GSM module} and battery or mains powered; this document considers this as the ideal setup.
-This setup doesn't require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone.
-This would be the ideal way because it minimizes the resources needed, which is crucial in an environment where those probably won't be available ---in the middle of the forest.
+This setup does not require a conventional access point because the internet access is granted through a GSM module, in a similar way than a mobile phone.
+This would be the ideal way because it minimizes the resources needed, which is crucial in an environment where those probably will not be available ---in the middle of the forest.
 
@@
       \item \href{https://www.itead.cc/wiki/RPI_SIM800_GSM/GPRS_ADD-ON_V2.0}{Itead Raspberry Pi GSM Board (SIM800)}.
-This is not the most interesting way, because although GPIO pins comes through, this does not mean the HAT supports \href{http://www.pi-in-the-sky.com/index.php?id=stacking-guide}{stacking}, and even if that does, it doesn't mean every other HAT would.
+This is not the most interesting way, because although GPIO pins comes through, this does not mean the HAT supports \href{http://www.pi-in-the-sky.com/index.php?id=stacking-guide}{stacking}, and even if that does, it does not mean every other HAT would.
       \item \href{https://tutorials-raspberrypi.com/raspberry-pi-gsm-module-mobile-internet-lte-3g-umts/}{USB GSM module}.
 This is apparently the most promising option.
-These kind of GSM modules are in general compliant with GNU/Linux systems and using a regular SIM card, they would be eventually able to provide an internet access point.
+This kind of GSM modules are in general compliant with GNU/Linux systems and using a regular SIM card, they would be eventually able to provide an internet access point.
 Of course, this will also increment the size of the whole device.
@@
 \subsubsubsection{Dragino GPS and LoRa HAT}
-Gateways don't actually need a diagram or detailed description because there are multiple devices which could play this role.
-These usually are generic devices due to LoRaWAN protocol versatility.
+Gateways do not actually need a diagram or detailed description because there are multiple devices which could play this role.
+These are usually generic devices due to the LoRaWAN protocol versatility.
 LoRaWAN is a cloud-based medium access control (MAC) layer protocol, but actually acts as a network layer protocol for managing communication between gateways and nodes, similar to a routing protocol.
@@
 
-Nevertheless, there are important considerations about the said above, for example, nodes are not actually associated with an specific gateway.
+Nevertheless, there are important considerations about all this, for example, nodes are not actually associated with an specific gateway.
 Instead, data transmitted by a node is typically received by multiple gateways.
@@
 \hypertarget{Raspi_HAT}{Here} is the important point.
-This Dragino HAT for the Raspberry Pi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is LoRaWAN compatible but isn't LoRaWAN compliant}.
+This Dragino HAT for the Raspberry Pi, mounts an SX1276 transceiver, which is known as node-class transceiver, so in conclusion, \textbf{this Dragino LoRa GPS HAT is LoRaWAN compatible but is not LoRaWAN compliant}.
 The main reason the SX1276 transceiver is not suitable to work as a gateway is that \textbf{it is actually a single channel transceiver}.
@@
 
-After having obtained the gateway ID (\cmd{b8:27:eb:ff:ff:1b:14:9b} for this Dragino HAT), the reader can proceed then to install the system service by performing \cmd{sudo make install}.
+After having obtained the gateway ID (\cmd{b8:27:eb:ff:ff:1b:14:9b} for this Dragino HAT), the reader can proceed to install the system service by performing \cmd{sudo make install}.
 As the content of \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/Makefile#L22}{\cmd{Makefile}} shows at the line 22, \cmd{sudo make install} will copy the system service \cmd{dual\_chan\_pkt\_fwd.service} to the path \cmd{/lib/systemd/system/} and it will enable it in order to start the service ---the service will start the controller as the reader can see at line 9 in \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{\cmd{dual\_chan\_pkt\_fwd.service}}--- at every system startup.\footnote{These kind of services are the way in which \cmd{systemd} manages the daemons running in background.}
 
-The reader must be aware of the path where the \cmd{dual\_chan\_pkt\_fwd} binary is located, this is so important because the service \cmd{dual\_chan\_pkt\_fwd.service} is using the absolute path to run the \cmd{dual\_chan\_pkt\_fwd} binary; So this could result in an error if the binary is not properly located or service is not properly modified.
-
-Once the service is installed and all these points have been checked, the reader will be able to manage the service to
+The reader must be aware of the path where the \cmd{dual\_chan\_pkt\_fwd} binary is located, this is so important because the service \cmd{dual\_chan\_pkt\_fwd.service} is using the absolute path to run the \cmd{dual\_chan\_pkt\_fwd} binary.
+So this could result in an error if the binary is not properly located or the service is not properly modified.
+
+Once the service is installed and all these points have been checked, the reader will be able to manage the service to:
 
@@
 
-As it can be seen, this service was modified in order to reach properly the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even set up as suggested with \cmd{pi} user.}
+As it can be seen, this service was modified in order to properly reach the \cmd{dual\_chan\_pkt\_fwd} binary, which is not located in \cmd{/home/pi/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd} (the default location defined for the binary in the \href{https://github.com/dragino/dual_chan_pkt_fwd/blob/master/dual_chan_pkt_fwd.service#L9}{GitHub repository}); mainly because the user is not called \cmd{pi} anymore but \cmd{wyre}, the path needs to be modified to that path containing the binary (\cmd{/home/wyre/TFG/LoRa/dual\_chan\_pkt\_fwd/dual\_chan\_pkt\_fwd})\footnote{This is a personal setup and is completely optional, the reader can choose their own path or even set up as suggested with \cmd{pi} user.}
 
@@
 
-To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without completing all the required info the \enquote{Devices Management} tab won't be enabled.
+To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without completing all the required info the \enquote{Devices Management} tab will not be enabled.
 Actually this tab performs a series of checks regarding TTN keys.
@@
 These keys can be obtained from TTN as it can be seen in screenshot \ref{fig:dev_overview} ---this screenshot shows info about \cmd{node\_0}, though.
-We can see in picture \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because the final \cmd{Z} letter in the string doesn't belong to hexadecimal characters set.
+We can see in picture \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because the final \cmd{Z} letter in the string does not belong to hexadecimal characters set.
 

### 2020-07-08 docs/LaTex/Paper.tex
Final redaction fixes

@@
 \begin{itemize}
-  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} using the GPIO connector, so this makes it impossible to use it along the LoRa HAT.
+  \item \textbf{PoE} (Power-over-Ethernet); this is not actually a suitable option because Raspberry Pi does not support PoE by itself, it needs a \href{https://www.raspberrypi.org/products/poe-hat/}{separate HAT} using the GPIO connector, so this makes it impossible to use it along with the LoRa HAT.
 Apparently there are hacks to reduce this HAT size, but they also use pins in the GPIO connector \cite{RaspiPoEHack}.
@@
 Orion Context Broker works mainly with two elements; entities and subscriptions, however, for any created entity the context data does not persist in Mongo database, instead the context data will be replaced/updated; so creating a subscription which notifies to QuantumLeap, this generic enabler will store the context data in a \textbf{Crate database} to make it persistent.
-Also a \textbf{Redis database} is used for geocoding purposes, however this project will not deal with it ---this is why of this database is not even being included in the figure \ref{fig:GenView}.
+Also a \textbf{Redis database} is used for geocoding purposes, however this project will not deal with it ---this is why of this database is not even being included in figure \ref{fig:GenView}.
   \item \textbf{Grafana}, this container is intended to provide a web interface to visualize and manage the retrieved context data.
@@
 The terms above highlighted in bold are the deployed containers that will be detailed later, many of them are not even part of FIWARE platform, though.
-First this document will expose how to set up the gateway, create an application and add devices inside The Things Network stack.
+First, this document will show how to set up the gateway, create an application and add devices inside The Things Network stack.
 
@@
 \subsubsection{Grafana}\label{sssec:Grafana}
-Grafana is a powerful dashboard used at many scenarios in the software industry, it allows monitoring almost everything the reader can imagine; because of this, it is capable of acquiring the data from multiple sources.
+Grafana is a powerful dashboard used at many scenarios in the software industry, it allows monitoring almost everything the reader can imagine; because of this, it is capable of acquiring data from multiple sources.
 In this project it will be used in order to monitor the context data along the time.
@@
 To get the panel plotting the context data stored in CrateDB, the reader must specify a query to retrieve the data from the database.
-This can be done in the \enquote{Query} tab below the graph shown in picture \ref{fig:panel_setup}: here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should look as follows in example \ref{ej:CrateDB_query}
+This can be done in the \enquote{Query} tab below the graph shown in figure \ref{fig:panel_setup}: here the user could use the graphical interface or write the query manually by clicking on the pencil button at the right, so essentially the query should look as follows in example \ref{ej:CrateDB_query}
 
@@
 
-Grafana has a wide options set intended to monitor different kind of systems, it can plot data in a variety of graphics and it even allows exporting to csv format, this can be useful for a more deep analysis using other technologies like, for instance, R, a well-known language widely used for statistical analysis.
+Grafana has a wide options set intended to monitor different kind of systems, it can plot data in a variety of graphics and it even allows exporting to csv format, this can be useful for a more deep analysis using other software packages like, for instance, R, a well-known language widely used for statistical analysis.
 
 Of course, this context data could have been acquired directly from CrateDB through some custom application developed specifically for this task, however this is beyond the purpose of this project.
-Perhaps in a future, a feature like this could be implemented.
+In a future, a feature like this could be implemented.
 
@@
 
-When the requirements are satisfied and all docker containers are properly running\footnote{In a first place the GUI will retrieve the versions for Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap, so if they are not running the GUI will not open.} the GUI can be run performing \cmd{./api-gui.py}; this should open the window shown by figure \ref{fig:GUI_app_keys}
+When the requirements are satisfied and all docker containers are properly running\footnote{In a first place the GUI will retrieve the versions for Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap, so if they are not running the GUI will not open.} the GUI can be run typing \cmd{./api-gui.py}; this should open the window shown in figure \ref{fig:GUI_app_keys}
 
@@
 
-To check Orion Context Broker, IoTagent-LoRaWAN and QuantumLeap versions running in the docker containers the reader has to click on \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without completing all the required info the \enquote{Devices Management} tab will not be enabled.
+To check the Orion Context Broker, the IoTagent-LoRaWAN and the QuantumLeap versions running in the docker containers the reader has to click on the \enquote{About} tab, however, this first shown tab ---\enquote{Application Keys}--- is the most important one because without completing all the required information the \enquote{Devices Management} tab will not be enabled.
 Actually this tab performs a series of checks regarding TTN keys.
@@
 This tab provides a user-friendly interface to manage devices.
-In a similar way, this info will must be the same than the info provided by TTN interface, so as in figure \ref{fig:device_management_tab}
+In a similar way, this info must be the same than the info provided by the TTN interface, so as in figure \ref{fig:device_management_tab}.
 
@@
 This view allows the user to create and delete devices ---along with its entities and subscriptions--- nevertheless, the \enquote{Delete device and entities} button remains disabled until a device is chosen in the right list.
-In a similar way than \enquote{Applications Keys}, to add a new device to the FIWARE instance deployed by docker, all keys must fullfil a set of minimum requirements; as it can be seen in picture \ref{fig:adding_dev}
+In a similar way than \enquote{Applications Keys}, to add a new device to the FIWARE instance deployed by docker, all keys must fullfil a set of minimum requirements; as it can be seen in figure \ref{fig:adding_dev}
 
 \begin{itemize}
-  \item Device entity and ID must be also unique and it has to be at least 3 characters long. In fact, this ID must match the chosen ID in TTN interface when the device is created,
+  \item Device entity and ID must also be unique and it has to be at least 3 characters long. In fact, this ID must match the chosen ID in the TTN interface when the device is created.
   \item Application Session Key must be 32 characters long and also an hexadecimal string.
@@
 These keys can be obtained from TTN as it can be seen in screenshot \ref{fig:dev_overview} ---this screenshot shows info about \cmd{node\_0}, though.
-We can see in picture \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because the final \cmd{Z} letter in the string does not belong to hexadecimal characters set.
+We can see in figure \ref{fig:adding_dev}, that there was a problem with \enquote{Device EUI}, this was caused because the final \cmd{Z} letter in the string does not belong to hexadecimal characters set.
 
@@
 
-Finally, as indicated, the \enquote{About} tab shows the versions info as it can be seen in \ref{fig:about_tab}.
-The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should be always the latests, this is why retrieving them could be helpful.
+Finally, as indicated, the \enquote{About} tab shows the versions info as it can be seen in figure \ref{fig:about_tab}.
+The QuantumLeap version is specified in \cmd{docker-compose.yml} file as well as the CrateDB version, this is to ensure compatibility between QuantumLeap and CrateDB; however, Orion Context Broker and IoTagent-LoRaWAN versions are not specified so they should always be the latests, this is why retrieving them could be helpful.
 
@@
 \section{Proof of concept, deployment}
-Due to inconvenient circumstances at early 2020, this section is not as long as initially planned, instead of testing the system in a real production environment, this section will focus on summarizing needed steps/stages to properly deploy the system.
-Although these steps are already separately detailed, this section tries to put them together to give a consistent and practical point of view.
+Due to the unexpected circumstances related with the COVID19 pandemic during early 2020, this section is not as long as initially planned.
+Instead of testing the system in a real production environment, this section will focus on summarizing needed steps/stages to properly deploy the system.
+Although these steps had already been separately detailed, this section tries to put them together to give a consistent and practical point of view.
 
@@
 \subsection{Set up the gateway}
-This process has been widely detailed in \ref{sssec:Gateway} subsection.
-The most important point on this stage is to ensure the systemd service is working properly and to retrieve the Gateway ID, which will be needed in the next stage.
+This process has been widely detailed in the \ref{sssec:Gateway} subsection.
+The most important point in this stage is to ensure that the \cmd{systemd} service is working properly and to retrieve the Gateway ID, which will be needed in the next stage.
 
 \subsection{TTN Stack}
-Then, the reader should follow the three steps described in subsection \ref{sssec:TTN_Stack}: register a Gateway using its \enquote{Gateway ID}, add an application and register a device for that application.
+After setting up the gateway, the reader should follow the three steps described in subsection \ref{sssec:TTN_Stack}: register a Gateway using its \enquote{Gateway ID}, add an application and register a device for that application.
 This is the first step because the reader will need the security keys for the two following stages:
@@
 \begin{itemize}
-  \item Embed these keys in Arduino sketch. This sketch will require the \enquote{Application Session Key}, \enquote{Network Session Key} and \enquote{Device Address}.
-  \item Use them to create devices, and subscriptions in FIWARE.
-To create the device ---and also its corresponding entity and subscription when using the designed graphical interface shown in \ref{sssec:API_GUI} subsection--- the IoTagent will require the \enquote{Application ID}, \enquote{Application Access Key}, \enquote{Device ID}, \enquote{Application EUI} and again \enquote{Application Session Key}.
+  \item Embed these keys in the Arduino sketch. This sketch will require the \enquote{Application Session Key}, \enquote{Network Session Key} and \enquote{Device Address}.
+  \item Use them to create devices and subscriptions in FIWARE.
+To create the device ---and also its corresponding entity and subscription when using the designed graphical interface shown in the \ref{sssec:API_GUI} subsection--- the IoTagent will require the \enquote{Application ID}, \enquote{Application Access Key}, \enquote{Device ID}, \enquote{Application EUI} and again \enquote{Application Session Key}.
 \end{itemize}
 
-We must remember also that it is very important to choose the CayenneLPP as the payload decoder. So if these keys are available in TTN platform, the process can continue.
+We must remember also that it is very important to choose the CayenneLPP as the payload decoder. So if these keys are available in the TTN platform, the process can continue.
 
 \subsection{Customize the Arduino sketch for a specific device}
-As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}, this latter is the most important one, because several nodes can use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check that every newly created node has a different one in TTN platform.
+As indicated, this sketch will need the \enquote{Application Session Key}, the \enquote{Network Session Key} and \textbf{\enquote{Device Address}}.
+This latter is the most important one, because several nodes can use the same \enquote{Application Session Key} and \enquote{Network Session Key}, but the \enquote{Device Address} must be unique ---it is in fact also provided by TTN and the reader can check that every newly created node has a different one in TTN platform.
 
 Of course, to be able to compile the sketch some dependencies must be satisfied, such as CayenneLPP, ArduinoJson and lmic-arduino libraries.
-Moreover, the \textbf{linear potentiometer has been replaced for a rotary potentiometer}, again due to circumstances in the moment when the project is developed.
+Moreover, the \textbf{linear potentiometer has been replaced for a rotary potentiometer}, again due to the aforementioned circumstances in the moment when the project was developed.
 
@@
 
-It is important to note that analog pin 2 has been chosen the for analog signal acquisition.
-Besides, the reader can check the status of the device in the serial monitor, however, the reader may also note that serial monitor is setup at 115200 bits per second.
+It is important to note that analog pin 2 has been chosen for analog signal acquisition.
+Besides, the reader can check the status of the device in the serial monitor, however, the reader may also note the serial monitor is setup at 115200 bits per second.
 
 \subsection{Deploy Docker containers}
-Once the hardware has been properly setup, the next step is to deploy locally all Docker containers, as indicated in \ref{sssec:FIWARE} subsection.
+Once the hardware has been properly setup, the next step is to deploy locally all Docker containers, as indicated in the \ref{sssec:FIWARE} subsection.
 Using \cmd{docker-compose} this becomes an easy task.
 
-This will deploy essentially the IoTagent-LoRaWAN, Orion CB, QuantumLeap and Grafana. The status of these containers can be checked performing \cmd{docker ps -a}.
+This will deploy essentially the IoTagent-LoRaWAN, Orion CB, QuantumLeap and Grafana. The status of these containers can be checked typing \cmd{docker ps -a}.
 
 \subsection{Devices, entities and subscriptions}
-When Docker containers are deployed, it should be possible to send HTTP requests to create a device (along its entity) and a subscription to notify QuantumLeap any change in any entity.
+When Docker containers are deployed, it should be possible to send HTTP requests to create a device (along with its entity) and a subscription to notify QuantumLeap any change in any entity.
 This process can be done through two different ways as indicated previously,
@@
 \begin{itemize}
-  \item Using developed graphical user interface (recommended, see \ref{sssec:API_GUI} subsection).
+  \item Using the developed graphical user interface (recommended, see \ref{sssec:API_GUI} subsection).
   \item Using \cmd{curl} or Postman to perform the HTTP requests directly.
@@
 \subsection{Grafana}
-Once any device (along its entity) and particularly its subscription are properly created, CrateDB should start journaling the context data, therefore next step is to add CrateDB as a data source in Grafana.
+Once any device (along with its entity) and particularly its subscription are properly created, CrateDB should start journaling the context data, therefore next step is to add CrateDB as a data source in Grafana.
 
@@
     \begin{itemize}
-      \item In a first place, as indicated in its corresponding section, these kind of gateways are LoRaWAN compatible but not compliant.
+      \item In a first place, as indicated in its corresponding section, this kind of gateways are LoRaWAN compatible but not compliant.
 So it would be desirable to replace this HAT by another one, or simply to acquire a proper gateway for a production environment.
@@
 In fact, TTN offers the possibility to reset the frame counter in TTN console but this could be removed in the future.
-    \item Maybe the developed GUI could have handled also the context data plots, however this task is left as an objective to be addressed in future researches, so this is not considered as part of the project because Grafana is completely capable to handle this for now and perhaps forever because it is extremely powerful handling databases.
-    \item For a production environment it seems clear that Arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not hard because of the wide offer for cases and power supply systems.
+    \item The developed GUI could have handled also the context data plots, however this task is left as an objective to be addressed in future developments, so this is not considered as part of the project because Grafana is completely capable to handle this for now and perhaps forever because it is extremely powerful handling databases.
+    \item For a production environment it seems clear that Arduino devices should be encapsulated and the sensor (potentiometer) part must be improved; however this is not difficult because of the wide offer for cases and power supply systems.
 This would lead to a very compact device which could be also improved adding more sensors due to CayenneLPP encoding standard's versatility which allows sending multiple measurements of different sensors.
-    \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; it has potential to be developed further, depending on the needs of the researches.
+    \item Finally this project is a good candidate to replace expensive commercial and proprietary systems; it has potential to be developed further, depending on the needs of the researchers.
 \end{itemize}
